### 2015-05-11 Preface/abstract.tex
Starting Framework

<!DOCTYPE html>
<html lang="en" class="">
  <head prefix="og: http://ogp.me/ns# fb: http://ogp.me/ns/fb# object: http://ogp.me/ns/object# article: http://ogp.me/ns/article# profile: http://ogp.me/ns/profile#">
    <meta charset='utf-8'>
    <meta http-equiv="X-UA-Compatible" content="IE=edge">
    <meta http-equiv="Content-Language" content="en">

    <title>Dissertation/abstract.tex at master · srvanderplas/Dissertation · GitHub</title>
    <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub">
    <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub">
    <link rel="apple-touch-icon" sizes="57x57" href="/apple-touch-icon-114.png">
    <link rel="apple-touch-icon" sizes="114x114" href="/apple-touch-icon-114.png">
    <link rel="apple-touch-icon" sizes="72x72" href="/apple-touch-icon-144.png">
    <link rel="apple-touch-icon" sizes="144x144" href="/apple-touch-icon-144.png">
    <meta property="fb:app_id" content="1401488693436528">

      <meta content="@github" name="twitter:site" /><meta content="summary" name="twitter:card" /><meta content="srvanderplas/Dissertation" name="twitter:title" /><meta content="Dissertation on Graphics, Perception, and Statistics" name="twitter:description" /><meta content="https://avatars0.githubusercontent.com/u/2127125?v=3&amp;s=400" name="twitter:image:src" />
      <meta content="GitHub" property="og:site_name" /><meta content="object" property="og:type" /><meta content="https://avatars0.githubusercontent.com/u/2127125?v=3&amp;s=400" property="og:image" /><meta content="srvanderplas/Dissertation" property="og:title" /><meta content="https://github.com/srvanderplas/Dissertation" property="og:url" /><meta content="Dissertation on Graphics, Perception, and Statistics" property="og:description" />
      <meta name="browser-stats-url" content="https://api.github.com/_private/browser/stats">
    <meta name="browser-errors-url" content="https://api.github.com/_private/browser/errors">
    <link rel="assets" href="https://assets-cdn.github.com/">

    <meta name="pjax-timeout" content="1000">

    <meta name="msapplication-TileImage" content="/windows-tile.png">
    <meta name="msapplication-TileColor" content="#ffffff">
    <meta name="selected-link" value="repo_source" data-pjax-transient>
      <meta name="google-analytics" content="UA-3769691-2">

    <meta content="collector.githubapp.com" name="octolytics-host" /><meta content="collector-cdn.github.com" name="octolytics-script-host" /><meta content="github" name="octolytics-app-id" /><meta content="43035E75:73EA:13DF9794:554FFB32" name="octolytics-dimension-request_id" />

    <meta content="Rails, view, blob#show" name="analytics-event" />
    <meta class="js-ga-set" name="dimension1" content="Logged Out">
    <meta class="js-ga-set" name="dimension2" content="Header v3">
    <meta name="is-dotcom" content="true">
    <meta name="hostname" content="github.com">
    <meta name="user-login" content="">

    <link rel="icon" type="image/x-icon" href="https://assets-cdn.github.com/favicon.ico">

    <meta content="authenticity_token" name="csrf-param" />
<meta content="46M6nZd8O41FepIhcHI4ZYCsjSe6ESsDVe/1vz8VKLThlzMgmhgfGyKSaywXqI6zjmHByvz2Zb6FXf7qph0hug==" name="csrf-token" />

    <link href="https://assets-cdn.github.com/assets/github-c486157afcc5f58155a921bc675afb08733fbaa8dcf39ac2104d381dd9c82ac2.css" media="all" rel="stylesheet" />
    <link href="https://assets-cdn.github.com/assets/github2-da2e842cc3f0aaf33b727d0ef034243c12ab008fd09b24868b97719683b40ee7.css" media="all" rel="stylesheet" />

    <meta http-equiv="x-pjax-version" content="4426702614c8182f33d1780ad1169662">

  <meta name="description" content="Dissertation on Graphics, Perception, and Statistics">
  <meta name="go-import" content="github.com/srvanderplas/Dissertation git https://github.com/srvanderplas/Dissertation.git">

  <meta content="2127125" name="octolytics-dimension-user_id" /><meta content="srvanderplas" name="octolytics-dimension-user_login" /><meta content="17874215" name="octolytics-dimension-repository_id" /><meta content="srvanderplas/Dissertation" name="octolytics-dimension-repository_nwo" /><meta content="true" name="octolytics-dimension-repository_public" /><meta content="false" name="octolytics-dimension-repository_is_fork" /><meta content="17874215" name="octolytics-dimension-repository_network_root_id" /><meta content="srvanderplas/Dissertation" name="octolytics-dimension-repository_network_root_nwo" />
  <link href="https://github.com/srvanderplas/Dissertation/commits/master.atom" rel="alternate" title="Recent Commits to Dissertation:master" type="application/atom+xml">

  </head>

  <body class="logged_out  env-production windows vis-public page-blob">
    <a href="#start-of-content" tabindex="1" class="accessibility-aid js-skip-to-content">Skip to content</a>
    <div class="wrapper">

        <div class="header header-logged-out" role="banner">
  <div class="container clearfix">

    <a class="header-logo-wordmark" href="https://github.com/" data-ga-click="(Logged out) Header, go to homepage, icon:logo-wordmark">
      <span class="mega-octicon octicon-logo-github"></span>
    </a>

    <div class="header-actions" role="navigation">
        <a class="btn btn-primary" href="/join" data-ga-click="(Logged out) Header, clicked Sign up, text:sign-up">Sign up</a>
      <a class="btn" href="/login?return_to=%2Fsrvanderplas%2FDissertation%2Fblob%2Fmaster%2FPreface%2Fabstract.tex" data-ga-click="(Logged out) Header, clicked Sign in, text:sign-in">Sign in</a>
    </div>

    <div class="site-search repo-scope js-site-search" role="search">
      <form accept-charset="UTF-8" action="/srvanderplas/Dissertation/search" class="js-site-search-form" data-global-search-url="/search" data-repo-search-url="/srvanderplas/Dissertation/search" method="get"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /></div>
  <input type="text"
    class="js-site-search-field is-clearable"
    data-hotkey="s"
    name="q"
    placeholder="Search"
    data-global-scope-placeholder="Search GitHub"
    data-repo-scope-placeholder="Search"
    tabindex="1"
    autocapitalize="off">
  <div class="scope-badge">This repository</div>
</form>
    </div>

      <ul class="header-nav left" role="navigation">
          <li class="header-nav-item">
            <a class="header-nav-link" href="/explore" data-ga-click="(Logged out) Header, go to explore, text:explore">Explore</a>
          </li>
          <li class="header-nav-item">
            <a class="header-nav-link" href="/features" data-ga-click="(Logged out) Header, go to features, text:features">Features</a>
          </li>
          <li class="header-nav-item">
            <a class="header-nav-link" href="https://enterprise.github.com/" data-ga-click="(Logged out) Header, go to enterprise, text:enterprise">Enterprise</a>
          </li>
          <li class="header-nav-item">
            <a class="header-nav-link" href="/blog" data-ga-click="(Logged out) Header, go to blog, text:blog">Blog</a>
          </li>
      </ul>

  </div>
</div>

      <div id="start-of-content" class="accessibility-aid"></div>
          <div class="site" itemscope itemtype="http://schema.org/WebPage">
    <div id="js-flash-container">

    </div>
    <div class="pagehead repohead instapaper_ignore readability-menu">
      <div class="container">

<ul class="pagehead-actions">

  <li>
      <a href="/login?return_to=%2Fsrvanderplas%2FDissertation"
    class="btn btn-sm btn-with-count tooltipped tooltipped-n"
    aria-label="You must be signed in to watch a repository" rel="nofollow">
    <span class="octicon octicon-eye"></span>
    Watch
  </a>
  <a class="social-count" href="/srvanderplas/Dissertation/watchers">
    2
  </a>

  </li>

  <li>
      <a href="/login?return_to=%2Fsrvanderplas%2FDissertation"
    class="btn btn-sm btn-with-count tooltipped tooltipped-n"
    aria-label="You must be signed in to star a repository" rel="nofollow">
    <span class="octicon octicon-star"></span>
    Star
  </a>

    <a class="social-count js-social-count" href="/srvanderplas/Dissertation/stargazers">
      1
    </a>

  </li>

    <li>
      <a href="/login?return_to=%2Fsrvanderplas%2FDissertation"
        class="btn btn-sm btn-with-count tooltipped tooltipped-n"
        aria-label="You must be signed in to fork a repository" rel="nofollow">
        <span class="octicon octicon-repo-forked"></span>
        Fork
      </a>
      <a href="/srvanderplas/Dissertation/network" class="social-count">
        1
      </a>
    </li>
</ul>

        <h1 itemscope itemtype="http://data-vocabulary.org/Breadcrumb" class="entry-title public">
          <span class="mega-octicon octicon-repo"></span>
          <span class="author"><a href="/srvanderplas" class="url fn" itemprop="url" rel="author"><span itemprop="title">srvanderplas</span></a></span><!--
       --><span class="path-divider">/</span><!--
       --><strong><a href="/srvanderplas/Dissertation" class="js-current-repository" data-pjax="#js-repo-pjax-container">Dissertation</a></strong>

          <span class="page-context-loader">
            <img alt="" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
          </span>

        </h1>
      </div><!-- /.container -->
    </div><!-- /.repohead -->

    <div class="container">
      <div class="repository-with-sidebar repo-container new-discussion-timeline  ">
        <div class="repository-sidebar clearfix">

<nav class="sunken-menu repo-nav js-repo-nav js-sidenav-container-pjax js-octicon-loaders"
     role="navigation"
     data-pjax="#js-repo-pjax-container"
     data-issue-count-url="/srvanderplas/Dissertation/issues/counts">
  <ul class="sunken-menu-group">
    <li class="tooltipped tooltipped-w" aria-label="Code">
      <a href="/srvanderplas/Dissertation" aria-label="Code" class="selected js-selected-navigation-item sunken-menu-item" data-hotkey="g c" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches /srvanderplas/Dissertation">
        <span class="octicon octicon-code"></span> <span class="full-word">Code</span>
        <img alt="" class="mini-loader" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
</a>    </li>

      <li class="tooltipped tooltipped-w" aria-label="Issues">
        <a href="/srvanderplas/Dissertation/issues" aria-label="Issues" class="js-selected-navigation-item sunken-menu-item" data-hotkey="g i" data-selected-links="repo_issues repo_labels repo_milestones /srvanderplas/Dissertation/issues">
          <span class="octicon octicon-issue-opened"></span> <span class="full-word">Issues</span>
          <span class="js-issue-replace-counter"></span>
          <img alt="" class="mini-loader" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
</a>      </li>

    <li class="tooltipped tooltipped-w" aria-label="Pull requests">
      <a href="/srvanderplas/Dissertation/pulls" aria-label="Pull requests" class="js-selected-navigation-item sunken-menu-item" data-hotkey="g p" data-selected-links="repo_pulls /srvanderplas/Dissertation/pulls">
          <span class="octicon octicon-git-pull-request"></span> <span class="full-word">Pull requests</span>
          <span class="js-pull-replace-counter"></span>
          <img alt="" class="mini-loader" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
</a>    </li>

  </ul>
  <div class="sunken-menu-separator"></div>
  <ul class="sunken-menu-group">

    <li class="tooltipped tooltipped-w" aria-label="Pulse">
      <a href="/srvanderplas/Dissertation/pulse" aria-label="Pulse" class="js-selected-navigation-item sunken-menu-item" data-selected-links="pulse /srvanderplas/Dissertation/pulse">
        <span class="octicon octicon-pulse"></span> <span class="full-word">Pulse</span>
        <img alt="" class="mini-loader" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
</a>    </li>

    <li class="tooltipped tooltipped-w" aria-label="Graphs">
      <a href="/srvanderplas/Dissertation/graphs" aria-label="Graphs" class="js-selected-navigation-item sunken-menu-item" data-selected-links="repo_graphs repo_contributors /srvanderplas/Dissertation/graphs">
        <span class="octicon octicon-graph"></span> <span class="full-word">Graphs</span>
        <img alt="" class="mini-loader" height="16" src="https://assets-cdn.github.com/assets/spinners/octocat-spinner-32-e513294efa576953719e4e2de888dd9cf929b7d62ed8d05f25e731d02452ab6c.gif" width="16" />
</a>    </li>
  </ul>

</nav>

              <div class="only-with-full-nav">

<div class="clone-url open"
  data-protocol-type="http"
  data-url="/users/set_protocol?protocol_selector=http&amp;protocol_type=clone">
  <h3><span class="text-emphasized">HTTPS</span> clone URL</h3>
  <div class="input-group js-zeroclipboard-container">
    <input type="text" class="input-mini input-monospace js-url-field js-zeroclipboard-target"
           value="https://github.com/srvanderplas/Dissertation.git" readonly="readonly">
    <span class="input-group-button">
      <button aria-label="Copy to clipboard" class="js-zeroclipboard btn btn-sm zeroclipboard-button tooltipped tooltipped-s" data-copied-hint="Copied!" type="button"><span class="octicon octicon-clippy"></span></button>
    </span>
  </div>
</div>

<div class="clone-url "
  data-protocol-type="subversion"
  data-url="/users/set_protocol?protocol_selector=subversion&amp;protocol_type=clone">
  <h3><span class="text-emphasized">Subversion</span> checkout URL</h3>
  <div class="input-group js-zeroclipboard-container">
    <input type="text" class="input-mini input-monospace js-url-field js-zeroclipboard-target"
           value="https://github.com/srvanderplas/Dissertation" readonly="readonly">
    <span class="input-group-button">
      <button aria-label="Copy to clipboard" class="js-zeroclipboard btn btn-sm zeroclipboard-button tooltipped tooltipped-s" data-copied-hint="Copied!" type="button"><span class="octicon octicon-clippy"></span></button>
    </span>
  </div>
</div>

<p class="clone-options">You can clone with
  <a href="#" class="js-clone-selector" data-protocol="http">HTTPS</a> or <a href="#" class="js-clone-selector" data-protocol="subversion">Subversion</a>.
  <a href="https://help.github.com/articles/which-remote-url-should-i-use" class="help tooltipped tooltipped-n" aria-label="Get help on which URL is right for you.">
    <span class="octicon octicon-question"></span>
  </a>
</p>

  <a href="https://windows.github.com" class="btn btn-sm sidebar-button" title="Save srvanderplas/Dissertation to your computer and use it in GitHub Desktop." aria-label="Save srvanderplas/Dissertation to your computer and use it in GitHub Desktop.">
    <span class="octicon octicon-device-desktop"></span>
    Clone in Desktop
  </a>

                <a href="/srvanderplas/Dissertation/archive/master.zip"
                   class="btn btn-sm sidebar-button"
                   aria-label="Download the contents of srvanderplas/Dissertation as a zip file"
                   title="Download the contents of srvanderplas/Dissertation as a zip file"
                   rel="nofollow">
                  <span class="octicon octicon-cloud-download"></span>
                  Download ZIP
                </a>
              </div>
        </div><!-- /.repository-sidebar -->

        <div id="js-repo-pjax-container" class="repository-content context-loader-container" data-pjax-container>

<a href="/srvanderplas/Dissertation/blob/a3e0da906f9ab58ab9afcbd02db434d83beee7d4/Preface/abstract.tex" class="hidden js-permalink-shortcut" data-hotkey="y">Permalink</a>

<!-- blob contrib key: blob_contributors:v21:134af7cd8930ec141f2b094aec07169d -->

<div class="file-navigation js-zeroclipboard-container">

<div class="select-menu js-menu-container js-select-menu left">
  <span class="btn btn-sm select-menu-button js-menu-target css-truncate" data-hotkey="w"
    data-master-branch="master"
    data-ref="master"
    title="master"
    role="button" aria-label="Switch branches or tags" tabindex="0" aria-haspopup="true">
    <span class="octicon octicon-git-branch"></span>
    <i>branch:</i>
    <span class="js-select-button css-truncate-target">master</span>
  </span>

  <div class="select-menu-modal-holder js-menu-content js-navigation-container" data-pjax aria-hidden="true">

    <div class="select-menu-modal">
      <div class="select-menu-header">
        <span class="select-menu-title">Switch branches/tags</span>
        <span class="octicon octicon-x js-menu-close" role="button" aria-label="Close"></span>
      </div>

      <div class="select-menu-filters">
        <div class="select-menu-text-filter">
          <input type="text" aria-label="Filter branches/tags" id="context-commitish-filter-field" class="js-filterable-field js-navigation-enable" placeholder="Filter branches/tags">
        </div>
        <div class="select-menu-tabs">
          <ul>
            <li class="select-menu-tab">
              <a href="#" data-tab-filter="branches" data-filter-placeholder="Filter branches/tags" class="js-select-menu-tab">Branches</a>
            </li>
            <li class="select-menu-tab">
              <a href="#" data-tab-filter="tags" data-filter-placeholder="Find a tag…" class="js-select-menu-tab">Tags</a>
            </li>
          </ul>
        </div>
      </div>

      <div class="select-menu-list select-menu-tab-bucket js-select-menu-tab-bucket" data-tab-filter="branches">

        <div data-filterable-for="context-commitish-filter-field" data-filterable-type="substring">

            <a class="select-menu-item js-navigation-item js-navigation-open selected"
               href="/srvanderplas/Dissertation/blob/master/Preface/abstract.tex"
               data-name="master"
               data-skip-pjax="true"
               rel="nofollow">
              <span class="select-menu-item-icon octicon octicon-check"></span>
              <span class="select-menu-item-text css-truncate-target" title="master">
                master
              </span>
            </a>
            <a class="select-menu-item js-navigation-item js-navigation-open "
               href="/srvanderplas/Dissertation/blob/srvanderplas/SineIllusion/Submission2/SineIllusion-Rearrange.Rnw/Preface/abstract.tex"
               data-name="srvanderplas/SineIllusion/Submission2/SineIllusion-Rearrange.Rnw"
               data-skip-pjax="true"
               rel="nofollow">
              <span class="select-menu-item-icon octicon octicon-check"></span>
              <span class="select-menu-item-text css-truncate-target" title="srvanderplas/SineIllusion/Submission2/SineIllusion-Rearrange.Rnw">
                srvanderplas/SineIllusion/Submission2/SineIllusion-Rearrange.Rnw
              </span>
            </a>
        </div>

          <div class="select-menu-no-results">Nothing to show</div>
      </div>

      <div class="select-menu-list select-menu-tab-bucket js-select-menu-tab-bucket" data-tab-filter="tags">
        <div data-filterable-for="context-commitish-filter-field" data-filterable-type="substring">

        </div>

        <div class="select-menu-no-results">Nothing to show</div>
      </div>

    </div>
  </div>
</div>

  <div class="btn-group right">
    <a href="/srvanderplas/Dissertation/find/master"
          class="js-show-file-finder btn btn-sm empty-icon tooltipped tooltipped-s"
          data-pjax
          data-hotkey="t"
          aria-label="Quickly jump between files">
      <span class="octicon octicon-list-unordered"></span>
    </a>
    <button aria-label="Copy file path to clipboard" class="js-zeroclipboard btn btn-sm zeroclipboard-button tooltipped tooltipped-s" data-copied-hint="Copied!" type="button"><span class="octicon octicon-clippy"></span></button>
  </div>

  <div class="breadcrumb js-zeroclipboard-target">
    <span class='repo-root js-repo-root'><span itemscope="" itemtype="http://data-vocabulary.org/Breadcrumb"><a href="/srvanderplas/Dissertation" class="" data-branch="master" data-direction="back" data-pjax="true" itemscope="url"><span itemprop="title">Dissertation</span></a></span></span><span class="separator">/</span><span itemscope="" itemtype="http://data-vocabulary.org/Breadcrumb"><a href="/srvanderplas/Dissertation/tree/master/Preface" class="" data-branch="master" data-direction="back" data-pjax="true" itemscope="url"><span itemprop="title">Preface</span></a></span><span class="separator">/</span><strong class="final-path">abstract.tex</strong>
  </div>
</div>

  <div class="commit file-history-tease">
    <div class="file-history-tease-header">
        <img alt="@srvanderplas" class="avatar" data-user="2127125" height="24" src="https://avatars1.githubusercontent.com/u/2127125?v=3&amp;s=48" width="24" />
        <span class="author"><a href="/srvanderplas" rel="author">srvanderplas</a></span>
        <time datetime="2015-04-16T18:38:07Z" is="relative-time">Apr 16, 2015</time>
        <div class="commit-title">
            <a href="/srvanderplas/Dissertation/commit/edbc4ab89617eddc65e6287ffb64566e04e165ed" class="message" data-pjax="true" title="Smaller abstract to conform to guidelines">Smaller abstract to conform to guidelines</a>
        </div>
    </div>

    <div class="participation">
      <p class="quickstat">
        <a href="#blob_contributors_box" rel="facebox">
          <strong>1</strong>
           contributor
        </a>
      </p>

    </div>
    <div id="blob_contributors_box" style="display:none">
      <h2 class="facebox-header">Users who have contributed to this file</h2>
      <ul class="facebox-user-list">
          <li class="facebox-user-list-item">
            <img alt="@srvanderplas" data-user="2127125" height="24" src="https://avatars1.githubusercontent.com/u/2127125?v=3&amp;s=48" width="24" />
            <a href="/srvanderplas">srvanderplas</a>
          </li>
      </ul>
    </div>
  </div>

<div class="file">
  <div class="file-header">
    <div class="file-actions">

      <div class="btn-group">
        <a href="/srvanderplas/Dissertation/raw/master/Preface/abstract.tex" class="btn btn-sm " id="raw-url">Raw</a>
          <a href="/srvanderplas/Dissertation/blame/master/Preface/abstract.tex" class="btn btn-sm js-update-url-with-hash">Blame</a>
        <a href="/srvanderplas/Dissertation/commits/master/Preface/abstract.tex" class="btn btn-sm " rel="nofollow">History</a>
      </div>

        <a class="octicon-btn tooltipped tooltipped-nw"
           href="https://windows.github.com"
           aria-label="Open this file in GitHub for Windows"
           data-ga-click="Repository, open with desktop, type:windows">
            <span class="octicon octicon-device-desktop"></span>
        </a>

          <button type="button" class="octicon-btn disabled tooltipped tooltipped-n" aria-label="You must be signed in to make or propose changes">
            <span class="octicon octicon-pencil"></span>
          </button>

        <button type="button" class="octicon-btn octicon-btn-danger disabled tooltipped tooltipped-n" aria-label="You must be signed in to make or propose changes">
          <span class="octicon octicon-trashcan"></span>
        </button>
    </div>

    <div class="file-info">
        18 lines (14 sloc)
        <span class="file-info-divider"></span>
      5.283 kb
    </div>
  </div>

  <div class="blob-wrapper data type-tex">
      <table class="highlight tab-size-8 js-file-line-container">
      <tr>
        <td id="L1" class="blob-num js-line-number" data-line-number="1"></td>
        <td id="LC1" class="blob-code blob-code-inner js-file-line"><span class="pl-c1">\specialchapt</span>{ABSTRACT}</td>
      </tr>
      <tr>
        <td id="L2" class="blob-num js-line-number" data-line-number="2"></td>
        <td id="LC2" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L3" class="blob-num js-line-number" data-line-number="3"></td>
        <td id="LC3" class="blob-code blob-code-inner js-file-line">There has been quite a bit of research on statistical graphics and visualization, generally focused on new types of graphics, new software to create graphics, interactivity, and usability studies. Our ability to interpret and use statistical graphics hinges on the interface between the graph itself and the brain that perceives and interprets it, and there is substantially less research on the interplay between graph, eye, brain, and mind than is sufficient to understand the nature of these relationships. </td>
      </tr>
      <tr>
        <td id="L4" class="blob-num js-line-number" data-line-number="4"></td>
        <td id="LC4" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L5" class="blob-num js-line-number" data-line-number="5"></td>
        <td id="LC5" class="blob-code blob-code-inner js-file-line">The goal of the work presented here is to further explore the interplay between a static graph, the translation of that graph from paper to mental representation (the journey from eye to brain), and the mental processes that operate on that graph once it is transferred into memory (mind). Understanding the perception of statistical graphics should allow researchers to create more effective graphs which produce fewer distortions and viewer errors while reducing the cognitive load necessary to understand the information presented in the graph. </td>
      </tr>
      <tr>
        <td id="L6" class="blob-num js-line-number" data-line-number="6"></td>
        <td id="LC6" class="blob-code blob-code-inner js-file-line"><span class="pl-c">% </span></td>
      </tr>
      <tr>
        <td id="L7" class="blob-num js-line-number" data-line-number="7"></td>
        <td id="LC7" class="blob-code blob-code-inner js-file-line"><span class="pl-c">% Chapter \ref{litreview} contains a review of past literature which is relevant to the encoding and memory of statistical graphics, encompassing studies from psychology, psychophysics, meteorology, business, and statistics. The literature review encompasses many different areas of interdisciplinary research; to ensure that there is a reference for some of the more niche vocabulary in this dissertation, it begins with the physiology of the eye and includes a summary of some basic neuropsychology as well. From there, I discuss some of the psychology and psychophysics research that concerns visual perception, memory, and attention. This research roughly falls in the domain of `cognitive psychology&#39;, but includes studies from fields as diverse as meteorology and business. After establishing the underlying mechanisms of perception, I cover some of the research which focuses specifically on statistical graphics, and discuss some of the methodology used in these experiments.</span></td>
      </tr>
      <tr>
        <td id="L8" class="blob-num js-line-number" data-line-number="8"></td>
        <td id="LC8" class="blob-code blob-code-inner js-file-line"><span class="pl-c">% </span></td>
      </tr>
      <tr>
        <td id="L9" class="blob-num js-line-number" data-line-number="9"></td>
        <td id="LC9" class="blob-code blob-code-inner js-file-line"><span class="pl-c">% </span></td>
      </tr>
      <tr>
        <td id="L10" class="blob-num js-line-number" data-line-number="10"></td>
        <td id="LC10" class="blob-code blob-code-inner js-file-line"><span class="pl-c">% </span></td>
      </tr>
      <tr>
        <td id="L11" class="blob-num js-line-number" data-line-number="11"></td>
        <td id="LC11" class="blob-code blob-code-inner js-file-line"><span class="pl-c">% Chapter \ref{sineillusion} discusses an optical illusion which is frequently present in even simple statistical graphics, but is incredibly difficult to resolve without altering the graph itself. This illusion (and its relative obscurity in the literature on graph perception) serves as a reminder that our graphs are only useful if they are designed with the perceptual system in mind. The presence of this illusion (and the solution for reducing the illusion&#39;s effects) serve as an introduction and case study for whether we can actually trust the things we see; the remainder of this dissertation aims to quantify variables which may affect how we answer the simple question ``What am I seeing?&quot; in relation to statistical graphics. Chapter \ref{liefactor} expands on the discussion of the sine illusion, exploring the perceptual origin of the illusion and presenting a case study of an individual not affected by the illusion.</span></td>
      </tr>
      <tr>
        <td id="L12" class="blob-num js-line-number" data-line-number="12"></td>
        <td id="LC12" class="blob-code blob-code-inner js-file-line"><span class="pl-c">% </span></td>
      </tr>
      <tr>
        <td id="L13" class="blob-num js-line-number" data-line-number="13"></td>
        <td id="LC13" class="blob-code blob-code-inner js-file-line"><span class="pl-c">% The use of statistical graphics is encouraged because graphics summarize statistical models and results in a form that is easy for most people to understand. The lineup protocol provides a convenient way to compare different types of graphics which display the same data, but results from comparative studies\citep{hofmann2012graphical} have demonstrated that individuals are highly variable in their ability to identify the target plot in a lineup. The study presented in Chapter \ref{visualreasoning} aims to explore the association between spatial reasoning, pattern recognition, and the ability to identify a target plot in a lineup successfully.</span></td>
      </tr>
      <tr>
        <td id="L14" class="blob-num js-line-number" data-line-number="14"></td>
        <td id="LC14" class="blob-code blob-code-inner js-file-line"><span class="pl-c">% </span></td>
      </tr>
      <tr>
        <td id="L15" class="blob-num js-line-number" data-line-number="15"></td>
        <td id="LC15" class="blob-code blob-code-inner js-file-line"><span class="pl-c">% Finally, Chapter \ref{featurehierarchy} discusses a study which seeks to understand visually salient features of a plot, to aid researchers in creating graphics which are constructed to efficiently convey information visually. \citet{cleveland:1984} created a hierarchy of graphical tasks which have informed graph design for the past 30 years, ranking numerical estimations of graphical features by accuracy. Similarly, it would be useful to understand what features in a graph will dominate a viewer&#39;s perception: if linear trend information contradicts coloring of points, which feature will a viewer take away from the image? This experiment uses the lineup protocol to examine the features which are most salient to viewers, using pairwise comparisons of shape, color, linear trend, and outliers. </span></td>
      </tr>
      <tr>
        <td id="L16" class="blob-num js-line-number" data-line-number="16"></td>
        <td id="LC16" class="blob-code blob-code-inner js-file-line">
</td>
      </tr>
      <tr>
        <td id="L17" class="blob-num js-line-number" data-line-number="17"></td>
        <td id="LC17" class="blob-code blob-code-inner js-file-line">Taken together, the experiments presented here should lay a foundation for exploring the perception of statistical graphics. There has been considerable research into the accuracy of numerical judgments viewers make from graphs, and these studies are useful, but it is more effective to understand how errors in these judgments occur so that the root cause of the error can be addressed directly. Understanding how visual reasoning relates to the ability to make judgments from graphs allows us to tailor graphics to particular target audiences. In addition, understanding the hierarchy of salient features in statistical graphics allows us to clearly communicate the important message from data or statistical models by constructing graphics which are designed specifically for the perceptual system. </td>
      </tr>
</table>

  </div>

</div>

<a href="#jump-to-line" rel="facebox[.linejump]" data-hotkey="l" style="display:none">Jump to Line</a>
<div id="jump-to-line" style="display:none">
  <form accept-charset="UTF-8" action="" class="js-jump-to-line-form" method="get"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /></div>
    <input class="linejump-input js-jump-to-line-field" type="text" placeholder="Jump to line&hellip;" autofocus>
    <button type="submit" class="btn">Go</button>
</form></div>

        </div>

      </div><!-- /.repo-container -->
      <div class="modal-backdrop"></div>
    </div><!-- /.container -->
  </div><!-- /.site -->

    </div><!-- /.wrapper -->

      <div class="container">
  <div class="site-footer" role="contentinfo">
    <ul class="site-footer-links right">
        <li><a href="https://status.github.com/" data-ga-click="Footer, go to status, text:status">Status</a></li>
      <li><a href="https://developer.github.com" data-ga-click="Footer, go to api, text:api">API</a></li>
      <li><a href="https://training.github.com" data-ga-click="Footer, go to training, text:training">Training</a></li>
      <li><a href="https://shop.github.com" data-ga-click="Footer, go to shop, text:shop">Shop</a></li>
        <li><a href="https://github.com/blog" data-ga-click="Footer, go to blog, text:blog">Blog</a></li>
        <li><a href="https://github.com/about" data-ga-click="Footer, go to about, text:about">About</a></li>

    </ul>

    <a href="https://github.com" aria-label="Homepage">
      <span class="mega-octicon octicon-mark-github" title="GitHub"></span>
</a>
    <ul class="site-footer-links">
      <li>&copy; 2015 <span title="0.03132s from github-fe141-cp1-prd.iad.github.net">GitHub</span>, Inc.</li>
        <li><a href="https://github.com/site/terms" data-ga-click="Footer, go to terms, text:terms">Terms</a></li>
        <li><a href="https://github.com/site/privacy" data-ga-click="Footer, go to privacy, text:privacy">Privacy</a></li>
        <li><a href="https://github.com/security" data-ga-click="Footer, go to security, text:security">Security</a></li>
        <li><a href="https://github.com/contact" data-ga-click="Footer, go to contact, text:contact">Contact</a></li>
    </ul>
  </div>
</div>

    <div class="fullscreen-overlay js-fullscreen-overlay" id="fullscreen_overlay">
  <div class="fullscreen-container js-suggester-container">
    <div class="textarea-wrap">
      <textarea name="fullscreen-contents" id="fullscreen-contents" class="fullscreen-contents js-fullscreen-contents" placeholder=""></textarea>
      <div class="suggester-container">
        <div class="suggester fullscreen-suggester js-suggester js-navigation-container"></div>
      </div>
    </div>
  </div>
  <div class="fullscreen-sidebar">
    <a href="#" class="exit-fullscreen js-exit-fullscreen tooltipped tooltipped-w" aria-label="Exit Zen Mode">
      <span class="mega-octicon octicon-screen-normal"></span>
    </a>
    <a href="#" class="theme-switcher js-theme-switcher tooltipped tooltipped-w"
      aria-label="Switch themes">
      <span class="octicon octicon-color-mode"></span>
    </a>
  </div>
</div>

    <div id="ajax-error-message" class="flash flash-error">
      <span class="octicon octicon-alert"></span>
      <a href="#" class="octicon octicon-x flash-close js-ajax-error-dismiss" aria-label="Dismiss error"></a>
      Something went wrong with that request. Please try again.
    </div>

      <script crossorigin="anonymous" src="https://assets-cdn.github.com/assets/frameworks-06533d99d3f0ca9115a0563dec4017e1bfad7758231e12b95a178cef2cdc3d4c.js"></script>
      <script async="async" crossorigin="anonymous" src="https://assets-cdn.github.com/assets/github-59cbed91abaeebed35de0a12f8c3adfea46b14dd56c5019b18a7dcdaf3c88a92.js"></script>

  </body>
</html>