### 2015-05-11 Preface/acknowl.tex
Add dedication, title page, etc.

\specialchapt{ACKNOWLEDGEMENTS}

First, I would like to thank Dr. Heike Hofmann for the guidance she has provide during my time at Iowa State University. Her patient tutelage has helped me immeasurably in my winding -- often stubbornly veering -- pursuit of a doctorate.

Great appreciation also extends to my Ph.D. committee members, who have graciously given their time and advice. A special thanks to Dr. Bob Stephenson, whom also served as a study advisor and mentor in statistics education.

I would also like to thank my undergraduate advisor, Dr. Jon Anderson, whom sparked my interest in statistics and fostered that early curiosity.

Lastly, I owe an enormous debt of gratitude to my friends and family -- especially my mother and father -- for their love and support throughout my life. Thank you, I love you.