### 2015-05-11 Preface/titlepage.tex

% Template Titlepage File
% \title{This is the title of a thesis
% submitted to Iowa State University\\
% Note that only the first letter of
% the first word and proper names
% are capitalized}
\title{Working Title}
\author{Karsten Tait Maurer}
\degree{DOCTOR OF PHILOSOPHY}
\major{Statistics}
\level{doctoral}
\mprof{Heike Hofmann}
\members{Robert Stephenson \\ Max Morris \\ Mark Kaiser \\ Alejandro Andreotti }
\notice

%comment these out for the actual dissertation :)
\level{doctoral}
\committee{4}
\format{dissertation}
% \submit{my awesome committee}

\maketitle