### 2015-05-11 Preface/dedication.tex

\chapter*{DEDICATION}

I dedicate this thesis to my love, Madeline, for the support she has given me in the pursuit of my goals. I am so excited to marry her this summer and begin our life together. Watch out world, here we come!