### 2015-05-24 DefenseSlides/DefenseSlidesCommitteeHour.Rnw
push defense slides files

\documentclass{beamer} 
% \usepackage{graphicx}
\usepackage{graphics}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{bbm}

\usepackage{amsmath,amsthm,amssymb}
\usepackage{graphicx,float,wrapfig,subfig,tabularx,ulem}
\usepackage{color}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{mathrsfs}

\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
}

\setbeamercovered{transparent}
\renewcommand{\ni}{\noindent}

\newcommand{\distas}[1]{\mathbin{\overset{#1}{\sim}}}%
\newcommand{\Cov}{\text{Cov}}%

<<setup,echo=F,include=F,eval=T>>=
### Preliminaries 
## Load Packages
library(car)
library(ggplot2)
library(subselect)
library(qtlmt)
library(multcomp)
library(lattice)
library(gridExtra)
library(reshape2)
library(plyr)
library(boot)
library(psy)
library(plyr)
library(graphics)
library(xtable)
library(dbData)
library(devtools)
library(RColorBrewer)
library(hexbin)
library(MASS)
library(scales)
@

<<BinnedScatterPrep, echo=FALSE, eval=TRUE, cache=TRUE, include=FALSE>>=
source("BinningLossFunctions.R")
spatlossdatSquareBins <- read.csv("Data/SpatialLossData.csv",header=TRUE)

connect <- dbConnect(dbDriver("MySQL"), host="mysql2.stat.iastate.edu", 
                     port=3306, user="dbaccess", dbname="baseball")
pitch <- new("dataDB", co=connect, table="Pitching")
pitchdat <- dbGetQuery(connect,  sprintf("select G, SO from Pitching"))

d1 <- dbData(pitch, vars=c( "G", "SO")) # Games, Strike Outs

xtd.data <- function(x) {
# replicate each line as often as given by the last entry
  k <- dim(x)[2]

  res <- data.frame(sapply(1:(k-1), function(var) {
    	return(rep(x[,var], x[,k]))
		}
	))
	names(res) <- names(x)[1:(k-1)]
	return(res)
}
d1.ext <- xtd.data(d1)
dbDisconnect(connect)
@

<<surveyprep, echo=FALSE, eval=TRUE, cache=TRUE, include=FALSE>>=
# Preliminaries
dat <- read.csv("Data/SurveyResults.csv",header=T)

#list of questions that were inversely coded
flipcoded <- c(3,5,6,7,10,11)
# recode to match positive worded response
for (i in flipcoded){
  dat[,i] <- 6-dat[,i]
}
dat$student <- 1:nrow(dat)
#centered version
cdat <- dat[1:12]
for(i in 1:ncol(cdat)){
  cdat[,i] <- cdat[,i]-3
}

names(dat) <- c("Eas1","Con1","Eng1","Eas2","Con2","Eng2",
                "Eas3","Con3","Eng3","Eas4","Con4","Eng4",
                "Section","Student")

#reshape data
meltdat <- melt(dat[,c(1:12,14)], measure.vars=c(1:12))
meltdat$set <- factor(rep(rep(c("Ease","Concept","Engagement"),4),each=nrow(dat)),levels=c("Ease","Concept","Engagement"))
meltdat$set4 <- rep(rep(c("Ease","Concept","Engagement"),4),each=nrow(dat))
meltdat$set4[meltdat$variable=="Con1" | meltdat$variable=="Con3"] <- "Positive Concept"
meltdat$set4[meltdat$variable=="Con2" | meltdat$variable=="Con4"] <- "Negative Concept"
meltdat$set4 <- factor(meltdat$set4, levels=c("Ease","Positive Concept","Negative Concept","Engagement"))
meltdat$centered <- as.factor(meltdat$value-3)
meltdat$value <- as.factor(meltdat$value)
meltdat$question.in.set <- rep(paste("Item", 1:4,sep=" "),each=3*nrow(dat))
meltdatNoNA <- meltdat[which(meltdat$value != "NA"),]
#create short DF of summary stats
sumstats <- ddply(meltdat,.(variable),summarize,
                  mean = round(mean(as.numeric(as.character(centered)), na.rm=TRUE),2),
                    sd = round(sd(as.numeric(as.character(centered)),na.rm=T),2) )

# Set up for fluctuation diagrams
ggfluctuation <- function (table, type = "size", floor = 0, ceiling = max(table$freq, na.rm = TRUE)) 
{
#    .Deprecated()
    if (is.table(table)) 
        table <- as.data.frame(t(table))
    oldnames <- names(table)
    names(table) <- c("x", "y", "result")
    table <- transform(table, x = as.factor(x), y = as.factor(y), 
        freq = result)

    if (type == "size") {
        table <- transform(table, freq = sqrt(pmin(freq, ceiling)/ceiling), 
            border = ifelse(is.na(freq), "grey90", ifelse(freq > 
                ceiling, "grey30", "grey50")))
        table[is.na(table$freq), "freq"] <- 1
        table <- subset(table, freq * ceiling >= floor)
    }

    if (type == "size") {
        nx <- length(levels(table$x))
        ny <- length(levels(table$y))
        p <- ggplot(table, aes_string(x = "x", y = "y", height = "freq", 
            width = "freq", fill = "border")) + geom_tile(colour = "white") + 
            scale_fill_identity() + theme(aspect.ratio = ny/nx)
    }

    else {
        p <- ggplot(table, aes_string(x = "x", y = "y", fill = "freq")) + 
        geom_tile(colour = "white") + scale_fill_gradient2(expression(w[ij]), low = "red", mid="white", high = "blue")
    }
    p$xlabel <- oldnames[1]
    p$ylabel <- oldnames[2]
    p
} 
theme_fluct <- theme(panel.background = element_rect(fill = "white", 
                colour = NA), panel.border = element_rect(fill = NA, 
                colour = "grey50"), panel.grid.major = element_line(colour = "grey90", 
                size = 0.2), panel.grid.minor = element_line(colour = "grey98", 
                size = 0.5))

var(cdat[,1]+cdat[,4])
4*cov(cdat[,1],cdat[,4])/var(cdat[,1]+cdat[,4])
cov(cdat[,1],cdat[,4])/ (sd(cdat[,1])*sd(cdat[,4]))
2*(1- (var(cdat[,1])+var(cdat[,4]))/var(cdat[,1]+cdat[,4]))
cor(cdat[,1],cdat[,4])
@

\title[Applications of Technology and Large Data in Statistics Education and Statistical Graphics]{Applications of Technology and Large Data in Statistics Education and Statistical Graphics}
\author[K. Maurer]{Karsten Maurer}
\date{June 4, 2015}
\institute[ISU]{Iowa State University}
\begin{document}

\begin{frame}
    \maketitle
\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Presentation Outline}  
  Quantifying Loss in Binned Scatterplots
    \begin{itemize}
      \item Scatterplot Adaptations
      \item Binning
      \item Loss Functions
      \item Impact of Binning Specification on Loss
      \item Results and Recommendations
    \end{itemize}    
  Shiny Database Sampler
    \begin{itemize}
      \item Development
      \item Applications
      \item Student User Survey
      \item Results
    \end{itemize}

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{}

\begin{center}
\large Loss and Binned Scatterplots\\
\normalsize Chapter 4
\end{center}

\end{frame}

% 
% %-------------------------------------------------
% 
% \begin{frame}
%     \frametitle{Outline}
% Bivariate Data Visualization   
%   \begin{itemize}
%     \item Traditional scatterplot
%     \item Adaptations for large data
%     \item Binned Scatterplots
%   \end{itemize} 
% \pause
% 
% Spatial Binning Algorithms
%   \begin{itemize}
%     \item Loss functions
%     \item Binning Properties
%   \end{itemize} 
% \pause
% 
% Frequency Binning Algorithms
%   \begin{itemize}
%     \item Loss functions
%     \item Binning Properties
%   \end{itemize} 
% \pause
% 
% Recommendations for Specification
% \end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Bivariate Data Visualization }
Example Data from Baseball   
  \begin{itemize}
    \item Sean Lahman's Baseball database (\url{http://www.seanlahman.com/baseball-archive/})
    \item Variables are Games Played and Strike Outs
    \item Observations: 42583 MLB pitcher seasons
    \item Data from 1871-2009
  \end{itemize}
Goal: Visualize relationship between games played and strike outs

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Traditional Scatterplot}
<<TradScatterplot, echo=FALSE , warning=FALSE, fig.width=4, fig.height=4, out.width='.6\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
qplot(data=d1, x=G, y=SO, geom="point", main="Games and Strike Outs in ML Baseball", size=I(2)) + theme_bw() + xlab("Games (Count)")+ ylab("Strike Outs (Count)")
@

\center Heavy Over-plotting

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Adaptations to Scatterplots}
Lots of Recommendations for Dealing with Scatterplot Over-plotting

\vspace{.5cm}

Change Point Rendering
  \begin{itemize}
    \item Open Points
    \item Alpha-Blending
    \item Generalized Scatterplots
  \end{itemize} \pause

Plotting Binned Aggregations
  \begin{itemize}
    \item Sunflower Plots
    \item Bubble Plots
    \item Binned Scatterplots
  \end{itemize}

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Open-Circle Scatterplot}
<<OpenCircleScatterplot, echo=FALSE , warning=FALSE, fig.width=4, fig.height=4, out.width='.6\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
qplot(data=d1, x=G, y=SO, geom="point", main="Games and Strike Outs in ML Baseball", shape="o", size=I(5)) + scale_shape_identity() + theme_bw() + xlab("Games (Count)")+ ylab("Strike Outs (Count)")
@

\center Helps slightly, but still allows for overplotting

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Alpha-Blended Scatterplot}
<<AlphaBlendScatterplot, echo=FALSE , warning=FALSE, fig.width=4, fig.height=4, out.width='.6\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
qplot(data=d1, x=G, y=SO, geom="point", main="Games and Strike Outs in ML Baseball", size=I(2), alpha=I(0.05)) + theme_bw() + xlab("Games (Count)")+ ylab("Strike Outs (Count)")
@

\center Fairly effective, but downplays outliers

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Bubble Plot}
<<BubblePlot, echo=FALSE , warning=FALSE, fig.width=4, fig.height=4, out.width='.6\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
d1.hex.obj <- hexbin(d1.ext$G, d1.ext$SO)
d1.hex <- data.frame(hcell2xy(d1.hex.obj))
names(d1.hex) <- c("G", "SO")
d1.hex$Freq <-  d1.hex.obj@count

qplot(data=d1.hex, x=G, y=SO, geom="point", size=Freq+1) + 
  scale_size_area("Log Frequency", max_size=4, trans="log", breaks=c(1, 10, 100, 1000, 10000))+ theme_bw() + xlab("Games (Count)")+ ylab("Strike Outs (Count)") + theme(legend.position="bottom")
@

\center Mapping frequency to size of point at bin center

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Binned Scatterplot}
<<BinnedScatterplotBaseball, echo=FALSE , warning=FALSE, fig.width=4, fig.height=4, out.width='.6\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
binout1 <- RectBin2d(pitchdat$G,pitchdat$SO,.5,-.5,5,25,type="standard")
qplot(binxs, binys, geom="tile", fill=binfreq, data=binout1[[1]],
      main="Binned Scatterplot") +
  xlab("Games (Count)")+ ylab("Strike Outs (Count)") +
  scale_fill_gradient("Frequencies", low="#56B1F7", high="#132B43") + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,.1,.1,.1),"cm"),
                     legend.text=element_text(size=12) )
@
\center Mapping frequency to shade of tile

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Binned Scatterplot}
<<LogBinnedScatterplotBaseball, echo=FALSE , warning=FALSE, fig.width=4, fig.height=4, out.width='.6\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
binout1 <- RectBin2d(pitchdat$G,pitchdat$SO,.5,-.5,5,25,type="standard")
qplot(binxs, binys, geom="tile", fill=log(binfreq), data=binout1[[1]],
      main="Binned Scatterplot with Log Frequencies") +
  xlab("Games (Count)")+ ylab("Strike Outs (Count)") +
  scale_fill_gradient("Log Frequencies", low="#56B1F7", high="#132B43") + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,.1,.1,.1),"cm"),
                     legend.text=element_text(size=12) )
@
\center Mapping $\log$(frequency) to shade of tile.

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Binning Functions}

Goal for Binned Scatterplot:
  \begin{itemize}
    \item Reduced Binned Data Triples $(x^\ast, y^\ast, c)$
    \item $(x^\ast,y^\ast)$ = bin center
    \item $c$ = bin frequency count
  \end{itemize}

Need binning function to uniquely map points to bins
  \begin{itemize}
    \item $b(.) : (x,y) \rightarrow (x^\ast,y^\ast)$
  \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Rectangular Binning Functions}

Rectangular binning bins each dimension univariately
  \begin{itemize}
    \item $b(x_i,y_i) = (b_X(x_i),b_Y(y_i))$
    \item Forms rectangular binned grid
  \end{itemize}
\pause

\vspace{.25cm}

Generally $b_X(.) : x_i \rightarrow x^\ast_j$ is defined as 
%
\begin{eqnarray}\label{rectbin}
b_X(x_i) = \left\{\begin{array}{ll} 
  x^\ast_{1} &\text{ for all } x_i = \beta_{0} \\
  x^\ast_j & \text{ for all } x_i \in (\beta_{j-1} , \beta_j] 
  \end{array}\right\}
\end{eqnarray}

where $(\beta_{j-1},\beta_{j}]$ for $j \in \{1,\dots,J\}$ are sequence of 
$J$ adjacent intervals spanning range of X data.

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Rectangular Binning Functions}

Standard Rectangular Binning
  \begin{itemize}
    \item Use equally spaced intervals
    \item Bin width $= \omega_X$
    \item $ \{\beta_j \text{ }|\text{ } \beta_j = \beta_{j-1} + \omega_X \}$
    \item $x_j^\ast = (\beta_{j-1} + \beta_j) /2$
  \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Reduced Binned Data}

\tiny
\begin{table}[H]
\begin{minipage}[t]{.33\textwidth}\centering\subfloat[Original Data, 12 rows]{
\begin{tabular}{rr}\hline
       $x$ &  $y$  \\ \hline
-7.7325 & -9.6340 \\ 
  -8.1176 & -1.4529 \\ 
  -5.8996 & -3.2033 \\ 
  -7.0375 & -5.5563 \\ 
  -3.6354 & -3.9315 \\ 
  -8.7639 & 0.9874 \\ 
  -2.9781 & 8.6802 \\ 
  0.8210 & -8.6118 \\ 
  5.4477 & -8.4555 \\ 
  4.6849 & -5.6620 \\ 
  9.4785 & 1.1133 \\ 
  1.7579 & 5.3759 \\  \hline
\end{tabular}}
\end{minipage}\hfil
\begin{minipage}[t]{.33\textwidth}\centering\subfloat[Binned Data Centers, 12 rows]{
\begin{tabularx}{.7\textwidth}{rr} \hline
  $b_X(x)$ & \hspace{.4cm} $b_Y(y)$  \\ \hline
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & -5 \\ 
  -5 & 5 \\ 
  -5 & 5 \\ 
  5 & -5 \\ 
  5 & -5 \\ 
  5 & -5 \\ 
  5 & 5 \\ 
  5 & 5 \\  \hline
\end{tabularx}}
\end{minipage}\hfil
\begin{minipage}[t]{.33\textwidth}\centering\subfloat[Reduced Binned Data, 4 rows]{
\begin{tabularx}{.7\textwidth}{rrr}\hline
 \hspace{.2cm} $x^\ast$   &   \hspace{.2cm} $y^\ast$ & \hspace{.4cm}c \\ \hline
-5 & -5 & 5 \\ 
  -5 & 5 & 2 \\ 
  5 & -5 & 3 \\ 
  5 & 5 & 2 \\  \hline
\end{tabularx}}
\vfil
\end{minipage}
\caption{\label{DataReductionExample}Original, Binned and Reduced Binned Data Tables, with data storage sizes.  Binned using standard rectangular approach with origin $(\beta_{0,x},\beta_{0,y})$ = (-10,-10) and bin widths $\omega_x = \omega_y = 10$. }
\end{table}
\normalsize
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Rectangular Binning Functions}

Random Rectangular Binning
  \begin{itemize}
    \item Define set of bin centers: $\{x_j^\ast \text{ }|\text{ } x_j^\ast > x_{j-1}^\ast \}$
    \item Assign points at random to surrounding bin centers
    \item Probabilities inversely proportional do distance from bin centers
  \end{itemize}

\begin{eqnarray}\label{randbin1}
b^r_X(x_i) = \left\{\begin{array}{ll} 
  x^\ast_j & \text{ with probability } (x^\ast_{j+1} - x_i)/(x^\ast_{j+1} -x^\ast_{j}) \\
  x^\ast_{j+1} &\text{ with probability }  (x_i - x^\ast_{j})/(x^\ast_{j+1} -x^\ast_{j})
  \end{array}\right\}
\end{eqnarray}

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Standard vs. Random Rectangular Binning}
<<StandVsRandDemo, echo=FALSE , warning=FALSE, fig.width=8, fig.height=4, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
n=12
param1 = 1
param2 = 1.6
set.seed(52)
x <- rbeta(n,param1, param2)*20-10
y <- rbeta(n,param1, param2)*20-10
# standard bin centers
xcent <- rep(-5,n) ; xcent[x>0] <- 5
ycent <- rep(-5,n) ; ycent[y>0] <- 5
# standard bin gridlines
temp <- data.frame(   
  x=c( -10,   0,  10, -10, -10, -10), 
  xend=c( -10,   0,  10,  10, 10, 10), 
  y=c( -10, -10 , -10 , -10, 0 , 10), 
  yend=c(10, 10 , 10, -10, 0, 10 ))

p1 <- qplot(xcent, ycent, geom="point", size=I(5), shape=I(3), color=I("gray")) +
      geom_segment(aes(x = x, y = y, xend = xend, yend = yend), data=temp) + 
      geom_segment(aes(x = x, y = y, xend = xcent, yend = ycent)) + 
      geom_point(aes(x=x,y=y), size=I(3))+ 
      theme_bw() + xlab("") + ylab("") +
      ggtitle("Standard Binning") +
      theme(panel.grid.major= element_blank(),panel.grid.minor= element_blank(),
            legend.position="none",plot.margin = unit(c(0,1,0, 1),"cm"),
            aspect.ratio=1)

# random bin centers
set.seed(14)
rxcent <- RandRectBin1d(x, -10, 10)
rycent <- RandRectBin1d(y, -10, 10)

p2 <- qplot(xcent, ycent, geom="point", size=I(5), shape=I(3), color=I("gray")) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), data=temp) + 
  geom_segment(aes(x = x, y = y, xend = rxcent, yend = rycent)) + 
  geom_point(aes(x=x,y=y), size=I(3))+ 
  theme_bw() + xlab("") + ylab("") +
  ggtitle("Random Binning") +
  theme(panel.grid.major= element_blank(),panel.grid.minor= element_blank(),
        legend.position="none",plot.margin = unit(c(0,1,0, 1),"cm"),
            aspect.ratio=1)

grid.arrange(p1,p2,nrow=1)
@

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Spatial Loss}

Create loss function for spatial information
    \begin{itemize}
        \item Use Euclidean distance between points and bin centers
        \item Spatial loss for $i^{th}$ observation: $L_i^S = \sqrt{ \left(x_i-b_X(x_i)\right)^2 + \left(y_i-b_Y(y_i)\right)^2 } $ 
        \item \textit{Total} spatial loss: $L^S = \sum_{i=1}^{n} L_i^S$
    \end{itemize} \pause

\vspace{.25cm}

Slight issue with spatial loss for random binning
    \begin{itemize}
        \item Many possible random assignments result in same reduced binned data
        \item Find assignment with minimum total spatial loss
        \item Call this \textit{net} spatial loss
    \end{itemize}

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Post Processing Random Rectangular Binning}
<<RandNetSpatialLossDemo, echo=FALSE , warning=FALSE, fig.width=7.5, fig.height=5, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
n=12
param1 = 1
param2 = 1.6
set.seed(52)
x <- rbeta(n,param1, param2)*20-10
y <- rbeta(n,param1, param2)*20-10
# standard bin centers
xcent <- rep(-5,n) ; xcent[x>0] <- 5
ycent <- rep(-5,n) ; ycent[y>0] <- 5
# standard bin gridlines
temp <- data.frame(   
  x=c( -10,   0,  10, -10, -10, -10), 
  xend=c( -10,   0,  10,  10, 10, 10), 
  y=c( -10, -10 , -10 , -10, 0 , 10), 
  yend=c(10, 10 , 10, -10, 0, 10 ))

# random bin centers
set.seed(14)
rxcent <- RandRectBin1d(x, -10, 10)
rycent <- RandRectBin1d(y, -10, 10)
# point 1 and 11 are special 
linecolor <- rep("blue", 12)
linecolor[c(1,11)] <- "red"

p2 <- qplot(xcent, ycent, geom="point", size=I(5), shape=I(3), color=I("gray")) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), data=temp) + 
  geom_segment(aes(x = x, y = y, xend = rxcent, yend = rycent, color=linecolor)) + 
  scale_colour_manual(values = c("darkblue","red"))+
  geom_point(aes(x=x,y=y), size=I(3))+ 
  theme_bw() + xlab("") + ylab("") +
  ggtitle("Random Binning") +
  theme(panel.grid.major= element_blank(),panel.grid.minor= element_blank(),
        legend.position="none",plot.margin = unit(c(0,1,0, 1),"cm"),
            aspect.ratio=1)

# post process random bin centers to visually 
# equivalent allocation with minimum spatial loss
rxcentpost <- rxcent
rycentpost <- rycent
rxcentpost[1] <- rxcent[11]
rycentpost[1] <- rycent[11]
rxcentpost[11] <- rxcent[1]
rycentpost[11] <- rycent[1]
# point 1 and 11 are special 
linecolor <- rep("blue", 12)
linecolor[c(1,11)] <- "red"

p3 <- qplot(xcent, ycent, geom="point", size=I(5), shape=I(3), color=I("gray")) +
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), data=temp) + 
  geom_segment(aes(x = x, y = y, xend = rxcentpost, yend = rycentpost, color=linecolor)) + 
  scale_colour_manual(values = c("darkblue","red"))+
  geom_point(aes(x=x,y=y), size=I(3))+ 
  theme_bw() + xlab("") + ylab("") +
  ggtitle("Post-Processed Random Binning") +
  theme(panel.grid.major= element_blank(),panel.grid.minor= element_blank(),
        legend.position="none",plot.margin = unit(c(0,1,0, 1),"cm"),
            aspect.ratio=1)

allbin <- data.frame(type=rep(c("Standard","Random","Post-Processed Random"),each=n),
                     binxs = c(xcent,rxcent,rxcentpost), 
                     binys = c(ycent,rycent,rycentpost))
allbinfreq <- ddply(allbin, .(type,binxs,binys),summarise,
                     freq = length(binxs))

p5 <- qplot(xcent, ycent, geom="point", size=I(5), shape=I(3), color=I("gray")) +
  geom_tile(aes(x=binxs, y=binys, fill=freq),data=subset(allbinfreq,type=="Random")) + 
  scale_fill_gradient(low="#56B1F7", high="#132B43", guide="legend") + 
  theme_bw() + xlab("") + ylab("") +
  theme(panel.grid.major= element_blank(),panel.grid.minor= element_blank(),
        plot.margin = unit(c(0,1,0, 1),"cm"),legend.position="bottom",
            aspect.ratio=1)+
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), data=temp)

p6 <- qplot(xcent, ycent, geom="point", size=I(5), shape=I(3), color=I("gray")) +
  geom_tile(aes(x=binxs, y=binys, fill=freq),data=subset(allbinfreq,type=="Post-Processed Random")) + 
  scale_fill_gradient(low="#56B1F7", high="#132B43", guide="legend") + 
  theme_bw() + xlab("") + ylab("") +
  theme(panel.grid.major= element_blank(),panel.grid.minor= element_blank(),
        plot.margin = unit(c(0,1,0, 1),"cm"),legend.position="bottom",
            aspect.ratio=1)+
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend), data=temp)

grid.arrange(p2,p3,p5,p6,nrow=2)
@

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Rectangular Binning Specifications}

Rectangular Binning Specification
    \begin{itemize}
        \item Binning function
        \item Bin dimensions 
        \item Binning origin
    \end{itemize} \pause

Properties of Interest
    \begin{itemize}
        \item Net spatial loss
        \item Computation time
        \item Visually misleading features 
    \end{itemize}

\end{frame}

%---------------------------------------------------------------

\begin{frame}
    \frametitle{Rectangular Binning Specifications}

Standard versus Random Rectangular Binning
  \begin{itemize}
    \item Standard binning computationally faster
    \item Standard binning is superior for spatial loss
    \item Random binning locally smooths bin frequencies
  \end{itemize}

Small versus Large Bins 
  \begin{itemize}
    \item Net spatial loss decreases as bin dimensions decrease
    \item Computation time increases as bin dimensions decrease
    \item Small bins allow fine bivariate structure to show through
    \item Large bins emphasize large scale bivariate structure
  \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study}

Explore Properties with Three Simulated Data Sets

100000 observations from:

\begin{itemize}
\item Set I: $x_i$ and $y_i \sim \text{iid Exp}(\lambda=11)$

\item Set II: $x_i$ and $y_i \sim \text{iid Normal}(\mu=50,\sigma^2=11^2)$

\item Set III: $x_i$ and $y_i \sim \text{iid Uniform}(a=0,b=100)$
\end{itemize}

Data then rounded to nearest even integer to induce a \textit{data resolution} of 2 units

\end{frame}

%--------------------------------------------------------------

% \begin{frame}
%     \frametitle{Simulated Data}
<<SimulatedDataScatters, echo=FALSE, cache=TRUE, include=FALSE, warning=FALSE, fig.width=7, fig.height=3.5, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
n=100000
set.seed(31415)
unifdat <- data.frame(x= runif(n, min=0, max=100),
                      y= runif(n, min=0, max=100),
                      type="Uniform",
                      record = "Fine")
normdat <- data.frame(x= rnorm(n, mean=50, sd=11),
                     y= rnorm(n, mean=50, sd=11),
                     type="Normal",
                     record = "Fine")
expdat <- data.frame(x= rexp(n, .11),
                     y= rexp(n, .11),
                     type="Exponential",
                     record = "Fine")
datlist <- list(unifdat,normdat,expdat)
### round all data to resolution alpha
alpha <- 2
for(i in 1:3){
  datlist[[i]]$roundx <- ((datlist[[i]]$x + 1)%/%alpha)*alpha
  datlist[[i]]$roundy <- ((datlist[[i]]$y + 1)%/%alpha)*alpha
}

# plot the original data
# make rounded data find overlapping points to make plot rendering faster
dat <- rbind(datlist[[1]],datlist[[2]],datlist[[3]])
smallround <- ddply(dat,.(type,record,roundx,roundy),summarise,
      overlap = length(roundx))
smallround <-cbind(data.frame(x=smallround$roundx,y=smallround$roundy), 
                   smallround[,1:4])
smallround$record <- "Coarse"
alldat <- rbind(dat,smallround)
alldat$type <- factor(alldat$type, levels=c("Exponential","Normal","Uniform"))

qplot(x,y,geom="point",data=subset(alldat, record=="Fine"),facets=.~type, size=I(1)) + 
  theme_bw() + theme(aspect.ratio=1) +
  #ggtitle("Scatterplots of Simulated Data") +
  xlab("X") + ylab("Y") + 
  scale_x_continuous(breaks=c(0,25,50,75,100)) + 
  scale_y_continuous(breaks=c(0,25,50,75,100))
@

% \end{frame}

% %-------------------------------------------------
% 
% \begin{frame}
%     \frametitle{Spatial Loss vs Bin Size}
% <<SpatialLossPlot, echo=FALSE , warning=FALSE, fig.width=7, fig.height=3.5, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
% spatlossdatSquareBins$bintype <- factor(spatlossdatSquareBins$bintype, levels=c("standard","random"))
% qplot(widthx,totalSpatialLoss,geom="path",colour=bintype, linetype=bintype,
%       data=spatlossdatSquareBins,group=bintype, facets=.~distrib, size=I(1)) +
%   theme_bw() + ylab("Net Spatial Loss") +
%   xlab("Size of Square Bins") + ylim(c(0, 1100000))
% @ 
% 
% \end{frame}
% 
% %-------------------------------------------------
% 
% \begin{frame}
%     \frametitle{Binning Time vs Bin Size}
% <<BinningTimePlot, echo=FALSE , warning=FALSE, fig.width=7, fig.height=3.5, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
% qplot(widthx,as.numeric(binTime),geom="path",colour=bintype, linetype=bintype,
%       data=spatlossdatSquareBins,group=bintype, facets=.~distrib, size=I(1)) +
%   theme_bw() + ylab("Binning Time (sec)") +
%   xlab("Size of Square Bins")
% @ 
% 
% \end{frame}
%

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Simulated Data}
<<SimulatedCoarseDataScatters, echo=FALSE, warning=FALSE, fig.width=7, fig.height=3.5, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
qplot(x,y,geom="point",data=subset(alldat, record=="Coarse"),facets=.~type, size=I(1)) + 
  theme_bw() + theme(aspect.ratio=1) +
  #ggtitle("Scatterplots of Simulated Data") +
  xlab("X") + ylab("Y") + 
  scale_x_continuous(breaks=c(0,25,50,75,100)) + 
  scale_y_continuous(breaks=c(0,25,50,75,100))
@

\center Scatterplots of simulated data recorded to nearest even integer

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Selecting Bin Width}

<<Stripes, echo=FALSE, warning=FALSE, fig.width=7, fig.height=3.5, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
setwd("C:\\Users\\Karsten\\Documents\\GitHub\\dbaccess")
allbindat <- read.csv("StripeBinData.csv",header=TRUE)
allbindat$BinningType <- factor(allbindat$BinningType, levels=c("Standard","Random"))

p1 <- qplot(binxs, binys, geom="tile", fill=binfreq, data=subset(allbindat,BinSize=="1 X 1 Unit")) +
  facet_grid(BinningType~BinSize)+
  ggtitle(" ") +   xlab("") + ylab("") +
  scale_fill_gradient(low="#56B1F7", high="#132B43",
                      limits=c(0, 450),name="Bin Count") + 
  theme_bw() + theme(legend.position="none",aspect.ratio=1,
                     panel.grid= element_blank())

p2 <- qplot(binxs, binys, geom="tile", fill=binfreq, data=subset(allbindat,BinSize=="4 X 4 Unit")) +
  facet_grid(BinningType~BinSize)+
  ggtitle(" ") +   xlab("") + ylab("") +
  scale_fill_gradient(low="#56B1F7", high="#132B43",
                      limits=c(0, 450),name="Bin Count") + 
  theme_bw() + theme(legend.position="none",aspect.ratio=1,
                     panel.grid= element_blank())

p3 <- qplot(binxs, binys, geom="tile", fill=binfreq, data=subset(allbindat,BinSize=="5 X 5 Unit")) +
  facet_grid(BinningType~BinSize)+
  ggtitle(" ") +   xlab("") + ylab("") +
  scale_fill_gradient(low="#56B1F7", high="#132B43",
                      limits=c(0, 450),name="Bin Count") + 
  theme_bw() + theme(legend.position="none",aspect.ratio=1,
                     panel.grid= element_blank())

grid.arrange(p1,p2,p3, nrow=1)
@   
%
\center Binned Scatterplots of Uniform Data \\

\center Bin dimensions for standard binning should be integer multiples of the resolution of the data to avoid \textit{artificial} stripes.

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Selecting a Binning Origin}

Specifying a binning origin to lower spatial loss
  \begin{itemize}
    \item Avoid heavy overhangs outside data boundaries
    \item Try to align bin centers with possible data values 
  \end{itemize}

A reliable default based on the data resolution
  \begin{itemize}
    \item Recorded to nearest $\alpha_x$ units in the X dimension 
    \item Recorded to nearest $\alpha_y$ units in the Y dimension
    \item $x_{(1)}$ and $y_{(1)}$ are the minimum data values in each dimension
    \item Set binning origin at $(x_{(1)}, y_{(1)}) - (\alpha_x/2,\alpha_y/2) $
  \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Selecting a Binning Origin}

<<ExponentialOriginShift, echo=FALSE, warning=FALSE, fig.width=7, fig.height=3.5, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
#create binned data for fine exponential with 10X10 bins at good and bad origin
binoutExp5 <- RectBin2d(datlist[[3]]$x,datlist[[3]]$y,0,0,10,10,type="standard")
binoutExp5offset <- RectBin2d(datlist[[3]]$x,datlist[[3]]$y,-9,-9,10,10,type="standard")

p1 <- qplot(binxs, binys, geom="tile", fill=log(binfreq), data=binoutExp5[[1]]) +  
  ggtitle("Origin at (0,0)") +
  scale_fill_gradient(low="#56B1F7", high="#132B43") + 
  theme_bw() + theme(legend.position="none",
                     aspect.ratio=1, panel.grid= element_blank()) +
  xlab("") + ylab("") + geom_hline(yintercept=0) + geom_vline(xintercept=0)
p2 <- qplot(binxs, binys, geom="tile", fill=log(binfreq), data=binoutExp5offset[[1]]) +  
  ggtitle("Origin at (-9,-9)") + 
  scale_fill_gradient(low="#56B1F7", high="#132B43") + 
  theme_bw() + theme(legend.position="none",
                     aspect.ratio=1, panel.grid= element_blank()) +
  xlab("") + ylab("") + geom_hline(yintercept=0) + geom_vline(xintercept=0)

grid.arrange(p1,p2,nrow=1)
@

\center Exponential data binned with 10X10 square bins.\\
Net spatial loss for binning origin at (0,0) is 7\% lower than (-9,-9).

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Selecting a Binning Origin}

<<OriginShiftCoarse, echo=FALSE, warning=FALSE, fig.width=7, fig.height=3.5, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
library(plyr)
setwd("C:\\Users\\Karsten\\Documents\\GitHub\\dbaccess")
lossOrigin <- read.csv("LossesByOriginBinwidth.csv",header=TRUE)
lossOrigin$binsize <- factor(paste(lossOrigin$widthx,"X",lossOrigin$widthx,sep=""),
                             levels=c("2X2","4X4","6X6","8X8","10X10"))

mins <- ddply(lossOrigin, .(widthx, bintype, distrib), summarise,
              minx = max(originx[which(totalSpatialLoss == min(totalSpatialLoss))]),
              minloss = min(totalSpatialLoss))
mins$binsize <- paste(mins$widthx,"X",mins$widthx,sep="")

qplot(originx,totalSpatialLoss/1000,geom="path",group=distrib, colour=distrib,
      data=subset(lossOrigin,bintype=="standard"), size=I(1))+ 
  facet_grid(distrib~binsize, scales="free_x", space="free_x") + theme_bw() + 
  theme(panel.grid= element_blank(), legend.position="none")+
  geom_vline(xintercept = -1, colour=I("darkgray")) +
  geom_point(aes(x=minx, y=minloss/1000), data=subset(mins, bintype=="standard"), size=3.5) +
  ylab("Net Spatial Loss (in thousands of units)") + xlab("Origin Offset") + 
  scale_x_continuous(breaks = seq(-10, 0, by = 1))
@

\center Net spatial loss for simulated data (with 2X2 resolution)  \\
Binning origin at (-1,-1) is good default option for all bin sizes

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Frequency Binning}

Bin count mapped to shade can be rendered continuously in HCL color space

\vspace{.25cm}

Why would we discretize the frequency shading?
  \begin{itemize}
    \item Make inability to accurately visually map bin shades to plot keys explicit
  \end{itemize}

<<Freq4NoBins, echo=FALSE, warning=FALSE, fig.width=8, fig.height=4, out.width='.9\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
binout1 <- RectBin2d(datlist[[2]]$x,datlist[[2]]$y,0,0,5,5,type="standard")
binoutFreqGroups4 <- freqBin(binout1, binType="standard", ncolor=4, logCount=FALSE)
binoutFreqGroups7 <- freqBin(binout1, binType="standard", ncolor=7, logCount=FALSE)
binoutFreqGroups10 <- freqBin(binout1, binType="standard", ncolor=10, logCount=FALSE)

# Plot with 4 Binned Frequencies
labels = unique(binoutFreqGroups4[[1]]$freqlabel)
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:4)/4)
# Plot with 4 Binned Frequencies
p4<- qplot(binxs, binys, geom="tile", fill=factor(freqgroup), data=binoutFreqGroups4[[1]],
           main="4 Standard Freq Bins") +
  xlab("") + ylab("")+
  scale_fill_manual("Frequencies", values=cols, 
                      guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top"),
                      labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.1,.1,.1,.1),"cm"),
                     legend.text=element_text(size=12) )

p0 <- qplot(binxs, binys, geom="tile", fill=binfreq, data=binout1[[1]],
      main="No Freq Binning") +
  xlab("") + ylab("")+
  scale_fill_gradient("Frequency", low="#56B1F7", high="#132B43",
                      guide = guide_colourbar(label.position="bottom", label.hjust=0.5,  
                                              title.position = "top",
                                              barwidth = unit(6, "cm"))) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     panel.grid= element_blank(),
                     plot.margin = unit(c(.1,.1,.1,.1),"cm") )

grid.arrange(p0,p4,nrow=1)
@

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Frequency Binning}

How do we discretize the frequency shading?
  \begin{itemize}
    \item Add second stage of binning
    \item Use one dimensional binning algorithms on bin counts, $c_\ell$
    \item Shade ordinally based on binned frequency groups
    \item May transform counts before frequency binning \\ (i.e. log counts)
  \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Frequency Binning}

Standard Frequency Binning
  \begin{itemize}
    \item Create k frequency groups using equal bin width
  \end{itemize}  
\pause

\vspace{.25cm}

Quantile Frequency Binning
  \begin{itemize}
    \item Create k frequency groups based on empirical quantiles of raw bin counts
    \item Roughly equal number of tiles in each shade
  \end{itemize}  
\pause

\vspace{.25cm}

Shade interpretability analogous to: "Histogram vs Boxplot"

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Frequency Loss}

Frequency Loss
  \begin{itemize}
    \item bin frequencies, $c_\ell$, for spatial bins $\ell \in \{1, \dots, \mathscr{L} \}$\\
    \item binned frequencies, $b_C(c_\ell)$
  \end{itemize}  
%
\begin{eqnarray}\label{totfreqloss}
L^F = \sum_{\ell=1}^{\mathscr{L}} (c_\ell-b_c(c_\ell))^2
\end{eqnarray}  
\pause

Log Frequency Loss
  \begin{itemize}
    \item If bin frequencies, $c_\ell$, are log transformed 
    \item binned log frequencies, $b_C(\log(c_\ell))$
  \end{itemize} 
  %
\begin{eqnarray}\label{logfrequency}
L^{\log F} = \sum_{\ell \in \mathscr{L}^\ast} (\log(c_\ell) - b_c(\log(c_\ell)))^2
\end{eqnarray}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Frequency Binning Specification}

3 Main Choices
  \begin{itemize}
    \item Number of frequency groups (k)
    \item Transform counts (yes/no)
    \item Binning algorithm (standard/quantile)
  \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Number of Frequency Groups}

<<Freq4710NoBins, echo=FALSE, warning=FALSE, fig.width=10, fig.height=4, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
binout1 <- RectBin2d(datlist[[2]]$x,datlist[[2]]$y,0,0,5,5,type="standard")
binoutFreqGroups4 <- freqBin(binout1, binType="standard", ncolor=4, logCount=FALSE)
binoutFreqGroups7 <- freqBin(binout1, binType="standard", ncolor=7, logCount=FALSE)
binoutFreqGroups10 <- freqBin(binout1, binType="standard", ncolor=10, logCount=FALSE)

# Plot with 4 Binned Frequencies
p4<- qplot(binxs, binys, geom="tile", fill=freqgroup, data=binoutFreqGroups4[[1]],
           main="4 Standard Freq Bins") +
  xlab("") + ylab("")+
  scale_fill_gradient("Frequency", low="#56B1F7", high="#132B43")+ 
  theme_bw() + theme(legend.position="none",aspect.ratio=1,
                     plot.margin = unit(c(0.1,.1,.1,.1),"cm") )

# Plot with 7 Binned Frequencies
p7<- qplot(binxs, binys, geom="tile", fill=freqgroup, data=binoutFreqGroups7[[1]],
            main="7 Standard Freq Bins") +
  xlab("") + ylab("")+
  scale_fill_gradient("Frequency", low="#56B1F7", high="#132B43")+ 
  theme_bw() + theme(legend.position="none",aspect.ratio=1,
                     plot.margin = unit(c(0.1,.1,.1,.1),"cm") )

# Plot with 10 Binned Frequencies
p10 <- qplot(binxs, binys, geom="tile", fill=freqgroup, data=binoutFreqGroups10[[1]],
           main="10 Standard Freq Bins") +
  xlab("") + ylab("")+
  scale_fill_gradient("Frequency", low="#56B1F7", high="#132B43")+ 
  theme_bw() + theme(legend.position="none",aspect.ratio=1,
                     plot.margin = unit(c(0.1,.1,.1,.1),"cm") )

p0 <- qplot(binxs, binys, geom="tile", fill=binfreq, data=binout1[[1]],
      main="No Freq Binning") +
  xlab("") + ylab("")+
  scale_fill_gradient("Frequency", low="#56B1F7", high="#132B43",
                      guide = guide_colourbar(label.position="bottom", label.hjust=0.5,  
                                              title.position = "top")) + 
  theme_bw() + theme(legend.position="none",aspect.ratio=1,
                     panel.grid= element_blank(),
                     plot.margin = unit(c(.1,.1,.1,.1),"cm") )

grid.arrange(p4,p7,p10,p0,nrow=1)
@

\center How many different shades can we distinguish simultaneously? \\ 
Research suggests 7 is max for the typical person $\Rightarrow$ use k $\le 7$

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Frequency Loss Simulation}

<<FreqLossLineplot, echo=FALSE, warning=FALSE, fig.width=11, fig.height=4, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
freqLossSimDat <- NULL
distrib <- c("Uniform", "Normal", "Exponential")
for (distidx in 1:3){
  binout1 <- RectBin2d(datlist[[distidx]]$x,datlist[[distidx]]$y,0,0,5,5,type="standard")
  for (ncol in 1:10){
    binoutFreq <- rbind(freqBin(binout1, binType="standard", ncolor=ncol, logCount=FALSE)[[2]],
                        freqBin(binout1, binType="quantile", ncolor=ncol, logCount=FALSE)[[2]], 
                        freqBin(binout1, binType="standard", ncolor=ncol, logCount=TRUE)[[2]], 
                        freqBin(binout1, binType="quantile", ncolor=ncol, logCount=TRUE)[[2]])
    binoutFreq$FreqBinType <- c("Standard", "Quantile", "Standard Log", "Quantile Log")
    binoutFreq$DistType <- distrib[distidx]
    binoutFreq$ncolor <- ncol 
    freqLossSimDat <- rbind(freqLossSimDat,binoutFreq)
  }
}

freqLossSimDat$FreqBinType <- factor(freqLossSimDat$FreqBinType , levels= c("Standard", "Quantile", "Standard Log", "Quantile Log"))

plotdat <- freqLossSimDat[(freqLossSimDat$FreqBinType!="Standard Log" & freqLossSimDat$FreqBinType!="Quantile Log" & freqLossSimDat$DistType=="Exponential"),]
p1 <- qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))

plotdat <- freqLossSimDat[(freqLossSimDat$FreqBinType!="Standard Log" & freqLossSimDat$FreqBinType!="Quantile Log" & freqLossSimDat$DistType=="Normal"),]
p2 <-  qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))

plotdat <- freqLossSimDat[(freqLossSimDat$FreqBinType!="Standard Log" & freqLossSimDat$FreqBinType!="Quantile Log" & freqLossSimDat$DistType=="Uniform"),]
p3 <-qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))

grid.arrange(p1,p2,p3, nrow=1)
@

\center Number of Frequency Bins vs. Frequency Loss

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Log Frequency Loss Simulation}

<<LogFreqLossLineplot, echo=FALSE, warning=FALSE, fig.width=11, fig.height=4, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
plotdat <- freqLossSimDat[((freqLossSimDat$FreqBinType=="Standard Log" |freqLossSimDat$FreqBinType=="Quantile Log")  & freqLossSimDat$DistType=="Exponential"),]
p4 <- qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color=FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Log Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))

plotdat <- freqLossSimDat[((freqLossSimDat$FreqBinType=="Standard Log" |freqLossSimDat$FreqBinType=="Quantile Log") & freqLossSimDat$DistType=="Normal"),]
p5 <- qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,max(plotdat$totalFreqLoss)) +
  xlab("Freq Bins") + ylab("Total Log Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))

plotdat <- freqLossSimDat[((freqLossSimDat$FreqBinType=="Standard Log" |freqLossSimDat$FreqBinType=="Quantile Log") & freqLossSimDat$DistType=="Uniform"),]
p6 <-qplot(ncolor, totalFreqLoss, geom="path",linetype=FreqBinType,
      data=plotdat , color = FreqBinType, size=I(1)) +
  facet_grid(.~DistType, scales="free_y") +  theme_bw() +
  scale_x_continuous(breaks=1:10) + ylim(0,2) +
  xlab("Freq Bins") + ylab("Total Log Freq Loss") + 
  theme(legend.position="bottom",plot.margin = unit(c(.1,.1,.1,.1),"cm"))

grid.arrange(p4,p5,p6, nrow=1)
@

\center Number of Log Frequency Bins vs. Log Frequency Loss

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Dealing With Skewed Bin Counts}

Distribution of bin counts are commonly skewed

<<SkewCounts, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
binout1 <- RectBin2d(datlist[[2]]$x,datlist[[2]]$y,0,0,5,5,type="standard")
p1 <- qplot(binxs, binys, geom="tile", fill=binfreq, data=binout1[[1]],
      main="No Freq Binning") +
  xlab("") + ylab("")+
  scale_fill_gradient("Frequency", low="#56B1F7", high="#132B43",
                      guide = guide_colourbar(label.position="bottom", label.hjust=0.5,  
                                              title.position = "top")) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     panel.grid= element_blank(),
                     plot.margin = unit(c(.1,.1,.1,.1),"cm") )
p2 <- qplot(binfreq, geom="histogram", data=binout1[[1]], main="Histogram of Raw Bin Counts")+
  theme_bw() + xlab("Bin Frequency") + ylab("Number of Bins")
grid.arrange(p1,p2, nrow=1)
@

\center Removes ability to see differences in low density bins

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Dealing With Skewed Bin Counts}

Transforming bin counts is one way to reduce visual impact of high density bins

<<SkewCountFreqBinning, echo=FALSE, warning=FALSE, fig.width=11, fig.height=4, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
binout1 <- RectBin2d(datlist[[2]]$x,datlist[[2]]$y,0,0,5,5,type="standard")
freqgrps = 5
binoutFreqGroups <- freqBin(binout1, binType="standard", ncolor=freqgrps, logCount=FALSE)
binoutQuantFreqGroups <- freqBin(binout1, binType="quantile", ncolor=freqgrps, logCount=FALSE)
binoutLogFreqGroups <- freqBin(binout1, binType="standard", ncolor=freqgrps, logCount=TRUE)
binoutQuantLogFreqGroups <- freqBin(binout1, binType="quantile", ncolor=freqgrps, logCount=TRUE)
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:freqgrps)/freqgrps)
# Plot with Quantile Binned Frequencies
labels = paste(unique(binoutQuantFreqGroups[[1]]$freqlabel), "\n", c("1st Pentile","2nd Pentile","3rd Pentile","4th Pentile","5th Pentile"), sep="")
p3 <- qplot(binxs, binys, geom="tile", fill=factor(freqgroup), data=binoutQuantFreqGroups[[1]],
            main="Quantile Freq Binning") +
    xlab("") + ylab("")+
  scale_fill_manual("Frequencies", values=cols, 
                      guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top"),
                      labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,.1,.1,.1),"cm"),
                     legend.text=element_text(size=12) ) 
# Plot with standard binned Log Frequencies
labels = unique(binoutLogFreqGroups[[1]]$freqlabel)
p4 <- qplot(binxs, binys, geom="tile", fill=factor(freqgroup), data=binoutLogFreqGroups[[1]],
            main="Standard Log Freq Binning") +
    xlab("") + ylab("")+
  scale_fill_manual("Log Frequencies", values=cols, 
                      guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                           title.position = "top"),
                      breaks=unique(binoutLogFreqGroups[[1]]$freqgroup),
                      labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,.1,.1,.1),"cm") ,
                     legend.text=element_text(size=12)) 
grid.arrange(p4,p3, nrow=1)
@

\center Quantile frequency binning adds a second option for overcoming skew in visualization

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Practical Takeaways for Binned Scatterplots}

Spatial Binning
  \begin{itemize}
      \item Standard binning algorithm superior to random binning in spatial loss and computation speed
      \item Smaller bins superior for precisely displaying spatial information     
      \item Make bin dimensions integer multiples of data resolution
      \item Quality default for origin offset based on data resolution
  \end{itemize}  
\pause

\vspace{.25cm}

Frequency Binning
  \begin{itemize}
    \item Use 4 to 7 frequency bins
    \item Use standard log freq or quantile freq binning for desired interpretation
  \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Application to Baseball Data}

<<BaseballRecommended1, echo=FALSE, warning=FALSE,fig.width=4, fig.height=4, out.width='.5\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
# Standard binning ~ 50 in each direction
# so range(x)/50 = 100/50 = 2 , range(y)/50 = 500/50 = 10 
# origin at (min(pitchdat$G)-0.5, min(pitchdat$SO)-0.5) = (0.5, -0.5)
binout1 <- RectBin2d(pitchdat$G,pitchdat$SO,.5,-.5,2,10,type="standard")

qplot(binxs, binys, geom="tile", fill=binfreq, data=binout1[[1]],
      main="Standard Freq Binning") +
  xlab("Games (count)") + ylab("Strike Outs (count)")+
  scale_fill_gradient("Frequencies", low="#56B1F7", high="#132B43") + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,-.5,.1,-.5),"cm"),
                     legend.text=element_text(size=12) )
@

Use Standard Rectangular Binning 
  \begin{itemize}
    \item Want approx 50 bins in each dimension $\Rightarrow$ use 2X10 bins
    \item Binning origin at (0.5,-0.5)  $\Rightarrow$ net spatial loss is 3\% lower than for origin at (1,0)
  \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Application to Baseball Data}

<<BaseballRecommended2, echo=FALSE, warning=FALSE,fig.width=9, fig.height=4, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
binout1 <- RectBin2d(pitchdat$G,pitchdat$SO,.5,-.5,2,10,type="standard")

## quantile freq binning
binoutQuantFreqGroups <- freqBin(binout1, binType="quantile", ncolor=4, logCount=FALSE)
labels = paste(sort(unique(binoutQuantFreqGroups[[1]]$freqlabel)), "\n",c("1st Quartile","2nd Quartile","3rd Quartile","4th Quartile") , sep="")
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:5)/5)
p2 <- qplot(binxs, binys, geom="tile", fill=factor(freqlabel), data=binoutQuantFreqGroups[[1]],
            main="Quantile Freq Binning") +
  xlab("Games (count)") + ylab("Strike Outs (count)")+
  scale_fill_manual("Frequencies", values=cols, 
                    guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                         title.position = "top"),
                    labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,-1,-.1,-1),"cm"),
                     legend.text=element_text(size=12) )

## standard log freq binning
binoutLogFreqGroups <- freqBin(binout1, binType="standard", ncolor=4, logCount=TRUE)
binoutLogFreqGroups[[1]]$freqlabel <- factor(binoutLogFreqGroups[[1]]$freqlabel, 
                                             levels = unique(binoutLogFreqGroups[[1]]$freqlabel)[c(3,5,2,4,1)])
labels = levels(binoutLogFreqGroups[[1]]$freqlabel)
cols <- seq_gradient_pal(low="#56B1F7", high="#132B43", space="Lab")((0:5)/5)

p3 <- qplot(binxs, binys, geom="tile", fill=factor(freqlabel), data=binoutLogFreqGroups[[1]],
      main="Standard Log Freq Binning") +
  xlab("Games (count)") + ylab("Strike Outs (count)")+
  scale_fill_manual("Log Frequencies", values=cols, 
                    guide = guide_legend(label.position="bottom", label.hjust=0.5,  
                                         title.position = "top"),
                    labels=labels) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     legend.key.width = unit(1, "cm"),
                     panel.grid= element_blank(),
                     plot.margin = unit(c(0.5,-1,.1,-1),"cm"),
                     legend.text=element_text(size=12) )

# plot three binned scatterplots (raw, standard log freq, quantile freq)
grid.arrange(p2,p3, nrow=1)
@

\center After frequency binning with 4 shade groups

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{}

\begin{center}
\large Shiny Database Sampler \\
\normalsize Chapter 3
\end{center}

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Software Development}    
  Goals
    \begin{itemize}
      \item Connect intro students with big data
      \item Bypass technical hurdles of big data
    \end{itemize} \pause
  Technical Components
    \begin{itemize}
      \item \textit{Data storage}: MySQL
      \item \textit{Computation}: \texttt{R}, \texttt{shiny}, \texttt{RMySQL}
      \item \textit{User interface}: accessed through web browser
    \end{itemize} \pause
  Interface Design
    \begin{itemize}
      \item Quality characteristics of software design
      \item Cognitive load theory
      \item Human centered design
      \item Pedagogical value and curricular integration
    \end{itemize}
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{}

\begin{center}  
  \includegraphics[width=.9\textwidth]{Figure/SampleAndSummarizeWithPanelLabels.png}
\end{center}
\center \href{http://shiny.stat.iastate.edu/karstenm/ShinyDatabaseSampler/}{User Interface}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Applications}    
  Lab Assignment
    \begin{itemize}
      \item Groups assignment
      \item Questions on hypothetical sampling goals
      \item Use Shiny Database Samplers to gather and explore samples
    \end{itemize}
  Course Projects
    \begin{itemize}
      \item Group project
      \item Research question with focus on bivariate associations
      \item Shiny Database Sampler use by some for mock surveys
    \end{itemize}
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Survey}    
  Student User Survey
    \begin{itemize}
      \item 265 students from stat 104
      \item Survey followed lab assignment
      \item Likert scaled responses to 3 sets of 4 items
      \item Item Sets designed to focus on 3 attributes of interest
      \begin{itemize}
        \item Ease of use
        \item Connection to sampling concepts
        \item Engagement with data
      \end{itemize}
      \item Two per item set written in negative, then reverse scored
    \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Survey}

\begin{table}[hbtp]
\centering
\scriptsize
\begin{tabular}{cp{2.75in}c}
Topic Set & Item & Polarity\\[3pt]
  \hline
Ease & \it I found the web tool easy to use & $+$\\  
 & \it The layout of the web tool was intuitive & $+$  \\  
 &  \it Using the web tool was difficult & $-$ \\  
 &  \it Learning to use the web tool was hard & $-$  \\  [3pt]
  \hline
Concept &  \it The web tool helped me understand sampling concepts & $+$  \\   
&  \it I understand sampling ideas less after using the web tool & $-$ \\  
&  \it Sampling techniques are clearer after using the web tool & $+$  \\  
&  \it The web tool made me less sure how to randomly sample & $-$ \\  
   \hline
Engagement &  \it I did not enjoy working with the Census data & $-$  \\  
& \it I thought the Census data was boring & $-$  \\  
& \it Knowing that the Census data was from real people made it more interesting & $+$  \\   
& \it I liked analyzing the Census data & $+$ \\   
\end{tabular}
\caption{Survey Items}
\label{tab:surveyquestions}
\end{table}

\normalsize

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Survey}
<<fluctuationDiagrams,echo=F, message=FALSE, include=T,eval=T,fig.width=12, fig.height=6, out.width='.98\\linewidth', fig.pos='hbt',fig.align='center',tidy=F, cache=TRUE, fig.show='hold'>>=
#ease of use
e12 <- ggfluctuation(table(cdat$q1,cdat$q4)) +
  xlab("Ease Item 1") + ylab("Ease Item 2") + theme_fluct
e13 <- ggfluctuation(table(cdat$q1,cdat$q7)) + xlab("Ease Item 1") + 
  ylab("Ease Item 3") + theme_fluct
e14 <- ggfluctuation(table(cdat$q1,cdat$q10)) + xlab("Ease Item 1") + 
  ylab("Ease Item 4") + theme_fluct
e23 <- ggfluctuation(table(cdat$q4,cdat$q7)) + xlab("Ease Item 2") + 
  ylab("Ease Item 3") + theme_fluct
e24 <- ggfluctuation(table(cdat$q4,cdat$q10)) + xlab("Ease Item 2") + 
  ylab("Ease Item 4") + theme_fluct
e34 <- ggfluctuation(table(cdat$q7,cdat$q10)) + xlab("Ease Item 3") + 
  ylab("Ease Item 4") + theme_fluct

#concept
c12 <- ggfluctuation(table(cdat$q2,cdat$q5)) +
  xlab("Concept Item 1") + ylab("Concept Item 2") + theme_fluct
c13 <- ggfluctuation(table(cdat$q2,cdat$q8)) +
  xlab("Concept Item 1") + ylab("Concept Item 3") + theme_fluct
c14 <- ggfluctuation(table(cdat$q2,cdat$q11)) +
  xlab("Concept Item 1") + ylab("Concept Item 4") + theme_fluct
c23 <- ggfluctuation(table(cdat$q5,cdat$q8)) +
  xlab("Concept Item 2") + ylab("Concept Item 3") + theme_fluct
c24 <- ggfluctuation(table(cdat$q5,cdat$q11)) +
  xlab("Concept Item 2") + ylab("Concept Item 4") + theme_fluct
c34 <- ggfluctuation(table(cdat$q8,cdat$q11)) +
  xlab("Concept Item 3") + ylab("Concept Item 4") + theme_fluct

#Engagement with cdata
d12 <- ggfluctuation(table(cdat$q3,cdat$q6)) +
  xlab("Engagement Item 1") + ylab("Engagement Item 2") + theme_fluct
d13 <- ggfluctuation(table(cdat$q3,cdat$q9)) +
  xlab("Engagement Item 1") + ylab("Engagement Item 3") + theme_fluct
d14 <- ggfluctuation(table(cdat$q3,cdat$q12)) +
  xlab("Engagement Item 1") + ylab("Engagement Item 4") + theme_fluct
d23 <- ggfluctuation(table(cdat$q6,cdat$q9)) +
  xlab("Engagement Item 2") + ylab("Engagement Item 3") + theme_fluct
d24 <- ggfluctuation(table(cdat$q6,cdat$q12)) +
  xlab("Engagement Item 2") + ylab("Engagement Item 4") + theme_fluct
d34 <- ggfluctuation(table(cdat$q9,cdat$q12)) +
  xlab("Engagement Item 3") + ylab("Engagement Item 4") + theme_fluct

grid.arrange(e12,e13,e14,e23,e24,e34,
             c12,c13,c14,c23,c24,c34,
             d12,d13,d14,d23,d24,d34,
             nrow=3) 
@

\center Fluctuation Diagrams of All Item Pairs within Topic Sets

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Survey}

\begin{table}[hbtp]
\centering
\begin{tabular}{lrc}
  \hline
Set & Estimate & 95\% Confidence Interval \\ 
  \hline
Ease & 0.70 & (0.613 , 0.759)\\ 
  Concept & 0.53 & (0.410 , 0.637) \\ 
  Engagement & 0.72 & (0.643 , 0.776) \\ 
   \hline
\end{tabular}
\caption{Cronbach's $\alpha$ Estimates for each item set with 95\% central bootstrap confidence interval based on 10,000 bootstrap samples} 
\label{cronbachstuff}
\end{table}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Survey}
<<PCA6plot,echo=F, message=FALSE, include=T,eval=T,fig.width=8, fig.height=4, out.width='.9\\linewidth', fig.pos='hbt',fig.align='center',tidy=F, cache=TRUE, fig.show='hold'>>=
#average score per student per set
cdat$student <- 1:nrow(cdat)

# Try PCA with 4 values Ease, Engagement, Concept+ and Concept-
Set6Scores <- ddply(cdat,.(student),summarize,
                  EasePos = mean(c(q1,q4), na.rm=TRUE),
                  EaseNeg = mean(c(q7,q10), na.rm=TRUE),
                  ConceptsPos = mean(c(q2,q8), na.rm=TRUE),
                  ConceptsNeg = mean(c(q5,q11), na.rm=TRUE),
                  EngagedPos = mean(c(q9,q12), na.rm=TRUE),
                  EngagedNeg = mean(c(q3,q6), na.rm=TRUE)  )
pc6 <- princomp(Set6Scores[,2:7],na.rm=T)

pc6loadings <- loadings(pc6)[1:6,]
pc6vars <- data.frame( PrinComp = 1:6,
            PropOfVar = (pc6[[1]]^2)/sum(pc6[[1]]^2),
            CumPropOfVar = cumsum((pc6[[1]]^2)/sum(pc6[[1]]^2)) ,
            row.names=NULL)
colnames(pc6vars) <- c("PC" , "Prop. of Var", "Cumulative Prop. of Var")
load23dat <- data.frame(pc6loadings[,2:3], compname = row.names(pc6loadings))
require(grid)
ggplot(data=load23dat)+ 
  geom_hline(x=0, colour="darkgray") + geom_vline(y=0, colour="darkgray")+
  geom_point(aes(x=Comp.2, y=Comp.3),size=3.5) + xlim(c(-.73,.6)) + 
  theme_bw() + xlab("Component 2") + ylab("Component 3") +
  geom_segment(aes(x = 0, y = 0, xend = Comp.2, yend = Comp.3), colour="grey40") + 
  geom_text(aes(x=0.825*Comp.2, y=0.825*Comp.3, label=compname)) +coord_fixed() + 
  theme(plot.margin=unit(c(0,0,0,0), unit="cm")) + coord_fixed(ratio = 1)
@
\center Item Pair Loadings on Components 2 and 3 from the Principal Component Analysis with Six Topic/Polarity Item Pairs
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Survey}
<<OverallSetBars,echo=F, message=FALSE, include=T ,eval=T,fig.width=9, fig.height=2.5, out.width='.99\\linewidth', fig.pos='hbt',fig.align='center',tidy=F, cache=TRUE, fig.show='hold'>>=
meltdatNoNA$value <- as.numeric(as.character(meltdatNoNA$value))
qu.summary <- ddply(meltdatNoNA, .(set, question.in.set, value), summarize,
n=length(value) )
qu.summary$polarity <- rep(c("Pos","Pos","Neg","Neg","Pos","Neg","Pos","Neg","Neg","Neg","Pos","Pos"),each=5)
qu.summary$glyphShift <- .05
qu.summary$glyphShift[qu.summary$polarity=="Neg"] <- -.05

qu.means <- ddply(qu.summary, .(set, value, polarity), summarize, means=mean(n))

# #write.csv(qu.means, "quMeans.csv",row.names=FALSE)
# qu.means <- read.csv("quMeans.csv",header=T)

ggplot() +
  geom_bar(aes(x=as.factor(value-3), weight=means,fill=polarity),data=qu.means,width=.5, position="dodge", size=I(2)) + 
  theme_bw() + scale_colour_brewer("", palette="Set1") + 
  scale_shape_manual("", values=c("-","+")) +
#  scale_shape_manual("", values=15:18) +
#  geom_point(aes(x=as.character(value-3), y=n,  colour=question.in.set, shape=polarity), data=qu.summary, size=I(7),position="dodge") +
  facet_grid(facets=.~set) + 
  xlab("Responses (Post Reverse-Scoring All Negative Item)") + ylab("Count") 
@
\center Item Set Response Distributions by Polarity

\end{frame}

% %-------------------------------------------------
% 
% \begin{frame}
%     \frametitle{Survey}
%     
%   Four sets for final analysis
%     \begin{itemize}
%       \item Ease and Engagement item sets
%       \item Two polarity-based sets for Concept items
%     \end{itemize}
%   
%   \pause
% 
%   Are these sets measuring unique latent characteristics?
% 
% <<PCA4,echo=F, message=FALSE, include=F,eval=T,fig.width=7.5, fig.height=5, out.width='.49\\linewidth', fig.pos='hbt',fig.align='center',tidy=F, cache=TRUE, fig.show='hold'>>=
% #average score per student per set
% cdat$student <- 1:nrow(cdat)
% 
% # Try PCA with 4 values Ease, Engagement, Concept+ and Concept-
% Set4Scores <- ddply(cdat,.(student),summarize,
%                   Ease = mean(c(q1,q4,q7,q10), na.rm=TRUE),
%                   ConceptsPos = mean(c(q2,q8), na.rm=TRUE),
%                   ConceptsNeg = mean(c(q5,q11), na.rm=TRUE),
%                   Engaged = mean(c(q3,q6,q9,q12), na.rm=TRUE)  )
% pc4 <- princomp(Set4Scores[,2:5],na.rm=T)
% 
% summary(pc4)
% loadings(pc4)
% 
% pc4loadings <- loadings(pc4)[1:4,]
% pc4vars <- data.frame( PrinComp = 1:4,
%             PropOfVar = (pc4[[1]]^2)/sum(pc4[[1]]^2),
%             CumPropOfVar = cumsum((pc4[[1]]^2)/sum(pc4[[1]]^2)) ,
%             row.names=NULL)
% colnames(pc4vars) <- c("PC" , "Prop. of Var", "Cumulative Prop. of Var")
% @
% 
% \begin{table}[hbtp]
% \centering
% \footnotesize
% \begin{tabular}{lrrrrr}
% \hline 
% & Principal Component & 1 & 2 & 3 & 4 \\ 
% \hline
% Variances \hspace{.1in} & Prop. of Var & \Sexpr{sprintf("%.3f", pc4vars[1,2])} & \Sexpr{sprintf("%.3f", pc4vars[2,2])}  & \Sexpr{sprintf("%.3f", pc4vars[3,2])} & \Sexpr{sprintf("%.3f", pc4vars[4,2])}  \\
% & Cumu. Prop. of Var & \Sexpr{sprintf("%.3f", pc4vars[1,3])} & \Sexpr{sprintf("%.3f", pc4vars[2,3])}  & \Sexpr{sprintf("%.3f", pc4vars[3,3])} & \Sexpr{sprintf("%.3f", pc4vars[4,3])}  \\
% \hline
% Loadings & Ease & \Sexpr{sprintf("%.3f", pc4loadings[1,1])} & \Sexpr{sprintf("%.3f", pc4loadings[1,2])} & \Sexpr{sprintf("%.3f", pc4loadings[1,3])} & \Sexpr{sprintf("%.3f", pc4loadings[1,4])}\\
% & Pos. Concept &  \Sexpr{sprintf("%.3f", pc4loadings[2,1])} & \Sexpr{sprintf("%.3f", pc4loadings[2,2])} & \Sexpr{sprintf("%.3f", pc4loadings[2,3])} & \Sexpr{sprintf("%.3f", pc4loadings[2,4])}\\
% & Neg. Concept &  \Sexpr{sprintf("%.3f", pc4loadings[3,1])} & \Sexpr{sprintf("%.3f", pc4loadings[3,2])} & \Sexpr{sprintf("%.3f", pc4loadings[3,3])} & \Sexpr{sprintf("%.3f", pc4loadings[3,4])}\\
% & Engagement &  \Sexpr{sprintf("%.3f", pc4loadings[4,1])} & \Sexpr{sprintf("%.3f", pc4loadings[4,2])} & \Sexpr{sprintf("%.3f", pc4loadings[4,3])} & \Sexpr{sprintf("%.3f", pc4loadings[4,4])}\\
% \hline
% \end{tabular}
% \caption{Principal Component Analysis with Final Four Item Sets} 
% \label{pctable}
% \end{table}
% 
% \normalsize
% 
% \end{frame}
%-------------------------------------------------

\begin{frame}
    \frametitle{Results} 
      Ease of use
        \begin{itemize}
          \item Students found the interface easy to use
        \end{itemize}
      Connection to sampling concepts
        \begin{itemize}
          \item Student views were neutral toward the learning benefit
          \item However they did not believe it hurt their understanding
        \end{itemize}
      Engagement with data
        \begin{itemize}
          \item Students were moderately engaged with the census data
        \end{itemize}
\end{frame}

%--------------------------------------------------------

\begin{frame}
    \frametitle{Thanks!}

\center Thank you for attending! \\

\center Any questions?

\end{frame}

\end{document}