### 2015-05-24 DefenseSlides/JobsTalk/JobsTalk.Rnw

\documentclass{beamer} 
% \usepackage{graphicx}
\usepackage{graphics}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{bbm}

\addtobeamertemplate{navigation symbols}{}{%
    \usebeamerfont{footline}%
    \usebeamercolor[fg]{footline}%
    \hspace{1em}%
    \insertframenumber/\inserttotalframenumber
}

\setbeamercovered{transparent}
\renewcommand{\ni}{\noindent}

\newcommand{\distas}[1]{\mathbin{\overset{#1}{\sim}}}%
\newcommand{\Cov}{\text{Cov}}%

<<setup,echo=F,include=F,eval=T>>=
### Preliminaries 
## Load Packages
library(car)
library(ggplot2)
library(subselect)
library(qtlmt)
library(multcomp)
library(lattice)
library(gridExtra)
library(reshape2)
library(plyr)
library(boot)

## Load Data
setwd("C:\\Users\\Karsten\\Dropbox\\Dissertation\\CurriculumStudy\\DraftTISE")
dat <- read.csv("./Data/DataForStudyAnalysis.csv",header=T)
#change factor level order to put sweeney as default control group
dat$room <- factor(dat$room.x, levels=c("sweeney", "kildee"))
dat$lab5perc <- dat$lab5perc * 100
dat$hw2perc <- dat$hw2perc * 100
dat$treatment <- "Simulation-Based"
dat$treatment[which(dat$room=="sweeney")] <- "Traditional"
@

<<ARTISTModSelectDiag,echo=F,include=F,eval=T>>=
### Build MANCOVA model for ARTIST scaled multiple choice scores for CI and HT
## Backward Stepwise Selection Using AIC as selection criterion
#start with biggest possible model
mod2 <- lm(cbind(ConfMC,HypMC)~ midterm + Section +
             hw1perc +  hw2perc + hw3perc + hw4perc + hw5perc + hw6perc + hw7perc +
            lab1perc +lab2perc +lab3perc +lab4perc +lab5perc +lab6perc + lab7perc +  room , data=dat)
summary(manova(mod2))
#backward stepwise to reduce model
mod2backward <- mStep(mod2, k=2, trace=TRUE) #k=2 means to use AIC
summary(manova(mod2backward))
#removal of lab2, hw4 due to p-values over .1 
#also remove hw5 due to worries about colinearity with lab5 (same topics)
mod2small <- update(mod2backward, .~. -  lab2perc - hw4perc - hw5perc)
summary(manova(mod2small), tests=c("Pillai","Wilks"))
summarytab <- summary(manova(mod2small))$stats
#significant overall effects: midterm and lab5 (random sampling lab)
#weak overall effect of treatment
summary(mod2small)
mod2smallCI <- lm(ConfMC~ midterm + lab5perc + room, data=dat)
summary(mod2smallCI)
CIcis <- data.frame(confint(mod2smallCI,level=.95))
CIcis$ests <- mod2smallCI$coeff
mod2smallHT <- lm(HypMC~ midterm + lab5perc + room, data=dat)
summary(mod2smallHT)
HTcis <- data.frame(confint(mod2smallHT,level=.95))
HTcis$ests <- mod2smallHT$coeff
#individually there is a significant effect of treatment on CI score
  #multiple comparisons adjustment ?!?!

#individually there is no significant effect of treatment on HT score

#--------------------------------------------------
### Check Conditions for MANCOVA linear model fit
## check response correlations to avoid colinearity in multivariate model
#Rule of Thumb: ~ .3 to .55 then MANCOVA will work well
with(dat, cor(ConfMC,HypMC)) #Not overly correlated

## Normality of responses
#qplot(dat$ConfMC)
#qplot(dat$HypMC)
#each is reasonably normal univariately
#qplot(dat$ConfMC,dat$HypMC) + geom_jitter()
# reasonable bivariate normal

## check residual plots for homoscedesticity and linearity
qplot(mod2small$fitted.values[,1] ,mod2small$residuals[,1]) + 
  geom_hline(yintercept=0) + geom_jitter()
qplot(mod2small$fitted.values[,2] ,mod2small$residuals[,2]) + 
  geom_hline(yintercept=0) + geom_jitter()
# these seem fine (put into appendix)

## Independence of responses between students must be assumed
@

<<results, echo=FALSE, include=FALSE>>=
load("C:\\Users\\Karsten\\Dropbox\\Dissertation\\CurriculumStudy/SimulationStudy/SimulationResults.RData")
library(plyr)
eta <- (0:5)/10
results <- ldply(1:6, function(x) data.frame(eta=eta[x], resultlist[[x]]))
# beta1s and beta2s are actually tau1 and tau2
simresults = ddply(results, .(eta), summarise,
      trt1=mean(beta1s),
      trt2=mean(beta2s),
      type1.1=mean(pval1s<0.05),
      type1.2=mean(pval2s<0.05),      
      betamidterm1s = mean(betamidterm1s),
      pvalmidterms1s = mean(pvalmidterm1s),
      sebetamidterm1s = mean(SEbetamidterm1s),
      betamidterm2s = mean(betamidterm2s),
      pvalmidterms2s = mean(pvalmidterm2s),
      sebetamidterm2s = mean(SEbetamidterm2s),
      avgpillai = mean(pillais),
      avgpillaipval = mean(pillaipvals),
      sebeta1s = mean(SEbeta1s),
      sebeta2s = mean(SEbeta1s),
      sdbeta1 = sd(beta1s),
      sdbeta2 = sd(beta2s)
)

L <- length(resultlist)
M = nrow(resultlist[[1]]) 
errorsummary <- data.frame(T2ErrorRateBetaMidterm1 = rep(NA, L),
                           T2ErrorRateBetaMidterm2 = rep(NA, L),
                           T2ErrorRateBetaLab1 = rep(NA, L),
                           T2ErrorRateBetaLab2 = rep(NA, L),
                           T1ErrorRateBeta1 = rep(NA, L),
                           T1ErrorRateBeta2 = rep(NA, L),
                           T1ErrorRatePillai = rep(NA, L))

for (l in 1:L){
  errorsummary[l,1] <- 1- length(which(resultlist[[l]]$pvalmidterm1s <= 0.05)) / M
  errorsummary[l,2] <- 1- length(which(resultlist[[l]]$pvalmidterm2s <= 0.05)) / M
  errorsummary[l,3] <- 1- length(which(resultlist[[l]]$pvallab1s <= 0.05)) / M
  errorsummary[l,4] <- 1- length(which(resultlist[[l]]$pvallab2s <= 0.05)) / M
  errorsummary[l,5] <- length(which(resultlist[[l]]$pval1s <= 0.05)) / M
  errorsummary[l,6] <- length(which(resultlist[[l]]$pval2s <= 0.05)) / M
  errorsummary[l,7] <- length(which(resultlist[[l]]$pillaipvals <= 0.05)) / M
}
errorsummary$eta <- eta
#errorsummary
@

\title[Comparison of Learning Outcomes for Simulation-Based and Traditional Curricula for Statistical Inference in a Designed Educational Experiment]{Comparison of Learning Outcomes for Simulation-Based and Traditional Curricula for Statistical Inference in a Designed Educational Experiment}
\author[K. Maurer]{Karsten Maurer}
\date{February 2, 2015}
\institute[ISU]{Iowa State University}
\begin{document}

\begin{frame}
    \maketitle
\end{frame}

\begin{frame}
    \frametitle{Educational Background}

    \begin{itemize}
        \item BA in Statistics and History from University of Minnesota, Morris (2009)
        \item MS in Statistics from Iowa State University (2011)
        \item PhD in Statistics In Progress at Iowa State University (Planned Summer 2015)
    \end{itemize}

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Research Interests}
Statistics Education
        \begin{itemize}
          \item Technology for Stat Ed
          \item Curriculum Development
        \end{itemize}

Statistical Graphics
        \begin{itemize}
          \item Visualization of Large Data Sources
          \item Graphical Perception
        \end{itemize}

\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Presentation Outline}  
    \begin{itemize}
      \item Statistical-Inference Curricula
      \item Experimental Design
      \item Data Description
      \item Model Based Analysis
        \begin{itemize}
          \item Model Selection
          \item Results
          \item Model Assessment
        \end{itemize}
      \item Discussion and Conclusions
    \end{itemize}
\end{frame}

%--------------------------------------------------------------

\begin{frame}
    \frametitle{Curriculum Study Introduction}

Study Goal: Compare student learning outcomes for under two introductory curricula for statistical inference in a designed educational experiment\\

\vspace{.1in}

Two Curricula for Statistical Inference
  \begin{itemize}
    \item Traditional statistical inference curricula
    \item Simulation-based statistical inference curricula
  \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Statistical Inference Curricula}
   Traditional Curriculum for Statistical Inference 
    \begin{itemize}
        \item Result of Stat Ed reform movement of 1990's
        \item Summarized in GAISE College Report (2005)
        \item Statistical inference through theory-based approach
        \begin{itemize}
        \item Normal model, CLT, formulas and tables
        \end{itemize}
        \item Textbooks Examples
          \begin{itemize}
            \item Introduction to the Practice Statistics (Moore, 2005)
            \item Statistics: The Art and Science of Learning From Data (Agresti \& Franklin, 2012)
          \end{itemize}
    \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Statistical Inference Curricula}
   Simulation-Based Curriculum for Statistical Inference
    \begin{itemize}
        \item Builds from Traditional curriculum for statistical inference
        \item Simulation/randomization to teach core concepts of inference
        \item Statistical inference introduced through simulation-based methods
          \begin{itemize}
          \item Bootstrap confidence intervals
          \item Simulation/randomization-based hypothesis tests
          \end{itemize}
          \item Textbooks Examples
            \begin{itemize}
              \item Statistics: Unlocking the Power of Data (Lock, Lock, Morgan, Lock \& Lock, 2012)
              \item Introduction to Statistical Investigations (Tintle, Chance, Cobb, Rossman, Roy, Swanson \& VanderStoep, 2014)
            \end{itemize}
        \item May elect to additionally teach theory-based methods  
    \end{itemize}
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation-Based Statistical Inference Classroom Example}
   Can Dolphins Communicate? 
    \begin{itemize}
        \item Example from workshop section led by Alan Rossman
        \item Well known study from 1960s investigated if dolphins could communicate abstract ideas
        \item 2 trained dolphins: Buzz and Doris
        \item Trained:
        \begin{itemize}
          \item Press button 1 when light blinking
          \item Press button 2 when light steady.
          \item Correct $\rightarrow$ get fish!
        \end{itemize}
    \end{itemize}
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation-Based Statistical Inference Classroom Example}

\includegraphics[width=1\textwidth]{Dolphins.png}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation-Based Statistical Inference Classroom Example}

    \begin{itemize}
        \item Buzz pressed right button 15/16 times.
        \item Test H$_o$:  Buzz guessing
        \item Simulate under H$_o$ to make randomization distribution
        \item Using coins first then \texttt{\href{http://lock5stat.com/statkey/}{Statkey}} software
        \item Compare data to randomization distribution
    \end{itemize}
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Experimental Design: Subjects}
   Conducted within Stat 104: Introduction to Statistics
      \begin{itemize}
        \item Agricultural and biological science students
        \item (2 hour lecture + 2 hour lab) / week
        \item 16 week course
      \end{itemize} \pause
   116 Students enrolled in two sections of Stat 104
      \begin{itemize}
        \item 4 Students dropped by week 2 of course
        \item 11 Students did not consent to data release
        \item 101 Students consented to data release
      \end{itemize}    
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Experimental Design: Random Assignment}
   Each student randomly assigned to inference curriculum

   \vspace{.1in}

   Resulted in four cohorts
      \begin{itemize}
        \item Cohort A: Section 1 / Simulation-Based
        \item Cohort B: Section 2 / Simulation-Based
        \item Cohort C: Section 1 / Traditional
        \item Cohort D: Section 2 / Traditional
      \end{itemize}

    101 students consented to release of course data for study
    \begin{itemize}
        \item 50 students from Traditional Curriculum
        \item 51 students from Simulation-Based Curriculum
    \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Experimental Design: Treatments}

  \includegraphics[width=1\textwidth]{CurriculaPaths2.pdf}
\center Curricula Schedules

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Experimental Design: Treatment Administration}

\includegraphics[keepaspectratio=true, width=1\textwidth]{LectureInstructionWithRoomsAndKey.pdf}

\center Lecture: Instructor and Room Schedule

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Experimental Design: Treatment Administration}

\includegraphics[keepaspectratio=true, width=1\textwidth]{LabInstructionWithRooms.pdf}

\center Lab: Instructor and Room Schedule

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Data Description}
Course Administration
    \begin{itemize}
        \item Enrollment Section
        \item Statistical Inference Curriculum
    \end{itemize} \pause
Pre-Treatment Measures
    \begin{itemize}
        \item Homework 1-7 Scores
        \item Lab 1-7 Scores 
        \item Midterm Exam Score
    \end{itemize} \pause
Learning Outcomes: ARTIST scaled question sets
    \begin{itemize}
        \item Hypothesis Testing (HT) score out of 10 
        \item Confidence Intervals (CI) score out of 10  
    \end{itemize}
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Data Description}
<<ScoreHistsByGroup, echo=FALSE , warning=FALSE, fig.width=15, fig.height=8, out.width='1\\linewidth', fig.pos='h',fig.align='center'>>=
dat$treatment <- "Simulation-Based"
dat$treatment[which(dat$room=="sweeney")] <- "Traditional"

smalldat <- dat[,c("treatment","midterm","ConfMC","HypMC","AppliedCI","AppliedHT")]
meltsmall <- melt(smalldat, id=c("treatment"))
trtsumstats <- ddply(dat, .(treatment), summarize,
                   avgmidterm = round(mean(midterm),1 ), 
                   sdmidterm = round(sd(midterm), 2 ), 
                   avgConfMC = round(mean(ConfMC),1 ), 
                   sdConfMC = round(sd(ConfMC), 2 ), 
                   avgHypMC = sprintf("%.1f",round(mean(HypMC),1 )), 
                   sdHypMC = round(sd(HypMC), 2 ), 
                   avgAppliedCI = round(mean(AppliedCI),1 ), 
                   sdAppliedCI = round(sd(AppliedCI), 2 ), 
                   avgAppliedHT = round(mean(AppliedHT), 1 ), 
                   sdAppliedHT = round(sd(AppliedHT), 2 ) , 
                   avgLab5 = round(mean(lab5perc), 1 ), 
                   sdLab5 = round(sd(lab5perc), 2 )  )

titlesize =1.5
xaxissize =1.3
sumstatfont = 6
yscale <- c(0,25)

themeForMultHist <- theme(plot.margin=unit(c(1,0,1,0), "cm"),
                          axis.title.y= element_blank() ,
                          strip.text.y = element_text(size = 15),
                          #strip.background = element_blank(),
                          plot.title = element_text(size = rel(titlesize)),
                          axis.title.x = element_text(size = rel(xaxissize)))

p0 <- qplot(midterm, data=dat, facets = treatment ~ ., binwidth=8, origin=4) + theme_bw()  +  
  xlab("Score (Out of 100)") + ggtitle("Midterm Exam") +
  #ylim(yscale) +
  geom_text(aes(x, y, label=lab), 
    data=data.frame(x=-1,y=12, lab=c(paste("mean =",trtsumstats[1,2]),paste("mean =",trtsumstats[2,2])) ,
    treatment=c("Simulation-Based","Traditional")), hjust=0, vjust=1, size= rel(sumstatfont)  )  + 
  geom_text(aes(x, y, label=lab), 
    data=data.frame(x=-1,y=.85*12, lab=c(paste("sd =",trtsumstats[1,3]), paste("sd =",trtsumstats[2,3])) ,
    treatment=c("Simulation-Based","Traditional")), hjust=0, vjust=1, size= rel(sumstatfont)    )+
  themeForMultHist + 
  scale_x_continuous(limits=c(-1, 100),breaks=seq(from=0, to=100, by=25))

p1 <- qplot(lab5perc, data=dat, facets = treatment ~ ., binwidth=10) + theme_bw() + 
  #ylim(yscale) + 
  xlab("Score (Out of 100)") + ylab("")  +  ggtitle("Lab 5") +
  geom_text(aes(x, y, label=lab), 
    data=data.frame(x=-1,y=38, lab=c(paste("mean =",trtsumstats[1,12]),
                                    paste("mean =",trtsumstats[2,12])) ,
    treatment=c("Simulation-Based","Traditional")), hjust=0, vjust=1 , size= rel(sumstatfont)   )  + 
  geom_text(aes(x, y, label=lab), 
    data=data.frame(x=-1,y=.85*38, lab=c(paste("sd =",trtsumstats[1,13]),
                                    paste("sd =",trtsumstats[2,13])) ,
    treatment=c("Simulation-Based","Traditional")), hjust=0, vjust=1 , size= rel(sumstatfont)   )+
  themeForMultHist+ 
  scale_x_continuous(limits=c(-1, 101),breaks=seq(from=0, to=100, by=25))

p2 <- qplot(ConfMC, data=dat, facets = treatment ~ ., binwidth=1) + theme_bw() + 
  xlab("Score (Out of 10)") + ylab("")  + ggtitle("ARTIST CI") +
  #ylim(yscale) +
  geom_text(aes(x, y, label=lab), 
    data=data.frame(x=-1,y=12, lab=c(paste("mean =",trtsumstats[1,4]),
                                    paste("mean =",trtsumstats[2,4])) ,
    treatment=c("Simulation-Based","Traditional")), hjust=0, vjust=1 , size= rel(sumstatfont)   )  + 
  geom_text(aes(x, y, label=lab), 
    data=data.frame(x=-1,y=.85*12, lab=c(paste("sd =",trtsumstats[1,5]),
                                    paste("sd =",trtsumstats[2,5])) ,
    treatment=c("Simulation-Based","Traditional")), hjust=0, vjust=1, size= rel(sumstatfont)    )+
  themeForMultHist + 
  scale_x_continuous(limits=c(-1, 11),breaks=seq(from=0, to=10, by=2))

p3 <- qplot(HypMC, data=dat, facets = treatment ~ ., binwidth=1) + theme_bw() + 
  #ylim(yscale) + 
  xlab("Score (Out of 10)") + ylab("")  +  ggtitle("ARTIST HT") +
  geom_text(aes(x, y, label=lab), 
    data=data.frame(x=-1,y=13, lab=c(paste("mean =",trtsumstats[1,6]),
                                    paste("mean =",trtsumstats[2,6])) ,
    treatment=c("Simulation-Based","Traditional")), hjust=0, vjust=1 , size= rel(sumstatfont)   )  + 
  geom_text(aes(x, y, label=lab), 
    data=data.frame(x=-1,y=.85*13, lab=c(paste("sd =",trtsumstats[1,7]),
                                    paste("sd =",trtsumstats[2,7])) ,
    treatment=c("Simulation-Based","Traditional")), hjust=0, vjust=1 , size= rel(sumstatfont)   )+
  themeForMultHist + 
  scale_x_continuous(limits=c(-1, 11),breaks=seq(from=0, to=10, by=2))

grid.arrange(textGrob("", rot = 90, gp = gpar(cex = 1.5)),
             arrangeGrob(p0,p1,p2,p3, nrow=1), 
             #arrangeGrob(textGrob("Simulation-Based", vjust = 1,
             #                     gp = gpar(cex = 1.5), rot = 90),
             #            textGrob("Traditional", vjust = 1, 
             #                     gp = gpar(cex = 1.5), rot = 90)),
             widths=unit.c(unit(2, "lines"), unit(1, "npc") - unit(2, "lines")),
             nrow=1
             )
@

\center Histograms and summary statistics of scores by curriculum group

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model}
Modeling approach based on data characteristics
    \begin{itemize}
        \item Bivariate responses: ARTIST scores for CI and HT
        \item Binary curricula indicator
        \item Continuous covariates: pre-treatment scores
    \end{itemize} \pause
Multivariate Analysis of Covariance (MANCOVA) Model
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model}

Bivariate MANCOVA Model Structure

\begin{eqnarray}\label{eq:mancova}
y_{ik} = \tau_{k} \mathbbm{1}_{\{i \in T\}} + \beta_{0k} + \sum_{p=1}^{P}x_{ip}\beta_{pk} + \epsilon_{ik},
\end{eqnarray}

 \begin{tabular}{lp{5in}}
 $y_{ik}$ &  $k^{th}$ response ($k \in \left\{1,2\right\}$) from  student $i$, $ 1 \le i \le n$,\\
 $\tau_{k}$ & treatment effect on response $k$, and \\
 $\mathbbm{1}_{\{i \in T\}}$ & indicator function student $i$ in Simulation-Based group\\
 $\beta_{0k}$ & common intercept for response $k$, and \\
 $\beta_{kp}$, & $1 \le p \le P$ model coefficients of $P$ covariates. \\
 $x_{ip}$ & $p^{th}$ pre-treatment covariate score of student $i$\\
 %$\epsilon}_{ik}$ & error for $k^{th}$ response from $i^{th}$ student. 
 \end{tabular}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model}

Bivariate MANCOVA Model Error Structure

\[
\vec{\epsilon}_{i} = 
\begin{bmatrix}
  \epsilon_{i1} \\ \epsilon_{i2} 
 \end{bmatrix}  
 \distas{iid} \text{MVN} \left( 
 \begin{bmatrix}
  0 \\ 0 
 \end{bmatrix},
 \Sigma = \begin{bmatrix}
  \sigma_{11}^2 & \sigma_{12}^2 \\ 
  \sigma_{21}^2 & \sigma_{22}^2
 \end{bmatrix}
 \right)
\]

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model Selection}

Model Selection Process
    \begin{itemize}
        \item Bivariate MANCOVA model fit initially with all covariates
        \item Backward stepwise selection process based on AIC
        \item Additional covariates removed due to collinearity
    \end{itemize}
Selected Parameters for ARTIST Model
    \begin{itemize}
        \item Midterm exam score
        \item Lab 5 score
        \item Curriculum treatment effect
    \end{itemize}
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model Results: Bivariate Effects}

\begin{table}[hbtp]
\centering
\begin{tabular}{lrrr} \hline
 & Pillai's $\Lambda$ & Approx. F & Pr($>$$|$F$|$)\\ 
 \hline  
Midterm & \Sexpr{sprintf("%.4f",round(summarytab[1,2],4))} & \Sexpr{sprintf("%.4f",round(summarytab[1,3],4))} &
\Sexpr{sprintf("%.4f",round(summarytab[1,6],6))}\\ 
Lab 5 & \Sexpr{sprintf("%.4f",round(summarytab[2,2],4))} & \Sexpr{sprintf("%.4f",round(summarytab[2,3],4))} &
\Sexpr{sprintf("%.4f",round(summarytab[2,6],6))}\\ 
Treatment & \Sexpr{sprintf("%.4f",round(summarytab[3,2],4))} & \Sexpr{sprintf("%.4f",round(summarytab[3,3],4))} &
\Sexpr{sprintf("%.4f",round(summarytab[3,6],6))}\\ 
\hline
\end{tabular}
\end{table}

\center Pillai's $\Lambda$ tests for overall covariate effects

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model Results: Univariate Effects}

\begin{table}[H]
\centering
\begin{tabular}{rccc}
  \hline
 & Covariate Values & Estimate & 95\% Confidence Interval \\ 
  \hline
Intercept & 1 & \Sexpr{sprintf("%.4f",round(HTcis[1,3],6))} & \Sexpr{paste("(",sprintf("%.4f",round(HTcis[1,1],6)),",", sprintf("%.4f",round(HTcis[1,2],6)),")",sep=" ")} \\ 
  Midterm & \{0,1,...100\} & \Sexpr{sprintf("%.4f",round(HTcis[2,3],6))} & \Sexpr{paste("( ",sprintf("%.4f",round(HTcis[2,1],6)),",", sprintf("%.4f",round(HTcis[2,2],6)),")",sep=" ")} \\ 
  Lab 5  &  \{0,1,...100\} & \Sexpr{sprintf("%.4f",round(HTcis[3,3],6))} & \Sexpr{paste("( ",sprintf("%.4f",round(HTcis[3,1],6)),",", sprintf("%.4f",round(HTcis[3,2],6)),")",sep=" ")} \\ 
  Treatment &  \{0,1\} & \Sexpr{sprintf("%.4f",round(HTcis[4,3],6))} & \Sexpr{paste("( ",sprintf("%.4f",round(HTcis[4,1],6)),",", sprintf("%.4f",round(HTcis[4,2],6)),")",sep=" ")} \\ 
   \hline
\end{tabular}
\end{table}

\center ARTIST Model coefficients for hypothesis testing topic score

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model Results: Univariate Effects}

\begin{table}[hbtp]
\centering
\begin{tabular}{rccc}
  \hline
 & Covariate Values & Estimate & 95\% Confidence Interval \\ 
  \hline
Intercept & 1 & \Sexpr{sprintf("%.4f",round(CIcis[1,3],6))} & \Sexpr{paste("(",sprintf("%.4f",round(CIcis[1,1],6)),",", sprintf("%.4f",round(CIcis[1,2],6)),")",sep=" ")} \\ 
  Midterm & \{0,1,...100\} & \Sexpr{sprintf("%.4f",round(CIcis[2,3],6))} & \Sexpr{paste("( ",sprintf("%.4f",round(CIcis[2,1],6)),",", sprintf("%.4f",round(CIcis[2,2],6)),")",sep=" ")} \\ 
  Lab 5  &  \{0,1,...100\} & \Sexpr{sprintf("%.4f",round(CIcis[3,3],6))} & \Sexpr{paste("( ",sprintf("%.4f",round(CIcis[3,1],6)),",", sprintf("%.4f",round(CIcis[3,2],6)),")",sep=" ")} \\ 
  Treatment &  \{0,1\} & \Sexpr{sprintf("%.4f",round(CIcis[4,3],6))} & \Sexpr{paste("( ",sprintf("%.4f",round(CIcis[4,1],6)),",", sprintf("%.4f",round(CIcis[4,2],6)),")",sep=" ")} \\ 
   \hline
\end{tabular}
\end{table}
\center ARTIST Model coefficients for confidence interval topic score

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model Assessment}

Model Assumptions Needing Assessment
    \begin{itemize}
        \item Normality of errors
        \item Linearity 
        \item Constant variance
        \item Independence between students
    \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model Assessment}
<<ARTISTModDiagPlots, echo=FALSE , warning=FALSE, fig.width=4, fig.height=4, out.width='.33\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
#qqp1 <- 
qplot(sample=mod2small$residuals[,1], stat="qq", xlab="Theoretical Normal Quantiles",
       ylab="ARTIST CI Score Residual") +  theme_bw() + theme(aspect.ratio=1)

#qqp2 <- 
  qplot(sample=mod2small$residuals[,2], stat="qq", xlab="Theoretical Normal Quantiles",
      ylab="ARTIST HT Score Residual") +  theme_bw() + theme(aspect.ratio=1)

#grid.arrange(qqp1,qqp2,nrow=2)

qplot(mod2small$residuals[,1],mod2small$residuals[,2],
      xlab="ARTIST CI Score Residual",  ylab="ARTIST HT Score Residual")  +  theme_bw()
@
\center Normal quantile plots (left \& center) \\

\center Bivariate residuals scatterplot (right)

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model Assessment}
<<residualsARTIST, echo=FALSE , message=FALSE, warning=FALSE, fig.width=8, fig.height=3.2, out.width='.9\\linewidth', fig.pos='h',fig.align='center'>>=
## check residual plots for homoscedesticity and linearity
jitheight <- .1
p1 <- qplot(mod2small$fitted.values[,1] ,mod2small$residuals[,1], geom = "jitter") + 
  geom_hline(yintercept=0) +geom_jitter(position = position_jitter(height = jitheight )) +
  xlab("ARTIST CI Score Fitted Value") + ylab("ARTIST CI Score Residual") +  theme_bw() + geom_smooth()
p2 <- qplot(mod2small$fitted.values[,2] ,mod2small$residuals[,2]) + 
  geom_hline(yintercept=0) +geom_jitter(position = position_jitter(height = jitheight )) +
  xlab("ARTIST HT Score Fitted Value") + ylab("ARTIST HT Score Residual") +  theme_bw() + geom_smooth()
grid.arrange(p1,p2,nrow=1)
# these seem fine (put into appendix)
@

\center ARTIST Model residual plots overlaid with Loess smoother
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model Assessment}

Recall:
\[
\vec{\epsilon}_{i} = 
\begin{bmatrix}
  \epsilon_{i1} \\ \epsilon_{i2} 
 \end{bmatrix}  
 \distas{iid} \text{MVN} \left( 
 \begin{bmatrix}
  0 \\ 0 
 \end{bmatrix},
 \Sigma = \begin{bmatrix}
  \sigma_{11}^2 & \sigma_{12}^2 \\ 
  \sigma_{21}^2 & \sigma_{22}^2
 \end{bmatrix}
 \right)
\]

Assumption of Independence
    \begin{itemize}
        \item Student score pairs are related:  $\Cov[\epsilon_{i1}, \epsilon_{i2}] = \sigma_{12}^2$
        \item Scores between students independent: $\Cov[\epsilon_{ik}, \epsilon_{j\ell}] = 0$, $\forall i\ne j$
    \end{itemize} \pause

    \vspace{.1in}

Impact of Violation of Independence
    \begin{itemize}
        \item Type I error rate when testing for curriculum treatment effect
        \item Bias in parameter estimation
    \end{itemize}
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Model Assessment}

Simulation Study
    \begin{itemize}
        \item Simulate responses under independence violation
        \item Assume MANCOVA model as generative model
          \begin{itemize}
            \item Set coefficient values, $\beta_{pk}$ for intercept, midterm and lab 5 based estimates from reduced MANCOVA
            \item No curriculum treatment effect (i.e. $\tau_1 = \tau_2 = 0$)
          \end{itemize}
    \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study}

Assumption: 
\[
\Cov[\epsilon_{ik}, \epsilon_{j\ell}] = 0, \  \text{ for all } i \ne j
\]

\vspace{.05in}

Independence Violation:\\

\begin{itemize}
\item Stage 1:
\end{itemize}
    \[
\Cov[\epsilon_{ik}, \epsilon_{j\ell}] = \eta \sigma_{k\ell}^2, \ \text{ where } \eta \in [0,1]  \text{ for all } i \ne j, \text{ and } k,\ell \in \{1,2\}
\]

\begin{itemize}
\item Stage 2:
\end{itemize}
    \[
\Cov[\epsilon_{ik}, \epsilon_{j\ell}] = c \eta \sigma_{k\ell}^2, \ \text{ where } \eta \in [0,1]  \text{ for all } i \ne j, \text{ and } k,\ell \in \{1,2\}    
\]
\hspace{1.5in} $c = 1$ for all i,j classmates\\
\hspace{1.5in} $c \in [0,1]$ for all i,j non-classmates

\end{frame}

% %-------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study}

\center No Violation of Between Student Independence

\begin{equation} \hspace*{-1cm} \footnotesize
\begin{bmatrix} 
 \epsilon_{11} \\ \vdots \\ \epsilon_{n1} \\ \epsilon_{12} \\ \vdots \\ \epsilon_{n2} 
\end{bmatrix} \sim \text{N}\left(\tiny
\vec{0} , 
 \left[\begin{array}{ccccc|ccccc}
\sigma_{11}^2 & 0 & \cdots & 0 & 0 & \sigma_{12}^2 & 0 & \cdots & 0 & 0\\
0 & \sigma_{11}^2 & \ddots & 0 & 0 & 0 & \sigma_{12}^2 & \ddots & 0 & 0\\
\vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
0 & 0 & \ddots & \sigma_{11}^2 & 0 & 0 & 0 & \ddots & \sigma_{12}^2 & 0\\
0 & 0 & \cdots & 0 & \sigma_{11}^2 & 0 & 0 & \cdots & 0 & \sigma_{12}^2\\
\hline
\sigma_{12}^2 & 0 & \cdots & 0 & 0 & \sigma_{22}^2 & 0 & \cdots & 0 & 0\\
0 & \sigma_{12}^2 & \ddots & 0 & 0 & 0 & \sigma_{22}^2 & \ddots & 0 & 0\\
\vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
0 & 0 & \ddots & \sigma_{12}^2 & 0 & 0 & 0 & \ddots & \sigma_{22}^2 & 0\\
0 & 0 & \cdots & 0 & \sigma_{12}^2 & 0 & 0 & \cdots & 0 & \sigma_{22}^2\\  
 \end{array}\right]
 \right) \nonumber 
 \end{equation}

\end{frame}

% %-------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study}

\center Stage 1 Violation of Between Student Independence

\begin{equation} \hspace*{-.6cm} \footnotesize
\begin{bmatrix} 
 \epsilon_{11} \\ \vdots \\ \epsilon_{n1} \\ \epsilon_{12} \\ \vdots \\ \epsilon_{n2} 
\end{bmatrix}\sim \text{N}\left(\tiny
\vec{0} , 
 \left[\begin{array}{ccccc|ccccc}
  \sigma_{11}^2 & \eta \sigma_{11}^2 & \cdots & \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2\\
  \eta \sigma_{11}^2 & \sigma_{11}^2 & \ddots & \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \eta \sigma_{12}^2 & \sigma_{12}^2 & \ddots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2\\
\vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
\eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \ddots & \sigma_{11}^2 & \eta \sigma_{11}^2 & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \ddots & \sigma_{12}^2 & \eta \sigma_{12}^2\\
\eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \cdots & \eta \sigma_{11}^2 & \sigma_{11}^2 & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \sigma_{12}^2\\
\hline
\sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \sigma_{22}^2 & \eta \sigma_{22}^2 & \cdots & \eta \sigma_{22}^2 & \eta \sigma_{22}^2\\
  \eta \sigma_{12}^2 & \sigma_{12}^2 & \ddots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \eta \sigma_{22}^2 & \sigma_{22}^2 & \ddots & \eta \sigma_{22}^2 & \eta \sigma_{22}^2\\
\vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
\eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \ddots & \sigma_{12}^2 & \eta \sigma_{12}^2 & \eta \sigma_{22}^2 & \eta \sigma_{22}^2 & \ddots & \sigma_{22}^2 & \eta \sigma_{22}^2\\
\eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \sigma_{12}^2 & \eta \sigma_{22}^2 & \eta \sigma_{22}^2 & \cdots & \eta \sigma_{22}^2 & \sigma_{22}^2\\  
 \end{array}\right]
 \right) \nonumber 
 \end{equation}

\end{frame}

%---------------------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study}

\center Stage 2 Violation of Between Student Independence

\begin{equation} \hspace*{-1cm} \footnotesize
\begin{bmatrix} 
 \epsilon_{11} \\ \vdots \\ \epsilon_{n1} \\ \epsilon_{12} \\ \vdots \\ \epsilon_{n2} 
\end{bmatrix}\sim \text{N}\left(\tiny
\vec{0} , 
 \left[\begin{array}{ccccc|ccccc}
  \sigma_{11}^2 &  \eta \sigma_{11}^2 & \cdots & c \eta \sigma_{11}^2 & c \eta \sigma_{11}^2 & \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & c \eta \sigma_{12}^2 & c\eta \sigma_{12}^2\\
  \eta \sigma_{11}^2 & \sigma_{11}^2 & \ddots & c\eta \sigma_{11}^2 & c\eta \sigma_{11}^2 & \eta \sigma_{12}^2 & \sigma_{12}^2 & \ddots & c\eta \sigma_{12}^2 & c\eta \sigma_{12}^2\\
\vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
c\eta \sigma_{11}^2 & c\eta \sigma_{11}^2 & \ddots & \sigma_{11}^2 & \eta \sigma_{11}^2 & c\eta \sigma_{12}^2 & c\eta \sigma_{12}^2 & \ddots & \sigma_{12}^2 & \eta \sigma_{12}^2\\
c\eta \sigma_{11}^2 & c\eta \sigma_{11}^2 & \cdots & \eta \sigma_{11}^2 & \sigma_{11}^2 & c\eta \sigma_{12}^2 & c\eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \sigma_{12}^2\\
\hline
\sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & c\eta \sigma_{12}^2 & c\eta \sigma_{12}^2 & \sigma_{22}^2 & \eta \sigma_{22}^2 & \cdots & c\eta \sigma_{22}^2 & c\eta \sigma_{22}^2\\
\eta \sigma_{12}^2 & \sigma_{12}^2 & \ddots & c\eta \sigma_{12}^2 & c\eta \sigma_{12}^2 & \eta \sigma_{22}^2 & \sigma_{22}^2 & \ddots & c\eta \sigma_{22}^2 & c\eta \sigma_{22}^2\\
\vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
c\eta \sigma_{12}^2 & c\eta \sigma_{12}^2 & \ddots & \sigma_{12}^2 & \eta \sigma_{12}^2 & c\eta \sigma_{22}^2 & c\eta \sigma_{22}^2 & \ddots & \sigma_{22}^2 & \eta \sigma_{22}^2\\
c\eta \sigma_{12}^2 & c\eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \sigma_{12}^2 & c\eta \sigma_{22}^2 & c\eta \sigma_{22}^2 & \cdots & \eta \sigma_{22}^2 & \sigma_{22}^2\\  
 \end{array}\right]
 \right) \nonumber 
 \end{equation}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study}

Simulation Process: Repeat M times
%
\begin{enumerate}
\item Simulate an error vector, $\vec{\epsilon}^{\hspace{.1cm}(m)}$, under violated error structure.
\item Derive responses:
%
\[
\begin{bmatrix}
  \vec{y}_1^{\hspace{.1cm}(m)} \\ \vec{y}_2^{\hspace{.1cm}(m)}
  \end{bmatrix}   =
   % %----
 \begin{bmatrix}
  X\vec{\beta}_1 \\ X\vec{\beta}_2
  \end{bmatrix} + 
  \vec{\epsilon}^{\hspace{.1cm}(m)}
\]
%
\item Randomly permute treatment label to break association with covariates
\item Fit full ARTIST Model (assuming independence)
\item Record estimates, standard errors, etc. 
\end{enumerate}
\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study}

Stage 1:
    \begin{itemize}
        \item Repeat $M = 25,000$ simulations at each $\eta$ in $\{0,0.1,0.2,0.3,0.4,0.5\}$
    \end{itemize}

    \vspace{.1in}

Stage 2:
    \begin{itemize}
        \item Repeat $M = 10,000$ simulations at each combination of $\eta$ in $\{0,0.1,0.2,0.3,0.4,0.5\}$ and $c$ in $\{0.9, 0.925, 0.95, 0.975, 1\}$
    \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study: Stage 1 Results}

\begin{table}[hbtp]
\centering
\begin{tabular}{cccc} \hline
$\eta$  & Pillai's Test & t-test (CI) & t-test (HT) \\
& $\text{H}_o : \tau_1 = \tau_2 = 0$  & $\text{H}_o : \tau_1 = 0$ &  $\text{H}_o : \tau_2 = 0$\\ 
 \hline  
\Sexpr{sprintf("%.1f",round(errorsummary[1,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[1,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[1,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[1,6],4))} \\
\Sexpr{sprintf("%.1f",round(errorsummary[2,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[2,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[2,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[2,6],4))} \\
\Sexpr{sprintf("%.1f",round(errorsummary[3,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[3,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[3,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[3,6],4))} \\
\Sexpr{sprintf("%.1f",round(errorsummary[4,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[4,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[4,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[4,6],4))} \\
\Sexpr{sprintf("%.1f",round(errorsummary[5,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[5,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[5,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[5,6],4))} \\
\Sexpr{sprintf("%.1f",round(errorsummary[6,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[6,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[6,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[6,6],4))} \\
\hline
\end{tabular}
\end{table}

\center Type I error rates for tests for curriculum treatment effect

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study: Stage 1 Results}

<<simhistogrid, echo=FALSE , message=FALSE, warning=FALSE, fig.width=10, fig.height=6, out.width='1\\linewidth', fig.pos='h',fig.align='center',cache=TRUE>>=
library(ggplot2)
library(reshape2)
etas <- seq(from=0, to=0.5, by=.1) 
plotbetadata <- data.frame(eta = rep(as.character(etas),each=M))
plotbetadata <- cbind(plotbetadata, rbind(resultlist[[1]][,c(1,2,7,8,13,14)],
                                          resultlist[[2]][,c(1,2,7,8,13,14)],
                                          resultlist[[3]][,c(1,2,7,8,13,14)],
                                          resultlist[[4]][,c(1,2,7,8,13,14)],
                                          resultlist[[5]][,c(1,2,7,8,13,14)],
                                          resultlist[[6]][,c(1,2,7,8,13,14)]) )
names(plotbetadata) <- c("eta","beta[11]","beta[12]","beta[21]","beta[22]","tau[1]","tau[2]")
plotbetadata$eta <- paste("eta==",plotbetadata$eta,sep="")
meltplotbetadata <- melt(plotbetadata, id=c("eta"))
trueeffects <- unique(meltplotbetadata[,1:2])
trueeffects$value <-  rep(c(0.046132,0.037915,0.017485, 0.008211,0,0),each=6)
number_ticks <- function(n) {function(limits) pretty(limits, n)}
qplot(value, geom="histogram", data=meltplotbetadata,fill=I("gray42")) + 
  facet_grid(eta~variable, scales = "free",labeller = label_parsed) +
  geom_vline(aes(xintercept = value), data=trueeffects, color="black")+ theme_bw()+
  ylab("") + xlab("Estimated Parameter Values") +
  scale_x_continuous(breaks=number_ticks(3))
@
\center Parameter estimates from 25,000 simulations at each $\eta$.

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study: Stage 1 Results}
\begin{table}[hbtp]
\centering
\begin{tabular}{ccccc} \hline
$\eta$ & $s_{\tau_1}$ & $\hat{\text{SE}}\left[\tau_1\right]$ & $s_{\tau_2}$ & $\hat{\text{SE}}\left[\tau_2\right]$\\ 
 \hline  
\Sexpr{sprintf("%.1f",round(simresults[1,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[1,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[1,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[1,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[1,17],4))} \\
\Sexpr{sprintf("%.1f",round(simresults[2,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[2,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[2,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[2,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[2,17],4))} \\
\Sexpr{sprintf("%.1f",round(simresults[3,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[3,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[3,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[3,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[3,17],4))} \\
\Sexpr{sprintf("%.1f",round(simresults[4,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[4,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[4,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[4,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[4,17],4))} \\
\Sexpr{sprintf("%.1f",round(simresults[5,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[5,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[5,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[5,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[5,17],4))} \\
\Sexpr{sprintf("%.1f",round(simresults[6,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[6,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[6,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[6,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[6,17],4))} \\
\hline
\end{tabular}
\end{table}

\center Standard deviation for simulated treatment effect estimates, $s_{\tau_k}$, and average standard errors for treatment effects, $\hat{\text{SE}}\left[\tau_k\right]$

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study: Stage 2 Results}

<<sim2plots, echo=FALSE , message=FALSE, warning=FALSE, fig.width=8, fig.height=3, out.width='1\\linewidth', fig.pos='h',fig.align='center', cache=TRUE>>=
library(ggplot2)
errorsummary2 <- read.csv("C:\\Users\\Karsten\\Dropbox\\Dissertation\\CurriculumStudy/SimulationStudy/Sim1000ClassCovResults.csv")
meltsim2 <- melt(errorsummary2, id=c("eta", "nonclassmate"))
meltsim2$variable2 <- factor(meltsim2$variable, labels = c("H[o]: tau[1]==0", "H[o]: tau[2]==0", "Pillai"))
qplot(eta, value, geom="path", data=meltsim2, colour=as.factor(nonclassmate), group=nonclassmate, size=I(1.2)) +
  geom_point(size=I(1.6),colour=("Black"))+
   facet_grid(. ~ variable2, labeller = label_parsed) +
  xlab(expression(eta)) + ylab("Type I Error Rate") + 
  theme_bw() + labs(colour='Non-classmate Variance
Scaling Parameter (c)') + geom_hline(yintercept=0.05, linetype=2)+
   geom_hline(yintercept=0, linetype=1)+
  ylim(c(0,.50))+
  scale_colour_manual(values = c("gray85","gray70","gray55","gray40","gray25"))
@
\center Type I error rates for tests of treatment effects

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Simulation Study: Stage 2 Results}

Type I error rates increase as $c$ increases

\vspace{.1in}

What does $c$ represent?
    \begin{itemize}
        \item covariance disparity attributable to \textit{non-treatment} factors 
        \item generative model must carry no curriculum treatment effect
    \end{itemize}

\vspace{.2in}

Cannot estimate $c$ due to classroom confounding with treatment

\vspace{.1in}

Steps taken to minimize non-treatment differences in classroom
    \begin{itemize}
        \item Alternation of instruction
        \item Identical curricula weeks 1 to 8
        \item Physical classrooms similar
    \end{itemize}

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Discussion and Conclusions}
Study Results:
    \begin{itemize}
      \item Simulation-based group scored significantly higher on CI 
        \begin{itemize}
          \item After controlling for midterm and lab 5 scores
          \item 7\% higher on ARTIST scale
        \end{itemize}
      \item No significant difference in HT scores
    \end{itemize} \pause

Discussion Issues:
    \begin{itemize}
      \item Criticisms of ARTIST sets
      \item Representation 
      \item Causality
      \item Treatment Complexity
    \end{itemize}

\end{frame}

% %-------------------------------------------------
% 
% \begin{frame}
%     \frametitle{Discussion and Conclusions}
% Criticisms of ARTIST sets
%     \begin{itemize}
%         \item Increasingly outdated
%         \item Lack of nationally normed reliability study
%     \end{itemize} 
% 
% \vspace{.1in}
% 
% Representation
%     \begin{itemize}
%         \item Ag/Bio students at large public Midwestern university
%         \item Results apply as far as students are representative
%     \end{itemize} 
% 
% \end{frame}
% 
% %-------------------------------------------------
% 
% \begin{frame}
%     \frametitle{Discussion and Conclusions}
% Causality requires assumptions
%     \begin{itemize}
%         \item Randomization eliminated effect of extraneous variables
%         \item Alternation eliminated instructor effect
%     \end{itemize} 
% 
% \vspace{.1in}
% 
% Treatment Complexity
%     \begin{itemize}
%         \item 8 week curricula (lecture, notes, lab, homework)
%         \item Cannot dissect to identify which parts led to improvement
%     \end{itemize} 
% 
% \end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Acknowledgements}

Co-investigator: Dennis Lock \\

\vspace{.2in}

Study Advisors: Dr. Heike Hofmann and Dr. Bob Stephenson \\

\end{frame}

%-------------------------------------------------

\begin{frame}
    \frametitle{Thanks!}

\center Thank you for attending! \\

\center Any questions?

\end{frame}

\end{document}