### 2015-06-30 Appendix/Appendix1.Rnw
Major changes following committee discussion in defense.

% Appendix1 file from standard thesis template
\chapter{ }

<<setupAppendix,echo=F,include=F,eval=T, cache=T>>=
### Preliminaries 
## Load Packages
library(car)
library(ggplot2)
library(subselect)
library(qtlmt)
library(multcomp)
library(lattice)
library(gridExtra)
library(reshape2)
library(plyr)
library(boot)

## Load Data
dat <- read.csv("DataForStudyAnalysis.csv",header=T)
#change factor level order to put sweeney as default control group
dat$room <- factor(dat$room.x, levels=c("sweeney", "kildee"))
dat$lab5perc <- dat$lab5perc * 100
dat$hw2perc <- dat$hw2perc * 100
dat$treatment <- "Simulation-Based"
dat$treatment[which(dat$room=="sweeney")] <- "Traditional"
@

%-------------------------------------------------------

\section{Appendix: Final Exam Used in Curricula Study}
\label{appendA}

The following appendix contains the questions and point rubrics for the ARTIST scaled questions and the applied theory-based inference problems as they appeared on the final exam.

\subsection{ARTIST Scaled Multiple Choice Question Set for Confidence Intervals}
\label{appendA1}

1. Answer the following general multiple choice questions regarding confidence intervals.  There is only one correct answer for each (circle the best option).

i. Two different samples will be taken from the same population of test scores where the population mean and standard deviation are unknown. The first sample will have 25 data values, and the second sample will have 64 data values. A 95\% confidence interval will be constructed for each sample to estimate the population mean. Which confidence interval would you expect to have greater precision (a smaller width) for estimating the population mean?

a. I expect the confidence interval based on the sample of 64 data values to be more precise.\\
b. I expect both confidence intervals to have the same precision.\\
c. I expect the confidence interval based on the sample of 25 data values to be more precise.\\

ii. A 95\% confidence interval is computed to estimate the mean household income for a city. Which of the following values will definitely be within the limits of this confidence interval?

a. The population mean\\
b. The sample mean\\
c. The standard deviation of the sample mean\\
d. None of the above\\

iii. Each of the 110 students in a statistics class selects a different random sample of 35 Quiz scores from a population of 5000 scores they are given. Using their data, each student constructs a 90\% confidence interval for $\mu$ the average Quiz score of the
5000 students. Which of the following conclusions is correct?

a. About 10\% of the sample means will not be included in the confidence intervals.\\
b. About 90\% of the confidence intervals will contain $\mu$.\\
c. It is probable that 90\% of the confidence intervals will be identical.\\
d. About 10\% of the raw scores in the samples will not be found in these confidence intervals\\

iv. A 95\% confidence interval for the mean reading achievement score for a population of third grade students is (43, 49). The margin of error of this interval is:

a. 5\\
b. 3\\
c. 6\\

v. Justin and Hayley conducted a mission to a new planet, Planet X, to study arm length. They took a random sample of 100 Planet X residents and calculated a 95\% confidence interval for the mean arm length. What does a 95\% confidence interval for arm length tell us in this case? Select the best answer:

a. I am 95\% confident that this interval includes the sample mean (x) arm length.\\
b. I am confident that most (95\%) of all Planet X residents will have an arm length within this interval.\\
c. I am 95\% confident that most Planet X residents will have arm lengths within this interval.\\
d. I am 95\% confident that this interval includes the population mean arm length.\\

vi. Suppose that a random sample of 41 state college students is asked to measure the length of their right foot in centimeters. A 95\% confidence interval for the mean foot length for students at this university turns out to be (21.709, 25.091). If instead a 90\% confidence interval was calculated, how would it differ from the 95\% confidence interval?

a. The 90\% confidence interval would be narrower.\\
b. The 90\% confidence interval would be wider.\\
c. The 90\% confidence interval would be the same as the 95\% confidence interval.\\

vii. A pollster took a random sample of 100 students from a large university and computed a confidence interval to estimate the percentage of students who were planning to vote in the upcoming election. The pollster felt that the confidence interval was too wide to provide a precise estimate of the population parameter. What could the pollster have done to produce a narrower confidence interval that would produce a more precise estimate of the percentage of all university students who plan to vote in the upcoming election?

a. Increase the sample size to 150.\\
b. Increase the confidence level to 99\%.\\
c. Both a and b\\
d. None of the above\\

viii. A newspaper article states with 95\% confidence that 55\% to 65\% of all high school students in the United States claim that they could get a hand gun if they wanted one. This confidence interval is based on a poll of 2000 high school students in Detroit. How would you interpret the confidence interval from this newspaper article?

a. 95\% of large urban cities in the United States have 55\% to 65\% high school students who could get a hand gun.\\
b. If we took many samples of high school students from different urban cities, 95\% of the samples would have between 55\% and 65\% high school students who could get hand guns.\\
c. You cannot use this confidence interval to generalize to all teenagers in the United States because of the way the sample was taken.\\
d. We can be 95\% confident that between 55\% and 65\% of all United States high school students could get a hand gun.\\

ix. The Gallup poll (August 23, 2002) reported that 53\% of Americans said they would favor sending American ground troops to the Persian Gulf area in an attempt to remove Hussein from power. The poll also reported that the margin of error for this poll was 4\%. What does the margin of error of 4\% indicate?

a. There is a 4\% chance that the estimate of 53\% is wrong.\\
b. The percent of Americans who are in favor is probably higher than 53\% and closer to 57\%.\\
c. The percent of Americans who are in favor is estimated to be between 49\% and 57\%.\\

x. Suppose two researchers want to estimate the proportion of American college students who favor abolishing the penny. They both want to have about the same margin of error to estimate this proportion. However, Researcher 1 wants to estimate with 99\% confidence and Researcher 2 wants to estimate with 95\% confidence. Which researcher would need more students for her study in order to obtain the desired margin of error?

a. Researcher 1.\\
b. Researcher 2.\\
c. Both researchers would need the same number of subjects.\\
d. It is impossible to obtain the same margin of error with the two different confidence levels.\\

\subsection{ARTIST Scaled Multiple Choice Question Set for Hypothesis Testing}

2.   Answer the following general multiple choice questions regarding hypothesis testing.  There is only one correct answer for each (circle the best option).  As a helpful note the term ``statistically significant'' means that you reject the null hypothesis.

i. The makers of Mini-Oats cereal have an automated packaging machine that is set to fill boxes with 24 ounces of cereal. At various times in the packaging process, a random sample of 100 boxes is taken to see if the machine is filling the boxes with an average of 24 ounces of cereal. Which of the following is a statement of the null hypothesis being tested?

a. The machine is filling the boxes with the proper amount of cereal.\\
b. The machine is not filling the boxes with the proper amount of cereal.\\
c. The machine is not putting enough cereal in the boxes.\\

ii. A research article gives a p-value of .001 in the analysis section. Which definition of a p-value is the most accurate?

a. the probability that the observed outcome will occur again.\\
b. the probability of observing an outcome as extreme or more extreme than the one observed if the null hypothesis is true.\\
c. the value that an observed outcome must reach in order to be considered significant under the null hypothesis.\\
d. the probability that the null hypothesis is true.\\

iii. If a researcher was hoping to show that the results of an experiment were statistically significant they would prefer:

a. a large p-value\\
b. a small p-value\\
c. p-values are not related to statistical significance\\

iv. A researcher compares men and women on 100 different variables using a difference in means t-test.  He sets the level of significance at 0.05 and then carries out 100 independent t-tests (one for each variable) on these data.  If, for each test, the null hypothesis is actually true, about how many ``statistically significant'' results will be produced?

a. 0\\
b. 5\\
c. 10\\
d. none of the above\\

Problems (v) and (vi) refer to the following situation: Food inspectors inspect samples of food products to see if they are safe. This can be thought of as a hypothesis test where Null: the food is safe, and Alternative: the food is not safe. Identify each of the following statements as a Type I or a Type II error.

v. The inspector says the food is safe but it actually is not safe.

a. Type I\\
b. Type II\\

vi. The inspector says the food is not safe but is actually safe.

a. Type I\\
b. Type II\\

vii. A newspaper article claims that the average age for people who receive food stamps is 40 years. You believe that the average age is less than that. You take a random sample of 100 people who receive food stamps, and find their average age to be 39.2 years. You find that this is significantly lower than the age of 40 stated in the article (p-value < .05). What would be an appropriate interpretation of this result?

a. The statistically significant result indicates that the majority of people who receive food stamps is younger than 40.\\
b. Although the result is statistically significant, the difference in age is not of practical importance.\\
c. An error must have been made. This difference is too small to be statistically significant.\\

viii. A newspaper article stated that the US Supreme Court received 812 letters from around the country on the subject of whether to ban cameras from the courtroom. Of these 812 letters, 800 expressed the opinion that cameras should be banned. A statistics student was going to use this sample information to conduct a test of significance of whether more than 95\% of all American adults feel that cameras should be banned from the courtroom. What would you tell this student?

a. This is a large enough sample to provide an accurate estimate of the American public's opinion on the issue.\\
b. The necessary conditions for a test of significance are not satisfied, so no statistical test should be performed.\\
c. With such a large number of people favoring the notion that cameras be banned, there is no need for a statistical test.\\

ix. A researcher conducts an experiment on human memory and recruits 15 people to participate in her study. She performs the experiment and analyzes the results. She obtains a p-value of .17. Which of the following is a reasonable interpretation of her results?

a. This proves that her experimental treatment has no effect on memory.\\
b. There could be a treatment effect, but the sample size was too small to detect it.\\
c. She should reject the null hypothesis.\\
d. There is evidence of a small effect on memory by her experimental treatment.\\

x. It is reported that scores on a particular test of historical trivia given to high school students are approximately normally distributed with a mean of 85. Mrs. Rose believes that her 5 classes of high school seniors will score significantly better than the national average on this test. At the end of the semester, Mrs. Rose administers the historical trivia test to her students. The students score an average of 89 on this test. After conducting the appropriate statistical test, Mrs. Rose finds that the p-value is .0025. Which of the following is the best interpretation of the p-value?

a. A p-value of .0025 provides strong evidence that Mrs. Rose's class outperformed high school students across the nation.\\
b. A p-value of .0025 indicates that there is a very small chance that Mrs. Rose's class outperformed high school students across the nation.\\
c. A p-value of .0025 provides evidence that Mrs. Rose is an exceptional teacher who was able to prepare her students well for this national test.\\
d. None of the above.\\

\subsection{Applied Theory-Based Confidence Interval Question}

3. Farmer Cindy is in charge of the chickens on her family's farm, and is curious about the average number of eggs the entire flock produces in a month.  Observing the entire flock would be time consuming, so she approaches you asking what her options are. You inform her that 30 chickens should be selected at random to be observed for a month. After the month she observes the average number of eggs her sample of 30 chickens produced is 25 eggs with a standard deviation of 6 eggs.  
\begin{enumerate} 
\item (5 pts) Construct a 95\% confidence interval for the average number of eggs chickens from her entire population produce in a month.  (You don't need to check any conditions here)  
\item (3 pts) Interpret the 95\% confidence interval constructed in the previous part:
\item (2 pts) Cindy is disappointed with the width of the interval you provide for her, suggest to her two ways she could obtain a narrower confidence interval.
\end{enumerate}

\subsection{Applied Theory-Based Hypothesis Testing Question}

4. Cindy becomes concerned about a disease some of her chickens are catching that causes a decrease in the chicken's egg production.  She is interested in the proportion of her entire flock that has the disease, but detecting the disease requires taking blood from the chicken which is expensive and time consuming.  After working with you in the past, she understands that she can estimate this proportion by taking just a sample of her chickens! Assume she selects a sample of 100 chickens in the best possible way and observed that 15 of the chickens had the disease.

Cindy's pessimistic guess is that 25\% of the flock has the disease.  Complete the following steps to test if the proportion of her entire population diseased is less than 0.25.  
\begin{enumerate} 
\item (2 pts) State the Null and Alternative hypothesis:
\item (2 pts) Check the conditions for a hypothesis test:
\item (2 pts) Calculate the test statistic:
\item (2 pts) Find the p-value:
\item (3 pts) Make a decision about your hypothesis and state your conclusion in context of the problem. 
\end{enumerate}

%-------------------------------------------------------

\section{Appendix: Midterm Exam Used in Curricula Study}
\label{appMidtermExamQuestions}

The following appendix is the midterm exam in its entirety, along with point designations for each question.  Note that Problem 6 for the midterm is the ARTIST scaled question set for assessing topics related to "data collection". This ARTIST set was included with the goal to strengthen the midterm as a effective covariate for controlling for pre-treatment differences in the curricula groups in the model comparing learning outcomes using the ARTIST scores.

%load packages that will be invisible on slides
<<configForMidtermAppendix, echo=FALSE, include=FALSE >>=
### Setting up R ###
#load necessary packages
#library(productplots)
library(graphics)
library(ggplot2)
library(gridExtra)
library(reshape)
library(reshape2)
library(plyr)

#Wellbeing metrics
wbmetric <- "thriving" ; wbperc <- 55

#diamonds data for regression problem
set.seed(12)
index <- sample(1:nrow(diamonds),150)
diamondsamp <- diamonds[index,c(1,7)]
head(diamondsamp)

# smoking and mortality data from DASL 
# http://lib.stat.cmu.edu/DASL/Stories/SmokingandCancer.html
  # smoking = rate of smoking in percentage of average male smoking
  # Mortality = rate of lung cancer mortality in percentage of average male lung cancer mortality
smokedie <- read.delim("smokedie.txt",header=T, )

#Census data for sidebyside boxplot problem
censussamp <- read.csv("CensusDataForExams.csv",header=T)
censussamp$incinthous <- censussamp$inctot/1000
censussamp$gender <- "Female" ; censussamp$gender[censussamp$sex==1] <- "Male"
cenmale <- summary(censussamp[censussamp$gender=="Male", ]$inctot)
cenfemale <- summary(censussamp[censussamp$gender=="Female", ]$inctot)

#marraige role for problem 2
mr2 <- "wife"
mr1 <- "husband"

#iris data for side by side boxplot problem
liliris <- iris[iris$Species!="setosa",c(3,5)] #versionA
#liliris <- iris[iris$Species!="virginica",c(3,5)] #versionB
flower1 <- "Virginica" #VersionA
#flower1 <- "Setosa" #VersionB
flower2 <- "Versicolor"

# Marble Example for Discrete RV Problem
colornum <- c(10,8,3,2)
colorpts <- c(2,5,10,15)
totmarb <- sum(colornum)

# Housing data for Univariate Numerical Variable Problem
load("realestate.rda")
realestate <- realestate[1:nrow(realestate), ]
#Version A
xl <- "Home Price (dollars)" ; yl <- "Count" 
maintit <-"Histogram of Ames Home Prices"  ; homeattr <- "price"
plotattr <- realestate$price ; binw <- 40000

#Version B
#xl <- "Home Floorspace (Square Feet)" ; yl <- "Count" 
#maintit <-"Histogram of Ames Home Floorspaces" ; homeattr <- "floorspace in square feet"
#plotattr <- realestate$sqft ; binw <- 300

#nturns <- 11
nturns <- 12
@

%---------------------------------------------------------------------

\textbf{1. [28 points] Linear Regression:} A study was conducted in England in 1989 investigating the relationship between smoking and lung cancer mortality within different occupational groups of males. The data include a smoking index and a lung cancer mortality index for men in \Sexpr{nrow(smokedie)} occupational groups in England. \\
\indent The smoking index measures the percentage of the number of cigarettes smoked per day by men in the particular occupational group compared to the average number of cigarettes smoked per day by all men. The units associated with the smoking index are percent of average cigarettes smoked (\% avg CS).  The mortality index measures the percentage of the rate of deaths from lung cancer among men in the particular occupational group compared to the rate of deaths from lung cancer among all men. The units associated with the mortality index are percent of average lung cancer deaths (\% avg LCD).  \\
\indent The data for the \Sexpr{nrow(smokedie)} occupations are displayed in the scatterplot below. Use the information from the scatterplot, the sample statistics and linear regression equation listed below to answer all parts of this problem. \\
<< 'regprobplot' , echo=FALSE, warning=FALSE, fig.width=6, fig.height=6, out.width='.5\\linewidth', fig.pos='H',fig.align='center'>>=
mod1 <- lm(Mortality ~ Smoking, smokedie)
#str(summary(mod1))
int <- mod1$coefficients[1]
slo <- mod1$coefficients[2]

qplot(Smoking,Mortality, geom="point", data=smokedie) + geom_abline(intercept = int, slope = slo, size=I(1), colour=I("black"))+ ylab("Mortality (% avg LCD)") + xlab("Smoking (% avg CS)") + ggtitle("Scatterplot of Mortality Index by Smoking Index with Fitted Regression Line") + theme_bw()

ybar1 <- round(mean(smokedie$Mortality),digits=4)
xbar1 <- round(mean(smokedie$Smoking),digits=4)
sy1 <- round(sd(smokedie$Mortality),digits=4)
sx1 <- round(sd(smokedie$Smoking),digits=4)
rsq1 <- round(summary(mod1)$r.squared, 4)-0.0001
r1 <- round(sqrt(rsq1),4)
@
\begin{table}[hbtp]\centering 
\begin{tabular}{ |c|c|c|}
  \hline                        
 $\bar{x}$ = \Sexpr{xbar1}\% avg CS & $\bar{y}$ = \Sexpr{ybar1}\% avg LCD & r=\Sexpr{r1}\\
 $S_x$ = \Sexpr{sx1} \% avg CS & $S_y$ = \Sexpr{sy1}\% avg LCD & $R^2$ = \Sexpr{rsq1}\\
  \hline  
\end{tabular}
\end{table}

\begin{center}
\textbf{Prediction Equation:} \hspace{.2in}  $\hat{y} = \Sexpr{int} + x(\Sexpr{slo})$
\end{center}
\noindent\textbf{(a)} [2 points] What is the response variable?\\
\noindent\textbf{(b)} [2 points] What is the explanatory variable?\\
\noindent\textbf{(c)} [4 points] The correlation coefficient, r=\Sexpr{r1}.  Using this value, what can be said about the relationship between the smoking index and the lung cancer mortality index?\\
\noindent\textbf{(d)} [4 points] The slope of the linear regression model is \Sexpr{slo}. Show the calculations to prove that this is the value for the slope. Then, interpret this value in the context of the problem.\\
\noindent\textbf{(e)} [4 points] The intercept of the linear regression model is \Sexpr{int}. First, show the calculations to prove that this is the value for the intercept. If it is appropriate to interpret this value in the context of the problem, then do so. Otherwise, specify why it is inappropriate to interpret this value.\\
\noindent\textbf{(f)} [3 points] In the prediction equation for the regression line explain that the symbols $\hat{y}$ and x represent.\\
\noindent\textbf{(g)} [3 points] Using the prediction line given above, predict the morality index (in \% avg LCD) for an occupational group that has a smoking index of 110\% avg CS (round your answer to 2 decimal places).\\
\noindent\textbf{(h)} [3 points] Suppose that there was an occupational group in our data set with a smoking index of 110\% avg CS and a mortality index of 140\% avg LCD.  Calculate the residual for this occupational group (round your answer to 2 decimal places).\\
\noindent\textbf{(i)} [3 points] Interpret the $R^2$ value in terms of the context of the problem.\\
%--------------------------------------------------------------------------

\textbf{2. [10 points]} A Gallup poll conducted on February 26, 2014 investigated the wellbeing of American adults. The survey used the Gallup-Healthways Well-Being Index to classify the person's wellbeing as "thriving", "struggling" or "suffering". The telephone survey found that \Sexpr{wbperc}\% of 1500 randomly selected American adults were classified as \Sexpr{wbmetric}.  \\
For this study, identify the following:\\
\noindent\textbf{(a)} [2 points] Identify the population of interest.\\
\noindent\textbf{(b)} [2 points] Identify the sample.\\
\noindent\textbf{(c)} [2 points] Identify the population parameter.\\
\noindent\textbf{(d)} [2 points] Identify the sample statistic.\\
\noindent\textbf{(e)} [2 points] Identify the variable.\\

%--------------------------------------------------------------------------

\textbf{3. [10  points]} A study investigated the relationship between heights of husbands and wives. Heights were recorded as groupings of "Tall", "Medium" and "Short". The researcher recorded height pairings for 205 married couples and wants to know how the height of the husband associated the height of the wife. Use the information found in the contingency table below to answer the questions that follow.
\begin{table}[hbtp]\centering 
\begin{tabular}{ |cc|ccc|c|}
  \hline                        
  & & & Wife & & \\
  & & Tall & Medium & Short & Total \\
 \hline
          & Tall    & 18 & 28  & 14  & 60  \\
Husband   & Medium  & 20 & 51  & 28  & 99  \\
          & Short   & 12 & 25  & 9   &  46 \\
  \hline  
  & Total & 50 & 104 & 51 & 205 \\
  \hline
\end{tabular}
\end{table}
\noindent\textbf{(a)} [2  points] What is the probability that a randomly selected married couple will have a short \Sexpr{mr1}?\\
\noindent\textbf{(b)} [2  points] What is the probability that a randomly selected married couple will have a tall \Sexpr{mr2} \textit{and} a short \Sexpr{mr1}?\\
\noindent\textbf{(c)} [2  points] What is the probability that a randomly selected married couple will have a tall \Sexpr{mr2} \textit{or} a short \Sexpr{mr1}?\\
\noindent\textbf{(d)} [2  points] What is the probability that a randomly selected married couple will have a short \Sexpr{mr1} \textit{given} that we know the \Sexpr{mr2} is tall?\\
\noindent\textbf{(e)} [2  points] Are the events of short \Sexpr{mr1} and tall \Sexpr{mr2} \textit{independent}? Justify your answer mathematically.\\

%--------------------------------------------------------------------------

\textbf{4. [8 points]} Suppose that we have a box containing \Sexpr{totmarb} marbles of equal size and shape so that each marble is equally likely to be selected from the box.  There are \Sexpr{colornum[1]} red, \Sexpr{colornum[2]} green, \Sexpr{colornum[3]} yellow marbles and  \Sexpr{colornum[4]} black marbles.  Suppose we assign points to each color marble: \Sexpr{colorpts[1]} points for a red, \Sexpr{colorpts[2]} point for a green, \Sexpr{colorpts[3]} points for a yellow and \Sexpr{colorpts[4]} points for a black. Let X be the discrete random variable that counts the number of points that a randomly selected marble is worth.\\
\noindent\textbf{(a)} [2 points] Find the probability distribution of X? (fill out table and round all values to 2 decimal places)
\begin{table}[hbtp]\centering 
\begin{tabular}{ c|c }
x & P(x) \\
  \hline  
   &  \\
   &  \\
   &  \\  
\end{tabular}
\end{table}
\noindent\textbf{(b)} [4 points] What 2 properties must be true for the values in the P(x) column of the probability distribution in part 4(a)? Note that X is a discrete random variable.\\
\noindent\textbf{(c)} [2 points] What is the mean of X? (If you were unable to complete part 4(a), describe how you would find the mean) \\

%---------------------------------------------------------------------

\textbf{5. [10 points]} A survey on gender and income was conducted and a random sample of 3700 Adult U.S. residents was gathered. In the figure below are side-by-side boxplots of the income (in thousands of dollars) for male and female U.S. residents. Use this graphical display to answer the following questions.
<< 'sidebysidebox' , echo=FALSE, warning=FALSE, fig.width=9, fig.height=4.5, out.width='.8\\linewidth', fig.pos='H',fig.align='center'>>=
qplot(factor(gender), incinthous, data = censussamp, geom = "boxplot") + coord_flip() + xlab("Gender") + ylab("Income (in thousands of dollars)") + ggtitle("Side by Side Boxplots for Incomes") + theme_bw()
@
\begin{table}[H]\centering 
\begin{tabular}{ |c||c|c||c|c|c|c|c|}
  \hline                        
Gender  & mean & sd & min & Q1 & median & Q3 & max \\
 \hline
 Male & 39319 & 54242 & -10000 & 12000 & 25600 & 46700 & 685000 \\
 Female & 18786 & 21570 & -10000 & 4880 & 13500 & 25800 & 239000 \\
  \hline
\end{tabular}
\end{table}
\noindent\textbf{(a)} [5 points] Using full sentences, describe the distribution of income for \textit{only} the male residents of the U.S.  \\
\noindent\textbf{(b)} [5 points] Using full sentences, compare the distribution of male and female distributions of income. A complete answer will compare shapes, centers and spreads. Be specific about which measures of center and spread are being used to make these comparisons. (ie writing "the spread is bigger" will not earn full credit)  \\

%--------------------------------------------------------------------------

\underline{\textbf{6. [18 points]} Multiple Choice: \textit{Clearly} circle the selected answer.}\\
\noindent\textbf{(i)} [2 points] In a survey people are asked 'Which brand of toothpaste do you prefer?' The data gathered from this question would be what type of data? \\
\indent a. continuous \\
\indent b. categorical \\
\indent c. quantitative

\textbf{Items ii and iii refer to the following situation:}
A student is gathering data on the driving experiences of other college students. One of the variables measured is the type of car the student drives. These data are coded using the following method: 1 = subcompact, 2 = compact, 3 = standard size, 4 = full size, 5 = premium, 6 = mini van, 7 = SUV, and 8 = truck.

\noindent\textbf{(ii)} [2 points] What type of variable is this? \\
\indent a. categorical \\
\indent b. quantitative \\
\indent c. continuous

\noindent\textbf{(iii)} [2 points] The student plans to see if there is a relationship between the number of speeding tickets a student gets in a year and the type of vehicle he or she drives. Identify the response variable in this study. \\
\indent a. college students \\
\indent b. type of car \\
\indent c. number of speeding tickets \\
\indent d. average number of speeding tickets last year

\noindent\textbf{(iv)} [2 points] A researcher is studying the relationship between a vitamin supplement and cholesterol level. What type of study needs to be done in order to establish that the amount of vitamin supplement causes a change in cholesterol level? \\
\indent a. Survey \\
\indent b. Randomized experiment \\
\indent c. Time Series Study \\
\indent d. Survey

\noindent\textbf{(v)} [2 points] An instructor is going to model an experiment in his statistics class by comparing the effect of 4 different treatments on student responses. There are 40 students in the class. Which is the best way for the instructor to distribute the students to the 4 treatments for this experiment? \\ 
\indent a. Assign the first treatment to the first 10 students on the class list, the second treatment to the next 10 students, and so on. \\
\indent b. Assign a unique number to each student, then use random numbers to assign 10 students to the first treatment, 10 students to the second treatment, and so on. \\
\indent c. Assign the treatment as students walk into class, giving the first treatment to the first 10 students and the second treatment to the next 10 student, and so on. \\
\indent d. All of these are equally appropriate methods. \\
\indent e. None of these is an appropriate method.

\textbf{Items vi and vii refer to the following situation:}\\
Suppose two researchers wanted to determine if aspirin reduces the chance of a heart attack.

\noindent\textbf{(vi)} [2 points] Researcher 1 studied the medical records of 500 randomly selected patients. For each patient, he recorded whether the person took aspirin every day and if the person had ever had a heart attack. Then he reported the percentage of heart attacks for the patients who took aspirin every day and for those who did not take aspirin every day. What type of study did Researcher 1 conduct? \\ 
\indent a. Observational \\
\indent b. Experimental \\
\indent c. Survey \\
\indent d. None of the above

\noindent\textbf{(vii)} [2 points] Researcher 2 also studied 500 patients that visited a regional hospital in the last year. He randomly assigned half (250) of the patients to take aspirin every day and the other half to take a placebo everyday. Then after a certain length of time he reported the percentage of heart attacks for the patients who took aspirin every day and for those who did not take aspirin every day. What type of study did Researcher 2 conduct? \\ 
\indent a. Observational \\
\indent b. Experimental \\
\indent c. Survey \\
\indent d. None of the above

\noindent\textbf{(vii)} [2 points] The dean of a college would like to determine the feelings of students concerning a new registration fee that would be used to upgrade the recreational facilities on campus. All registered students would pay the fee each term. Which of the following data collection plans would provide the best representation of students' opinions at the school? \\
\indent a. Survey every 10th student who enters the current recreational facilities between the hours of 1:00 and 5:00 pm until 100 students have been asked. \\
\indent b. Randomly sample fifty student ID numbers and send a survey to all students in the sample. \\
\indent c. Place an ad in the campus newspaper inviting students to complete an online survey. Collect the responses of the first 200 students who respond. \\
\indent d. All of the above would be equally effective.

\noindent\textbf{(ix)} [2 points] A team in the Department of Institutional Review at a large university wanted to study the relationship between completing an internship during college and students' future earning potential. From the same graduating class, they selected a random sample of 80 students who completed an internship and 100 students who did not complete an internship and examined their salaries 5 years past graduation. They found that there was a statistically higher mean salary for the internship group than for the non-internship group. Which of the following interpretations do you think is the most appropriate?\\
\indent a. More students should take internships because having an internship produces a higher salary.\\
\indent b. There could be a confounding variable, such as student major, that explains the difference in mean salary between the internship and no internship groups. \\
\indent c. You cannot draw any valid conclusions because the samples are not the same size.\\

%--------------------------------------------------------------------------

\textbf{7. [8 points]} When playing the game Settlers of Catan\textbf{TM} there is a game piece known as the robber that blocks resources from being obtained.  On each player's turn, that player rolls a pair of 6 sided dice.  If the sum of the dice roll equals 7 then that player may move the robber. This means that there is a 6/36 chance of moving the robber with every roll of the dice, because there are 6 of the 36 possible combinations that sum to 7. Let X count the number of times that one player moves the robber in \Sexpr{nturns} turns.\\
\noindent\textbf{(a)} [4 points]  Check the 4 conditions to argue that X is a Binomial Random Variable  \\
\noindent\textbf{(b)} [2 points]  What is the probability that a player gets to move the robber 2 times during the \Sexpr{nturns} turns?  \\
\noindent\textbf{(c)} [2 points]  What is the mean number of times during \Sexpr{nturns} turns that a player will be able to move the robber?\\

%-----------------------------------------------------------------------

\textbf{8. [4 points]} A waiter from a restaurant keeps track of the information about tips he receives from dinning parties. Part of this data includes the tipping rate, as a proportion of the total bill, that was tipped. Below is a histogram of tipping rates from the 244 dinning parties served by the waiter. Use this histogram to answer the following question.
<< 'tipratehisto' , echo=FALSE, warning=FALSE, fig.width=9, fig.height=3.5, out.width='.7\\linewidth', fig.pos='H',fig.align='center'>>=
qplot(tip/total_bill, data=tips, binwidth=.04, xlim=c(0,.8)) + 
  xlab("Tipping Rate (proportion of total bill)") + ylab("Count") + ggtitle("Histogram of Tipping Rates") + theme_bw()
@
\noindent\textbf{(a)} [2 points]  What is the best measure of center to describe the distribution of tipping rates? Explain your answer. (Note: no calculations or estimated values are needed to answer this problem)  \\
\noindent\textbf{(b)} [2 points]  What is the best measure of spread to describe the distribution of tipping rates? Explain your answer. (Note: no calculations or estimated values are needed to answer this problem) \\

%-----------------------------------------------------------------------

\textbf{9. [4 points]} Below are histograms for two separate random samples of 30 quiz scores.  Which quiz will have the larger sample standard deviation for quiz scores? \textbf{Explain your choice briefly}. (Note: You do not, and should not, do any calculations to answer this question)
\begin{figure}[H]
\begin{center}
\includegraphics[keepaspectratio=TRUE,width=.2\textwidth]{CurriculumStudy/QuizScores1Version1.png}
\includegraphics[keepaspectratio=TRUE,width=.2\textwidth]{CurriculumStudy/QuizScores2Version1.png}
\end{center}
\end{figure}

%-------------------------------------------------------
\section{Appendix: MANCOVA Model Diagnostics}
\label{appModDiag}

For the ARTIST Model described in Subsection~\ref{ArtistModel} and the Applied Model described in Subsection~\ref{AppliedModel} a set of model diagnostics were conducted. Each of these MANCOVA models are parameterized as specified in Section~\ref{analysis}; as such the assessment of model assumptions will be very similar for each model.   Residual plots will be used to assess the assumptions of linearity and constant variance.  The univariate and bivariate normality of the error terms will be assessed visually using normal quantile plots and a scatterplot of the paired residuals from the model. Additionally, although it is not a modeling assumption, MANCOVA models are best behaved with low to moderate correlation between the response variables, because then then model is able to capture variance unique to each response.

% \subsection{Simulation Study for Violations of Independence}
% \label{simappend}
% 
% \km{To be removed after new simulation section added to main body.}\\
% 
% The ARTIST Model of the form (\ref{eq:mancova}) assumes  that errors from different students are independent, i.e.~$\Cov[\epsilon_{ik}, \epsilon_{jl}] = 0$, for all $i \neq j$ with $ 1 \le i,j \le n$. In our simulation setting, we simulate data from a model adapted from the model~(\ref{eq:mancova}), where there is no curriculum effect, i.e.~$\tau_{1} =  \tau_{2} = 0$ and where additionally the assumption of independent errors is violated by assuming a between student covariance of 
% \[
% \Cov[\epsilon_{ik}, \epsilon_{jl}] = \eta \sigma_{kl}^2, \ \ \ \text{ where } \eta \in [0,1]  \text{ for all } i \ne j. 
% \]
% $\eta$ scales the covariance between all students uniformly and can be adjusted to increase the violation of independence. Note that when $\eta = 0$ we have no violation of independence. Thus our generative model for the ARTIST responses are linear combinations of the midterm and lab~5 scores with correlated errors. This generative model for simulation can be organized using vector notation as follows:
% %
% \begin{eqnarray}\label{eq:linear}
% \begin{bmatrix}
%   \vec{y}_1 \\ \vec{y}_2
%   \end{bmatrix}  = 
% % %---- 
% \begin{bmatrix}
%  y_{11} \\ \vdots \\ y_{n1} \\ y_{12} \\ \vdots \\ y_{n2} 
% \end{bmatrix}  =
% % %----  
%  \begin{bmatrix}
%   \beta_{01} + x_{11}\beta_{11} + x_{12}\beta_{21}  \\
%   \vdots  \\ 
%   \beta_{01} + x_{n1}\beta_{11} + x_{n2}\beta_{21}  \\
%   \beta_{02} + x_{11}\beta_{12} + x_{12}\beta_{22}  \\
%   \vdots  \\ 
%   \beta_{02} + x_{n1}\beta_{12} + x_{n2}\beta_{22}  \\
%  \end{bmatrix} +
%  \begin{bmatrix}
%  \epsilon_{11} \\ \vdots \\ \epsilon_{n1} \\ \epsilon_{12} \\ \vdots \\ \epsilon_{n2} 
%  \end{bmatrix} = 
%  % %----
%  \begin{bmatrix}
%   X\vec{\beta}_1 \\ X\vec{\beta}_2
%   \end{bmatrix} + 
%   \vec{\epsilon},
% \end{eqnarray} 
% %
% \begin{eqnarray}\label{eq:error}
% \vec{\epsilon} \sim
% % %----  
%  \text{MVN}\left(
% \vec{0} \hspace{.1cm}
%   , 
%   \begin{bmatrix}
%    \sigma_{11}^2 & \sigma_{12}^2 \\ 
%    \sigma_{21}^2 & \sigma_{22}^2
%   \end{bmatrix}
% \otimes
%  \left( \eta 1_{n \times n} + (1- \eta) I_{n \times n}\right)
% %  \left[\begin{array}{ccccc|ccccc}
% %   \sigma_{11}^2 & \eta \sigma_{11}^2 & \cdots & \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2\\
% %   \eta \sigma_{11}^2 & \sigma_{11}^2 & \ddots & \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \eta \sigma_{12}^2 & \sigma_{12}^2 & \ddots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2\\
% % \vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
% % \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \ddots & \sigma_{11}^2 & \eta \sigma_{11}^2 & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \ddots & \sigma_{12}^2 & \eta \sigma_{12}^2\\
% % \eta \sigma_{11}^2 & \eta \sigma_{11}^2 & \cdots & \eta \sigma_{11}^2 & \sigma_{11}^2 & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \sigma_{12}^2\\
% % \hline
% % \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \sigma_{22}^2 & \eta \sigma_{22}^2 & \cdots & \eta \sigma_{22}^2 & \eta \sigma_{22}^2\\
% %   \eta \sigma_{12}^2 & \sigma_{12}^2 & \ddots & \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \eta \sigma_{22}^2 & \sigma_{22}^2 & \ddots & \eta \sigma_{22}^2 & \eta \sigma_{22}^2\\
% % \vdots & \ddots & \ddots &\ddots &\vdots & \vdots & \ddots & \ddots &\ddots &\vdots \\
% % \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \ddots & \sigma_{12}^2 & \eta \sigma_{12}^2 & \eta \sigma_{22}^2 & \eta \sigma_{22}^2 & \ddots & \sigma_{22}^2 & \eta \sigma_{22}^2\\
% % \eta \sigma_{12}^2 & \eta \sigma_{12}^2 & \cdots & \eta \sigma_{12}^2 & \sigma_{12}^2 & \eta \sigma_{22}^2 & \eta \sigma_{22}^2 & \cdots & \eta \sigma_{22}^2 & \sigma_{22}^2\\  
% %  \end{array}\right]
%  \right),
% \end{eqnarray}
% where $A \otimes B = \left( a_{ij} B \right)$ is the Kronecker product of matrices $A$ and $B$. $\vec{0}$ is the zero vector of length $2n$, $1_{n\times n}$ is the $n \times n$ matrix consisting of 1s only, and $I_{n \times n}$ is the identity matrix of dimension $n \times n$.
% Here, $x_{i1}$ is the midterm exam score for student $i$ and $x_{i2}$ is the lab~5 score for student $i$. In order to achieve practically relevant values for the coefficients $\beta_{kp}$ of the generative model we selected the estimated values from a MANCOVA model fit to the ARTIST responses using only the midterm exam and lab~5 scores.  Likewise, the variance components, $\sigma_{kl}$ for the generative model were selected using the estimated variance structure from the same reduced MANCOVA model for the ARTIST responses. 
% 
% The simulation procedure for assessing the error rates under violation independence between student responses was conducted in a five step process.
% For each $\eta$ in $\{0,0.1,0.2,0.3,0.4,0.5\}$ repeat the following process for $M = 25,000$ simulations each:
% %
% \begin{enumerate}
% \item Simulate an error vector, $\vec{\epsilon}^{\hspace{.1cm}(m)}$, from the multivariate normal distribution as defined in~(\ref{eq:error}).
% \item Derive corresponding responses based on model (\ref{eq:linear}) as:
% %
% \[
% \begin{bmatrix}
%   \vec{y}_1^{\hspace{.1cm}(m)} \\ \vec{y}_2^{\hspace{.1cm}(m)}
%   \end{bmatrix}   =
%    % %----
%  \begin{bmatrix}
%   X\vec{\beta}_1 \\ X\vec{\beta}_2
%   \end{bmatrix} + 
%   \vec{\epsilon}^{\hspace{.1cm}(m)}
% \]
% %
% \item Create assignment to curricula  for each student randomly. 
% 
% \item Fit the ARTIST Model (assuming independence) as given in model (\ref{eq:mancova}) to the simulated responses
% \item Record $\hat{\beta}_{kp}^{\hspace{.1cm}(m)}$, $\text{SE}\left[\hat{\beta}_{kp}^{\hspace{.1cm}(m)}\right]$, $\tau_k^{\hspace{.1cm}(m)}$, $\text{SE}\left[\tau_k^{\hspace{.1cm}(m)}\right]$ and $\hat{\sigma}_{kl}^{\hspace{.1cm}(m)}$ $\forall k,l \in \{1,2\}$ and $p \in \{0,1,2\}$. \\
% Also record Pillai's $\lambda^{\hspace{.1cm}(m)}$ testing for overall curriculum effect.
% %\item Repeat simulation steps 1 through 6, $M=25000$ times
% 
% \end{enumerate}
% 
% <<results, echo=FALSE, include=FALSE>>=
% load("SimulationResults.RData")
% library(plyr)
% eta <- (0:5)/10
% results <- ldply(1:6, function(x) data.frame(eta=eta[x], resultlist[[x]]))
% # beta1s and beta2s are actually tau1 and tau2
% simresults = ddply(results, .(eta), summarise,
%       trt1=mean(beta1s),
%       trt2=mean(beta2s),
%       type1.1=mean(pval1s<0.05),
%       type1.2=mean(pval2s<0.05),      
%       betamidterm1s = mean(betamidterm1s),
%       pvalmidterms1s = mean(pvalmidterm1s),
%       sebetamidterm1s = mean(SEbetamidterm1s),
%       betamidterm2s = mean(betamidterm2s),
%       pvalmidterms2s = mean(pvalmidterm2s),
%       sebetamidterm2s = mean(SEbetamidterm2s),
%       avgpillai = mean(pillais),
%       avgpillaipval = mean(pillaipvals),
%       sebeta1s = mean(SEbeta1s),
%       sebeta2s = mean(SEbeta1s),
%       sdbeta1 = sd(beta1s),
%       sdbeta2 = sd(beta2s)
% )
% 
% 
% L <- length(resultlist)
% M = nrow(resultlist[[1]]) 
% errorsummary <- data.frame(T2ErrorRateBetaMidterm1 = rep(NA, L),
%                            T2ErrorRateBetaMidterm2 = rep(NA, L),
%                            T2ErrorRateBetaLab1 = rep(NA, L),
%                            T2ErrorRateBetaLab2 = rep(NA, L),
%                            T1ErrorRateBeta1 = rep(NA, L),
%                            T1ErrorRateBeta2 = rep(NA, L),
%                            T1ErrorRatePillai = rep(NA, L))
% 
% for (l in 1:L){
%   errorsummary[l,1] <- 1- length(which(resultlist[[l]]$pvalmidterm1s <= 0.05)) / M
%   errorsummary[l,2] <- 1- length(which(resultlist[[l]]$pvalmidterm2s <= 0.05)) / M
%   errorsummary[l,3] <- 1- length(which(resultlist[[l]]$pvallab1s <= 0.05)) / M
%   errorsummary[l,4] <- 1- length(which(resultlist[[l]]$pvallab2s <= 0.05)) / M
%   errorsummary[l,5] <- length(which(resultlist[[l]]$pval1s <= 0.05)) / M
%   errorsummary[l,6] <- length(which(resultlist[[l]]$pval2s <= 0.05)) / M
%   errorsummary[l,7] <- length(which(resultlist[[l]]$pillaipvals <= 0.05)) / M
% }
% errorsummary$eta <- eta
% #errorsummary
% @
% 
% Histograms for the estimated parameters from simulations at each value of $\eta$ are displayed in Figure~\ref{fig:simhistogrid}.  Note that the histograms are each centered around the true values of the parameters from the generative model, indicating that the estimates remain unbiased as the violation of independence between student responses increases. The figure also displays a decreasing variability in the parameter estimates as $\eta$ increases. The standard deviation for the treatment effect estimates, $s_{\tau_k}$, can be found in table~\ref{tab:simvarcompare}; along with averages from standard errors for the treatment effects, $\hat{\text{SE}}\left[\tau_k\right]$.  Note that while the former is measuring the variability of the coefficients from all simulations, the latter is the average of variability estimates from each individual simulation.  The paired values are nearly equivalent, meaning that the standard errors used in the t-tests for treatment effects shrink at the same rate as the true sampling variability in the parameters estimates as the violation of independence increases. 
% 
% <<simhistogrid, echo=FALSE , message=FALSE, warning=FALSE, fig.width=10, fig.height=6, out.width='1\\linewidth', fig.pos='h',fig.align='center',fig.cap="ARTIST model parameter estimates from 25,000 simulations at each $\\eta$. True parameter values from the generative model are displayed as vertical lines in each cell.", cache=TRUE>>=
% library(ggplot2)
% library(reshape2)
% etas <- seq(from=0, to=0.5, by=.1) 
% plotbetadata <- data.frame(eta = rep(as.character(etas),each=M))
% plotbetadata <- cbind(plotbetadata, rbind(resultlist[[1]][,c(1,2,7,8,13,14)],
%                                           resultlist[[2]][,c(1,2,7,8,13,14)],
%                                           resultlist[[3]][,c(1,2,7,8,13,14)],
%                                           resultlist[[4]][,c(1,2,7,8,13,14)],
%                                           resultlist[[5]][,c(1,2,7,8,13,14)],
%                                           resultlist[[6]][,c(1,2,7,8,13,14)]) )
% names(plotbetadata) <- c("eta","beta[11]","beta[12]","beta[21]","beta[22]","tau[1]","tau[2]")
% plotbetadata$eta <- paste("eta==",plotbetadata$eta,sep="")
% meltplotbetadata <- melt(plotbetadata, id=c("eta"))
% trueeffects <- unique(meltplotbetadata[,1:2])
% trueeffects$value <-  rep(c(0.046132,0.037915,0.017485, 0.008211,0,0),each=6)
% number_ticks <- function(n) {function(limits) pretty(limits, n)}
% qplot(value, geom="histogram", data=meltplotbetadata,fill=I("gray42")) + 
%   facet_grid(eta~variable, scales = "free",labeller = label_parsed) +
%   geom_vline(aes(xintercept = value), data=trueeffects, color="black")+ theme_bw()+
%   ylab("") + xlab("Estimated Parameter Values") +
%   scale_x_continuous(breaks=number_ticks(3))
% @
% 
% \begin{table}[hbtp]
% \centering
% \begin{tabular}{ccccc} \hline
% $\eta$ & $s_{\tau_1}$ & $\hat{\text{SE}}\left[\tau_1\right]$ & $s_{\tau_2}$ & $\hat{\text{SE}}\left[\tau_2\right]$\\ 
%  \hline  
% \Sexpr{sprintf("%.1f",round(simresults[1,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[1,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[1,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[1,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[1,17],4))} \\
% \Sexpr{sprintf("%.1f",round(simresults[2,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[2,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[2,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[2,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[2,17],4))} \\
% \Sexpr{sprintf("%.1f",round(simresults[3,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[3,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[3,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[3,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[3,17],4))} \\
% \Sexpr{sprintf("%.1f",round(simresults[4,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[4,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[4,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[4,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[4,17],4))} \\
% \Sexpr{sprintf("%.1f",round(simresults[5,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[5,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[5,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[5,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[5,17],4))} \\
% \Sexpr{sprintf("%.1f",round(simresults[6,1],1))} & \Sexpr{sprintf("%.4f",round(simresults[6,14],4))} & \Sexpr{sprintf("%.4f",round(simresults[6,16],4))} &\Sexpr{sprintf("%.4f",round(simresults[6,15],4))}  &\Sexpr{sprintf("%.4f",round(simresults[6,17],4))} \\
% \hline
% \end{tabular}
% \caption{Comparison of the standard deviation for the treatment effect estimates, $s_{\tau_k}$, and average from standard errors for the treatment effects, $\hat{\text{SE}}\left[\tau_k\right]$ from simulations under each $\eta$.}
% \label{tab:simvarcompare}
% \end{table}
% %
% Table~\ref{tab:errorrates} displays that the Type I error rates from the Pillai's test and each of the individual t-tests for curriculum treatment effects are stable around the nominal $\alpha = 0.05$ level for all $\eta$ values. Thus the violation of independence between student responses is not detrimental to the Type I error rates in this setting, where no treatment effects exists in our generative model.  The stability in the Type I errors is linked to the finding that the variability estimates used in testing are shrinking at approximately the same rate as the true sampling variability in the parameters estimates, as discussed above.
% 
% Figure~\ref{fig:visual1} displays that the treatment effect estimates remain unbiased and the Type I error rates for testing the significance of the treatment on each response remain at the nominal level. We may additionally be curious about the Type II error rates; failing to find effects significant that truly exist in our generative model.  Figure~\ref{fig:visual2} displays that the midterm effect estimates are unbiased for the non-zero parameters in our generative model.  It also shows that the Type II error rate in fact \textit{decreases} as the dependence between student responses increases.  This implies that power for finding a truly significant effect \textit{increased} as the violation of independence increased.  
% 
% Our simulations, under this generative model, have demonstrated that parameter estimates remain unbiased and error rates are not detrimentally effected by increasing dependence between student responses.  Thus concerns over violating the modeling assumption of independence between student responses are alleviated. 
% 
% \begin{table}[hbtp]
% \centering
% \begin{tabular}{cccc} \hline
% $\eta$  & Pillai's Test & t-test (CI) & t-test (HT) \\
% & $\text{H}_o : \tau_1 = \tau_2 = 0$  & $\text{H}_o : \tau_1 = 0$ &  $\text{H}_o : \tau_2 = 0$\\ 
%  \hline  
% \Sexpr{sprintf("%.1f",round(errorsummary[1,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[1,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[1,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[1,6],4))} \\
% \Sexpr{sprintf("%.1f",round(errorsummary[2,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[2,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[2,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[2,6],4))} \\
% \Sexpr{sprintf("%.1f",round(errorsummary[3,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[3,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[3,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[3,6],4))} \\
% \Sexpr{sprintf("%.1f",round(errorsummary[4,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[4,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[4,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[4,6],4))} \\
% \Sexpr{sprintf("%.1f",round(errorsummary[5,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[5,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[5,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[5,6],4))} \\
% \Sexpr{sprintf("%.1f",round(errorsummary[6,8],1))} & \Sexpr{sprintf("%.4f",round(errorsummary[6,7],4))} & \Sexpr{sprintf("%.4f",round(errorsummary[6,5],4))} &\Sexpr{sprintf("%.4f",round(errorsummary[6,6],4))} \\
% \hline
% \end{tabular}
% \caption{Type I error rates for tests for curriculum treatment effect from 25,000 simulations at each $\eta$. All tests were conducted with a nominal significance level of $\alpha = 0.05$.}
% \label{tab:errorrates}
% \end{table}
% 
% 
% <<visual1, echo=FALSE , message=FALSE, warning=FALSE, fig.width=11, fig.height=3, out.width='.9\\linewidth', fig.pos='h',fig.align='center',fig.cap="Average curriculum treatment effects on each value in the simulated response pairs (left) and corresponding Type I error rates (right) from 25,000 simulations at each $\\eta$. The generative model set $\\tau_1 = \\tau_2 = 0$ and tests for significance were run using $\\alpha = 0.05$">>=
% library(ggplot2)
% library(reshape2)
% library(gridExtra)
% srm <- melt(simresults, id.var="eta", measure.var=c("trt1", "trt2"))
% levels(srm$variable) <- c("ConfMC","HypMC")
% p1 <- ggplot(aes(x=eta, y=value, group=variable), data=srm) + 
%   geom_hline(yintercept=0.00, colour="grey50") + theme_bw() + 
%   geom_point(size=4, aes(shape=variable)) + geom_line(aes(ltype=variable)) + ylim(c(-0.01,0.01)) + 
%   ylab("Estimated Treatment Effect") + xlab(expression(eta)) +
%   theme(legend.position="none")
% 
% srm <- melt(errorsummary, id.var="eta", measure.var=c("T1ErrorRateBeta1", "T1ErrorRateBeta2"))
% levels(srm$variable) <- c("ConfMC","HypMC")
% p2 <- ggplot(aes(x=eta, y=value, group=variable), data=srm) + 
%   geom_hline(yintercept=0.05, colour="grey50") + theme_bw() + 
%   geom_point(size=4, aes(shape=variable)) + geom_line(aes(ltype=variable)) + ylim(c(0,0.06)) + 
%   ylab("Type I Error Rate") + xlab(expression(eta)) +
%   scale_shape_discrete(name="Treatment Effect", labels=c(expression(paste("  ",tau[1])),expression(tau[2]))) +
%   theme(legend.text = element_text(size = 12))
% 
% grid.arrange(p1,p2,nrow=1, widths=c(1,1.4))
% @
% 
% <<visual2, echo=FALSE , message=FALSE, warning=FALSE, fig.width=11, fig.height=3, out.width='.9\\linewidth', fig.pos='h',fig.align='center',fig.cap="Average midterm exam effects on each value in the simulated response pairs (left) and corresponding Type II error rates (right) from 25,000 simulations at each $\\eta$. The generative model set $\\beta_{11} = 0.0461$ and $\\beta_{12} = 0.0379 $ and tests for significance were run using $\\alpha = 0.05$.">>=
% srm <- melt(simresults, id.var="eta", measure.var=c("betamidterm1s", "betamidterm2s"))
% levels(srm$variable) <- c("ConfMC","HypMC")
% p1 <- ggplot(aes(x=eta, y=value, group=variable), data=srm) +
%   #geom_hline(yintercept=c(0.046132,0.037915), colour="grey50")+
%   theme_bw() + 
%   geom_point(size=4, aes(shape=variable)) + geom_line(aes(ltype=variable)) +
%   ylab("Estimated Midterm Effect") + xlab(expression(eta)) + ylim(c(0, 0.05))+
%   theme(legend.position="none")
% 
% 
% srm <- melt(errorsummary, id.var="eta", measure.var=c("T2ErrorRateBetaMidterm1", "T2ErrorRateBetaMidterm2"))
% levels(srm$variable) <- c("ConfMC","HypMC")
% p2 <- ggplot(aes(x=eta, y=value, group=variable), data=srm) + 
%   theme_bw() + 
%   geom_point(size=4, aes(shape=variable)) + geom_line(aes(ltype=variable)) +
%   ylab("Type II Error Rate") + xlab(expression(eta)) +
%   scale_shape_discrete(name="Midterm Coefficients", labels=c(expression(paste("  ",beta[11])),expression(beta[12]))) +
%   theme(legend.text = element_text(size = 12))
% grid.arrange(p1,p2,nrow=1, widths=c(1,1.4))
% @
% 
% 
% A final consideration is that in practice we might be dealing with a situation where there is a more complex covariance structure, where students of the same classrooms have more related scores than students of different classrooms. The rational is that students within the same physical environment may interact in ways that cause their responses to be more similar than students of different classrooms. In this situation we adjust the generative model (\ref{eq:linear}) to assume the covariance between student to be
% \[
% \Cov[\epsilon_{ik}, \epsilon_{jl}] = c \eta \sigma_{kl}^2, 
% \]
% where $c = 1$ for all students $i,j$ in same classroom, and $c \in [0,1] $ for students $i,j$ in different classrooms.  In this way, $c$ scales the covariance back for all students not sharing a classroom. We will refer to $c$ as the ``non-classmate covariance scaling parameter".  
% %We will bear in mind that this difference should not reflect a treatment effect, merely a covariance structure for items on the final exam due to the sharing of physical environments (i.e. inter-student conversation, physical differences in classroom).  We do not want $c$ to represent the differences caused by the administration of a curricula because this is considered a treatment effect, which is not supposed to exist in the generative model. We will simulate with $c$ rather close to 1 in the generative model.
% The simulation procedure otherwise follows the same five stage procedure described previously with $M = 10,000$ simulations at each combination of $\eta$ in $\{0,0.1,0.2,0.3,0.4\}$ and $c$ in $\{0.9, 0.925, 0.95, 0.975, 1.0\}$.
% 
% Similar to the initial simulation, it was found that the parameter estimates remain unbiased as the violation of independence between student responses increases in this adapted generative model. There was also decreasing variability in the parameter estimates as $\eta$ and $c$ increase.
% 
% Figure~\ref{fig:sim2plots} displays the Type I error rates for testing for treatment effects univariately using t-tests and bivariately using Pillai's overall test. What is clear is that the type I error rate climbs quickly when the non-classmate covariance scaling parameter decreases. This means that when the variance structure for observations more clearly aligns for students in the same classroom, it becomes more likely to observe a difference in the scores between classrooms.  
% 
% Unfortunately, there is no way to estimate $c$ from our data, where classroom is entirely confounded with curricula treatment.  As such we will need to recognize that reliability of our results are contingent on the belief that the alternation of instruction, identical curricula administered with all students in the same room in weeks 1 to 8, and careful pedagogical preparation has minimized all non-treatment related differences that may have arisen due to the physically different classrooms in which the treatments were administered.
% 
% <<ResultsClassroomSimulation, echo=FALSE, include=FALSE>>=
% errorsummary2 <- read.csv("C:\\Users\\Karsten\\Dropbox\\Dissertation\\CurriculumStudy\\SimulationStudy/Sim1000ClassCovResults.csv")
% @
% 
% <<sim2plots, echo=FALSE , message=FALSE, warning=FALSE, fig.width=8, fig.height=3, out.width='.8\\linewidth', fig.pos='h',fig.align='center',fig.cap="Type I error rates for tests of treatment effects from 10,000 simulations at each combination of $\\eta \\in \\{0,0.1,0.2,0.3,0.4\\}$ and $c \\in \\{0.9,0.925,0.95,0.975,1.0\\}$. The nominal $\\alpha=0.05$ is displayed with the dashed line in each plot.", cache=TRUE>>=
% library(ggplot2)
% meltsim2 <- melt(errorsummary2, id=c("eta", "nonclassmate"))
% meltsim2$variable2 <- factor(meltsim2$variable, labels = c("H[o]: tau[1]==0", "H[o]: tau[2]==0", "Pillai"))
% qplot(eta, value, geom="path", data=meltsim2, colour=as.factor(nonclassmate), group=nonclassmate, size=I(1.2)) +
%   geom_point(size=I(1.6),colour=("Black"))+
%    facet_grid(. ~ variable2, labeller = label_parsed) +
%   xlab(expression(eta)) + ylab("Type I Error Rate") + 
%   theme_bw() + labs(colour='Non-classmate Variance
% Scaling Parameter (c)') + geom_hline(yintercept=0.05, linetype=2)+
%    geom_hline(yintercept=0, linetype=1)+
%   ylim(c(0,.50))+
%   scale_colour_manual(values = c("gray80","gray70","gray60","gray60","gray40"))
% @
%

\subsection{ARTIST Model Diagnostics}
\label{appARTISTModDiag}

The ARTIST Model is assessed to satisfactorily meet the modeling assumptions. The correlation between the ARTIST scores for confidence intervals and hypothesis tests is acceptable for modeling with MANCOVA with a correlation of $\Sexpr{round(with(dat, cor(ConfMC,HypMC)),3)}$ between the responses.  The normality of the errors is upheld by the plots in Figure~\ref{fig:ARTISTModDiag}.  The normal quantile plots only display a slight bend for the residuals from the Hypothesis Test topic scores, and the bivariate distribution of the residuals are visually consistent with bivariate normality.

<<ARTISTModSelect,echo=F,include=F,eval=T>>=
### Build MANCOVA model to be used elseware in appendix
mod2 <- lm(cbind(ConfMC,HypMC)~ midterm + Section +
             hw1perc +  hw2perc + hw3perc + hw4perc + hw5perc + hw6perc + hw7perc +
            lab1perc +lab2perc +lab3perc +lab4perc +lab5perc +lab6perc + lab7perc +  room , data=dat)
mod2backward <- mStep(mod2, k=2, trace=TRUE) #k=2 means to use AIC
mod2small <- update(mod2backward, .~. -  lab2perc - hw4perc - hw5perc)
summarytab <- summary(manova(mod2small))$stats
mod2smallCI <- lm(ConfMC~ midterm + lab5perc + room, data=dat)
CIcis <- data.frame(confint(mod2smallCI,level=.95))
CIcis$ests <- mod2smallCI$coeff
mod2smallHT <- lm(HypMC~ midterm + lab5perc + room, data=dat)
HTcis <- data.frame(confint(mod2smallHT,level=.95))
HTcis$ests <- mod2smallHT$coeff
@

<<ARTISTModDiag, echo=FALSE , warning=FALSE, fig.width=5, fig.height=5, out.width='.49\\linewidth', fig.pos='h',fig.align='center',fig.cap="Normal quantile plots (left) and bivariate scatterplot (right) for residuals of each response from ARTIST Model.",fig.show='hold'>>=
#--------------------------------------------------
### Check Conditions for MANCOVA linear model fit
## check response correlations to avoid colinearity in multivariate model
#Rule of Thumb: ~ .3 to .55 then MANCOVA will work well
#with(dat, cor(ConfMC,HypMC)) #Not overly correlated

### check for influencial and high leverage points
#infmeas <- influence.measures(mod2small)
#head(infmeas$infmat)
#head(infmeas$is.inf)
### no high leverage

#Normality of Residuals
#op <- par(mfrow = c(1, 3))      
#qqp1 <- qqPlot(mod2small$residuals[,1], xlab="Theoretical Normal Quantiles",
#       ylab="ARTIST CI Score Residual")

#qqp2 <- qqPlot(mod2small$residuals[,2], xlab="Theoretical Normal Quantiles",
#       ylab="ARTIST HT Score Residual")
#op

qqp1 <- qplot(sample=mod2small$residuals[,1], stat="qq", xlab="Theoretical Normal Quantiles",
       ylab="ARTIST CI Score Residual") +  theme_bw()

qqp2 <- qplot(sample=mod2small$residuals[,2], stat="qq", xlab="Theoretical Normal Quantiles",
      ylab="ARTIST HT Score Residual") +  theme_bw()

grid.arrange(qqp1,qqp2,nrow=2)

qplot(mod2small$residuals[,1],mod2small$residuals[,2],
      xlab="ARTIST CI Score Residual",  ylab="ARTIST HT Score Residual")  +  theme_bw()
# Univariate and Bivariate normality looks fine

# makeenv <- function(res){
# grav.qqboot <- boot(res,sort,R=999,sim="parametric",ran.gen=grav.gen)
# theo.qq <- qqnorm(res,plot=FALSE)
# theo.qq <- lapply(theo.qq,sort)
# env <- envelope(grav.qqboot,level=0.95)
# envdat <- data.frame(theoquant = c(theo.qq$x,theo.qq$x),
#                      envelope = c(env$point[1,],env$point[2,]),
#                      bound = rep(c("upper","lower"),each=length(res)))
# return(envdat)
# }
# 
# envdat <- makeenv(mod2small$residuals[,1])
# head(envdat)
# qplot(sample = mod2small$residuals[,1], stat = "qq") +
#   geom_path(aes(x=theoquant, y=envelope), data=subset(envdat,bound=="lower")) +
#   geom_path(aes(x=theoquant, y=envelope), data=subset(envdat,bound=="upper"))

## Independence of responses between students must be assumed
@

The residual plots in Figure~\ref{fig:residualsARTIST} display stripped bands of points that run shallowly downward, creating the optical illusion of a trend in the points.  This is because the model fits the discretely recorded ARTIST scores as continuous response variables, thus there is a discrete set of residuals possible for any particular fitted value.  Note that, as with all least-squares regression models, the residuals are uncorrelated with the fitted values.

The residual plots in Figure~\ref{fig:residualsARTIST} are overlain with Loess smoothers -- and corresponding 95\% confidence envelopes -- to check for violations of linearity.  There is no issue with the assumption of linearity in the prediction of the confidence interval score, but there is a slight significant dip in the pattern for hypothesis interval residuals.  The assumption of homoscedasticity appear to hold with the residuals spread fairly evenly at all levels of the fitted values. There are however a few outliers on the residual plots, specifically a few students who scored abnormally lower than predicted. These outliers were investigated and found to be low leverage and non-influential.

<<residualsARTIST, echo=FALSE , message=FALSE, warning=FALSE, fig.width=8, fig.height=3.2, out.width='.9\\linewidth', fig.pos='h',fig.align='center',fig.cap="ARTIST Model residual plots overlaid with Loess smoother and corresponding 95\\% confidence envelopes.">>=
## check residual plots for homoscedesticity and linearity
jitheight <- .1
p1 <- qplot(mod2small$fitted.values[,1] ,mod2small$residuals[,1], geom = "jitter") + 
  geom_hline(yintercept=0) +geom_jitter(position = position_jitter(height = jitheight )) +
  xlab("ARTIST CI Score Fitted Value") + ylab("ARTIST CI Score Residual") +  theme_bw() + geom_smooth()
p2 <- qplot(mod2small$fitted.values[,2] ,mod2small$residuals[,2]) + 
  geom_hline(yintercept=0) +geom_jitter(position = position_jitter(height = jitheight )) +
  xlab("ARTIST HT Score Fitted Value") + ylab("ARTIST HT Score Residual") +  theme_bw() + geom_smooth()
grid.arrange(p1,p2,nrow=1)
@

\subsection{Applied Model Diagnostics}
\label{appAppliedModDiag}

A correlation of $\Sexpr{round(with(dat, cor(AppliedCI,AppliedHT)),3)}$ between the pair of applied problem scores for confidence intervals and hypothesis tests is acceptable for modeling with MANCOVA.  The assumption of normality of errors in the Applied Model is potentially problematic. The normal quantile plot for the residuals from the applied confidence interval score in Figure~\ref{fig:AppliedModDiag} show a distinct curve. The scatterplot of the residual pairs from the Applied Model also appear to have a non-normal bivariate distribution.

The Loess smoothers do not significantly departing from a the horizontal lines at zero for the residual plots in Figure~\ref{fig:residuals2} and thus no departures from the assumption of linearity.  The residuals do however show signs of changing variance over the range of the fitted values. This appears to be driven by the upper bound on student scores for each question.

<<AppliedModDiag, echo=FALSE , warning=FALSE, fig.width=5, fig.height=5, out.width='.49\\linewidth', fig.pos='h',fig.align='center',fig.cap="Normal quantile plots (left) and bivariate scatterplot (right) for residuals of each response from Applied Model.",fig.show='hold'>>=
#--------------------------------------------------
### Check Conditions for MANCOVA linear model fit
## check response correlations to avoid colinearity in multivariate model
#Rule of Thumb: ~ .3 to .55 then MANCOVA will work well
#with(dat, cor(AppliedCI,AppliedHT)) #Not overly correlated

### check for influencial and high leverage points
#infmeas <- influence.measures(mod1small)
#head(infmeas$infmat)
#head(infmeas$is.inf)
### no high leverage

qqp1 <- qplot(sample=mod1small$residuals[,1], stat="qq", xlab="Theoretical Normal Quantiles",
       ylab="Applied CI Score Residual") +  theme_bw()

qqp2 <- qplot(sample=mod1small$residuals[,2], stat="qq", xlab="Theoretical Normal Quantiles",
      ylab="Applied HT Score Residual") +  theme_bw()

grid.arrange(qqp1,qqp2,nrow=2)

qplot(mod1small$residuals[,1],mod1small$residuals[,2],
      xlab="Applied CI Score Residual",  ylab="Applied HT Score Residual")  +  theme_bw()
@

<<residuals2, echo=FALSE , warning=FALSE, message=FALSE, fig.width=8, fig.height=3.2, out.width='.9\\linewidth', fig.pos='H',fig.align='center',fig.cap="Applied Model residual plots overlaid with Loess smoother and corresponding 95\\% confidence envelopes.">>=
p1 <- qplot(mod1small$fitted.values[,1] ,mod1small$residuals[,1]) + 
  geom_hline(yintercept=0) +geom_jitter(position = position_jitter(height = jitheight )) +
  xlab("Applied CI Score Fitted Value") + ylab("Applied CI Score Residual") +  theme_bw() + geom_smooth()
p2 <- qplot(mod1small$fitted.values[,2] ,mod1small$residuals[,2]) + 
  geom_hline(yintercept=0) +geom_jitter(position = position_jitter(height = jitheight )) +
  xlab("Applied HT Score Fitted Value") + ylab("Applied HT Score Residual") +  theme_bw() + geom_smooth()
grid.arrange(p1,p2,nrow=1)
@

%------------------------------------------------------------------------------
\chapter{ }
%------------------------------------------------------------------------------

\section{Appendix: Lab Assignment Using Shiny Database Sampler}
\label{labappend}

For this activity you will be using a tool called the Shiny Database Sampler to take a random sample of United States residents from US census data. The census data is the Public Use Microdata Sample (PUMS) which is a 3 million person subset of the entire Census data.  For this activity we treat our samples as though they are selected from the full census records.

We are going to explore how these random sampling plans relate to the goals of a sample survey. The tool will allow you to define either a simple random sampling plan or a stratified random sampling plan. In the following two scenarios we will explore the advantages and disadvantages of these two sampling plans. Access the tool at \url{http://shiny.stat.iastate.edu/karstenm/ShinyDatabaseSampler}.

\underline{Scenario 1:} Suppose that our goal is to estimate the mean age of all US residents. Similar to polling organizations we have a budget that allows us to survey around 1000 people. To collect our sample we decide to take a simple random sample of 1040 US residents. \\

\begin{enumerate}[(a)]
\item Is this study and example of an experiment or an observational study?  Explain your answer.
\item	Your colleague Bob claims that we are wasting our budget to get only 1040 people using random sampling. He says that we could get 20000 responses to the survey if we invested that money into a mailing campaign in Minneapolis. Explain why the random selection is important.
\item	Another colleague, Jill, asks why we do not stratify by state when we take the sample so that we get 20 people from each of the 50 states along with Puerto Rico and the District of Columbia. Explain why this idea would not create a representative sample to pursue our goal.
\end{enumerate}

Now that we have decided on our sampling plan, let's go collect our data.  The Shiny Database Sampler needs to be told 4 pieces of information in order to collect census records the way you want. (1) Choose the database called ``Census'', (2) select the ``simple random sample'' option, (3) enter a random seed, any number between 1 and 10000, you can do this by rolling a 10-sided die 4 times and (4) lastly tell it that we want ``1040'' random draws. Once you have drawn your samples the page will display basic summary statistics for the variables in the census.\\

\begin{enumerate}[(a)]
\setcounter{enumi}{3}
\item Report the 5-number summary and sample mean age.
\item	Use the 5-number summary to construct a box plot of age.
\item	Go to the ``Visualize'' tab.  Choose age as your Response Variable to Plot.  What type of variable is this?  By clicking on Make My Plot? a histogram of the sample of ages will be displayed.  Describe the shape of the data distribution of age.
\item	Is the relationship between the sample mean and sample median consistent with your description of shape?  Explain briefly.
\item	If our goal was to not only estimate the mean age of all the U.S. residents but also come up with estimates of the median age of all residents in each of the 50 states, plus the District of Columbia and Puerto Rico what is a drawback of using the simple random sample of 1040?  Hint: Set the Data Table to display 100 records per page and go to the page that has ``states'' 10 and 11 (Delaware and the District of Columbia).
\end{enumerate}

\underline{Scenario 2:} Suppose now that our goal has changed.  Now we wish to investigate the association between age and state of residency. We want to compare the median ages for different states. We still have a budget that allows us to survey around 1040 people. To collect our sample we decide to take a stratified random sample of 20 residents from each state in the United States plus the District of Columbia and Puerto Rico.

\begin{enumerate}[(a)]
\setcounter{enumi}{8}
\item  Explain in general why collecting a stratified random sample is a better plan than a simple random sample for answering this question.  
\end{enumerate}

Now that we have decided on our new sampling plan, let's go collect our data.  The Shiny Database Sampler will need to be told 5 pieces of information in order to collect census records the way you want this time. (1) Choose the database called ``Census'', (2) select the ``stratified random sample'' option, (3) enter a random seed, any number between 1 and 10000, you can do this by rolling a 10-sided die 4 times, (4) select ``state'' as strata variable and (5) lastly tell it that we want ``20'' random draws from each state, plus the District of Columbia and Puerto Rico.

It will take a minute or two to collect these data. It is sifting through millions of records and randomly selecting them from within state groups after all! Once you have drawn your samples you can take a peek at your data set in the main panel of the webpage. You will be able to answer the following questions using the summaries provided on the webpage.

You will notice that the summaries are all broken down by state, but the states are not given names, they are given a code number.  This is done on the census to save computer storage space (saving a ``19'' is much smaller than ``Iowa'').  A list of all the state codes is available at \url{https://www.census.gov/geo/reference/ansi\_statetables.html} (Click on FIPS Codes for the States and the District of Columbia).

\begin{enumerate}[(a)]
\setcounter{enumi}{9}
\item Report the mean and 5-number summary for the age of the sample from the state of Iowa (\texttt{state = 19}).
\item	Report the mean and 5-number summary for the age of the sample from the state of Alaska (\texttt{state = 2}).
\item	Compare the distribution of ages in Alaska and Iowa using the values from parts j and k.
\item	Making comparisons as we have done above would become tedious if we wanted to compare ages between all pairs of states in the country.  What would be a good way to visually display this information so aid in making these comparisons? Explain your answer.
\end{enumerate}

% %------------------------------------------------------------------
% \section{Appendix: Technical Implementation of Shiny Application}
% \label{appendConstruction}
% 
% \textbf{Construction schematics and description of sampling process and plotting process.}
% 
% 
% %------------------------------------------------------------------
% \section{Appendix: Database Descriptions}
% \label{appendDB}
% 
% \textbf{To be detailed when databases updated and origins better known.}

%------------------------------------------------------------------
\section{Appendix: Cronbach's $\alpha$ Properties}
\label{appendCronbach}

Recall the form of Cronbach's $\alpha$ from equation~(\ref{eq:alpha}):\\

$\alpha = K/(K-1) \cdot \left( 1- \sum_{i=1}^K \V{Y_i} \right.\left/  \V{\sum_{j=1}^K Y_j}  \right)$

\textit{Claim 1:} Perfect agreement in items leads to $\alpha = 1$

\textit{Proof:} Let $ Y = Y_1 = Y_2 = ... = Y_k$, thus having perfect agreement.

$\Rightarrow$  $\text{Cov}(Y_i, Y_j) = \V{Y} = \sigma^2_y $ \hspace{.1in} $\forall i\ne j$ \\

$\Rightarrow$  $\V{\sum_{j=1}^K Y_j }$ = $\sum_{i=1}^K \V{Y_i} + \sum_{i\ne j}\text{Cov}(Y_i, Y_j)$ =  $K\sigma^2_y + K(K-1)\sigma^2_y$ \\

$\Rightarrow$  $\alpha = \left(K/(K-1)\right) \left( 1- \sum_{i=1}^K \V{Y_i} \right.\left/  \V{\sum_{j=1}^K Y_j}  \right)$ = \\
\indent \hspace{.2in} $\left(K/(K-1)\right) \left( 1- K\sigma^2_y \right.\left/ K\sigma^2_y + K(K-1)\sigma^2_y \right)$ =\\
\indent \hspace{.2in} $\left(K/(K-1)\right)  (1- 1/K)$ = $\left(K/(K-1)\right) ((K-1)/K)$ = $1$ \\

\vspace{.25in} %------------------------------

\textit{Claim 2:} For independent items $\alpha = 0$

\textit{Proof:} Let $Y_1 = Y_2 = ... = Y_k$ be independent

$\Rightarrow$ $\sum_{i=1}^K \V{Y_i} = \V{\sum_{j=1}^K Y_j} $

$\Rightarrow$  $\alpha = \left(K/(K-1)\right) \left( 1- \sum_{i=1}^K \V{Y_i} \right.\left/  \V{\sum_{j=1}^K Y_j}  \right)$ = \\
\indent \hspace{.2in} $\alpha = \left(K/(K-1)\right) \left( 1- \V{\sum_{j=1}^K Y_j} \right.\left/  \V{\sum_{j=1}^K Y_j}  \right)$ = \\
\indent \hspace{.2in} $\alpha = \left(K/(K-1)\right) \left( 1- 1 \right)$ = 0

\vspace{.25in} %------------------------------

\textit{Claim 3:} Perfect disagreement in items leads to $\alpha = -\infty$

\textit{Proof:} Let $K=2$ and $Y_1 = -Y_2$, thus having perfect disagreement.

$\Rightarrow$  $\V{Y_1 + Y_2}$ = $\V{Y_1 - Y_1}$ = $\V{0}$ = $0$

$\Rightarrow$  $\alpha = \left(K/(K-1)\right) \left( 1- \sum_{i=1}^K \V{Y_i} \right.\left/  \V{\sum_{j=1}^K Y_j}  \right)$ = $(2/1)(1-2\sigma^2_y/0)$ = $-\infty$

%------------------------------------------------------------------------------
\chapter{ }
%------------------------------------------------------------------------------

\section{Appendix: Optimal Offset for Symmetric Data Recorded to Resolution $\alpha_x$}
\label{proof:offset}
The following proof assumes three conditions for standard rectangular binning of a set univariate data hold: (i) data are recorded to a resolution of $\alpha_x$ units in the $X$ dimension, (ii) points are symmetric distributed within standard rectangular bins, (iii) the bin width, $\omega_x$, is an integer multiples of $\alpha_x$ (i.e. $\omega_x = k\alpha_x$ for some $k \in \{1,2,\dots\}$).  Under these conditions it will be shown that spatial loss is minimized by setting the binning origin, $\beta_x$, to $\alpha_x/2$ units below the minimum data value (i.e. $\beta_x = x_{(1)} - \alpha_x/2$).\\

Let $x_1, x_2, \dots, x_k \in \mathbb{R}$ represent the values in a single bin such that $x_{i+1} = x_i + \alpha_x$ for some constant $\alpha_x \in \mathbb{R}$. Thus $x_j = x_1 + (j-1)\alpha_x$.

Suppose then that we bin the data using standard rectangular binning with origin, $\beta_x = x_1 - \theta$, and binwidth $\omega$; where $\theta$ is the \textit{origin offset} from the data. Thus

$b(x_j) = \beta_x + \omega/2 = (x_1 - \theta) + (k\alpha_x/2)$

Spatial Loss, $L^S = \sum_{i=1}^{k} ||x_i-b_(x_i)|| $ is definitionally minimized when $b_(x_i)$ is the \textit{geometric median}. The geometric median for $x_1, \dots, x_k = Q_x(.5) = (x_{\lceil\frac{k+1}{2}\rceil}+x_{\lfloor \frac{k+1}{2}\rfloor})/2$ , where $Q_x(\cdot)$ is the empirical quantile function.

Thus the optimal offset is the $\theta$ such that

$ b(x_i) = Q_x(.5) $ \\
$ \Rightarrow (x_1 - \theta) + (k\alpha_x/2) = (x_{\lceil\frac{k+1}{2}\rceil}+x_{\lfloor \frac{k+1}{2}\rfloor})/2 $ \\
$ \Rightarrow 2x_1 - 2\theta + k\alpha_x = (x_1 + (\lceil\frac{k+1}{2}\rceil - 1)\alpha_x ) + (x_1 + (\lfloor\frac{k+1}{2}\rfloor - 1)\alpha_x ) $ \\
$ \Rightarrow -2\theta + k\alpha_x = (\lceil\frac{k+1}{2}\rceil - 1)\alpha_x  + (\lfloor\frac{k+1}{2}\rfloor - 1)\alpha_x  $ \\
$ \Rightarrow -2\theta + k\alpha_x = ((k+1) - 2)\alpha_x  $ \\
$ \Rightarrow -2\theta = -\alpha_x $ \\
$ \Rightarrow \theta = \alpha_x/2 $ \\

Thus the optimal offset for reducing spatial loss in this scenario is $\theta = \alpha_x/2$.  This result holds for data that is symmetrically distributed within the bin since the median will not change.  It extends to multiple contiguous bins with resolution $\alpha_x$ data that has symmetrically distributed data withing each bin.

If the same conditions are extended to the two dimensional case, then the origin for minimal spatial loss is at $(x_{(1)}-\alpha_x/2, y_{(1)}-\alpha_y/2)$ where $\alpha_x$ and $\alpha_y$  are the data resolution for each dimension, respectively.